---
title: "Cross-lag SEM cort-brain models"
author: "Michelle.VanTieghem"
date: "Jan 23, 2019"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

# Readme
main analyses for cross-lag SEM \
Feb 2020: adding wave1/wave2 correlations (not in model) to verify cross-sectional relationships. \
 ICV is scaled to mm^3/ 100,000  for this model \


get useful packages
```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
```

function for chi-square test later
```{r}
# is p < 0.05? test with the chi-square distribution for this DF
chi_square_test <- function(chi_sq_diff, df_diff) {
  if (chi_sq_diff > qchisq(.95, df=df_diff)){
    p  = '< 0.05'
  } else{
    p = '> 0.05' 
  }
  return(list(chi_square_difference = chi_sq_diff, df_difference = df_diff, p.value = p))
}

```

# load data 
```{r}

# this one compiled more subjects with 2 waves of data. 
load("../../../data/5_SEM_cross_lag_data/cleaned2_all_2wave_cort_brain_wide_format_ages4to19.Rdata")
#names(df_wide)

# check scaling. 
range(df_wide$ICV_wave1_rs.c, na.rm = T)
range(df_wide$Hipp_wave2_rs.c, na.rm = T)

```

# 1) check for correlations between amyg and hipp and cort
for each TP - confirming cross-sectional relationships from SEM model. 
```{r}
# correlated very highly.
amyg_hipp_cor1 <- cor.test(df_wide$Amyg_wave1, df_wide$Hipp_wave1)
amyg_hipp_cor1
save(amyg_hipp_cor1, file = "model_output/amyg_hipp_correlation_wave1.Rdata")
amyg_hipp_cor2 <- cor.test(df_wide$Amyg_wave2, df_wide$Hipp_wave2)
amyg_hipp_cor2
save(amyg_hipp_cor2, file = "model_output/amyg_hipp_correlation_wave2.Rdata")

# not correlated
amyg_cort_cor1 <- cor.test(df_wide$Amyg_wave1, df_wide$waking_wave1)
amyg_cort_cor1
save(amyg_cort_cor1, file = "model_output/amyg_waking_correlation_wave1.Rdata")
amyg_cort_cor2 <- cor.test(df_wide$Amyg_wave2, df_wide$waking_wave2)
amyg_cort_cor2 
save(amyg_cort_cor2, file = "model_output/amyg_waking_correlation_wave2.Rdata")

# trend correlated
hipp_cort_cor1 <- cor.test(df_wide$Hipp_wave1, df_wide$waking_wave1)
hipp_cort_cor1
save(hipp_cort_cor1, file = "model_output/hipp_waking_correlation_wave1.Rdata")
# not corelated 
hipp_cort_cor2 <- cor.test(df_wide$Hipp_wave2, df_wide$waking_wave2)
hipp_cort_cor2
save(hipp_cort_cor2, file = "model_output/hipp_waking_correlation_wave2.Rdata")

```


# 2) cross-lagged SEM Cort & Hipp 
## basic model with cross-lag paths 

note on standardized coefficients
Two extra columns of standardized parameter values are printed. In the first column (labeled Std.lv), only the latent variables are standardized. In the second column (labeled Std.all), both latent and observed variables are standardized. The latter is often called the 'completely standardized solution'.
```{r }

model= 
'
#Regression paths cross-lagged
Hipp_wave2_rs.c ~  Hipp_wave1_rs.c + waking_wave1.c + Age_wave2.c +  IDENT_SUBTYPE + sex.c + ICV_wave2_rs.c + scanner_wave2
waking_wave2.c ~ Hipp_wave1_rs.c + waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE + sex.c 

# Regression paths at T1 
Hipp_wave1_rs.c ~ Age_wave1.c + IDENT_SUBTYPE  +  sex.c + ICV_wave1_rs.c 
waking_wave1.c ~ Age_wave1.c + IDENT_SUBTYPE +  sex.c

# covariance
waking_wave1.c ~~ Hipp_wave1_rs.c
waking_wave2.c ~~ Hipp_wave2_rs.c

# time 1 - time 2 things
ICV_wave1_rs.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Age_wave2.c 

# correlations at each time point 
Age_wave1.c ~ ICV_wave1_rs.c 
Age_wave2.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Hipp_wave1_rs.c 
Age_wave2.c ~ Hipp_wave2_rs.c
Age_wave1.c ~ waking_wave1.c
Age_wave2.c ~ waking_wave2.c
#sex.c	~	Hipp_wave1_rs.c		

ICV_wave1_rs.c	~	Hipp_wave1_rs.c	
ICV_wave2_rs.c	~	Hipp_wave2_rs.c	


IDENT_SUBTYPE	~	Hipp_wave1_rs.c
#IDENT_SUBTYPE ~ sex.c 
IDENT_SUBTYPE ~ ICV_wave1_rs.c
'

hipp_mod1 <- sem(model, data=df_wide, fixed.x=F, missing='fiml')

fit_measures <- data.frame(fitMeasures(hipp_mod1)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "hipp_mod1"

hipp_mod1_fit_measures <- fit_measures %>%
  spread(metric, value)

summary(hipp_mod1, fit.measures=TRUE, standardized=TRUE, rsquare=TRUE)
# parameterEstimates(hipp_mod1) just useful for confidence intervals.

```

## testing model without hipp1 --> waking 2
```{r }

model= 
'
#Regression paths cross-lagged
Hipp_wave2_rs.c ~  Hipp_wave1_rs.c + waking_wave1.c + Age_wave2.c +  IDENT_SUBTYPE + sex.c + ICV_wave2_rs.c + scanner_wave2
waking_wave2.c ~  waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE + sex.c 

# Regression paths at T1 
Hipp_wave1_rs.c ~ Age_wave1.c + IDENT_SUBTYPE  +  sex.c + ICV_wave1_rs.c 
waking_wave1.c ~ Age_wave1.c + IDENT_SUBTYPE +  sex.c 

# covariance
waking_wave1.c ~~ Hipp_wave1_rs.c
waking_wave2.c ~~ Hipp_wave2_rs.c

# time 1 - time 2 things
ICV_wave1_rs.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Age_wave2.c 

# correlations at each time point 
Age_wave1.c ~ ICV_wave1_rs.c 
Age_wave2.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Hipp_wave1_rs.c 
Age_wave2.c ~ Hipp_wave2_rs.c
Age_wave1.c ~ waking_wave1.c
Age_wave2.c ~ waking_wave2.c
#sex.c	~	Hipp_wave1_rs.c		

ICV_wave1_rs.c	~	Hipp_wave1_rs.c	
ICV_wave2_rs.c	~	Hipp_wave2_rs.c		

IDENT_SUBTYPE	~	Hipp_wave1_rs.c
#IDENT_SUBTYPE ~ sex.c 
IDENT_SUBTYPE ~ ICV_wave1_rs.c
'

hipp_mod2 <- sem(model, data=df_wide, fixed.x=F, missing='fiml')

fit_measures <- data.frame(fitMeasures(hipp_mod2)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "hipp_mod2"

hipp_mod2_fit_measures <- fit_measures %>%
  spread(metric, value)
```

## chisquare difference test
adding hipp wave1 --> cort wave 2 significantly improves model fit

```{r}
hipp_mod1_fit_measures$chisq # model with hipp has better (lower) chi-square
hipp_mod2_fit_measures$chisq # model without hipp has higher (worse) chi-square 

# is this difference significant?
diff_chi_sq <- abs(hipp_mod2_fit_measures$chisq - hipp_mod1_fit_measures$chisq)
diff_df <-  abs(hipp_mod2_fit_measures$df - hipp_mod1_fit_measures$df)

# yes!
chi_square_hipp_result <- chi_square_test(diff_chi_sq, diff_df)
chi_square_hipp_result
save(chi_square_hipp_result, file = "model_output/hipp_cross_lag_chi_sq_test.Rdata")

```


# 3) cross-lagged SEM Cort & Amyg 
## basic model
```{r }
df_wide <- df_wide %>%
  mutate(scanner_wave2.c = scanner_wave2 - mean(scanner_wave2, na.rm = T),
         GROUP.n =ifelse(df_wide$GROUP == "PI", 1, 0), 
         GROUP.c = GROUP.n - mean(GROUP.n))
model= 
'
#Regression paths cross-lagged
Amyg_wave2_rs.c ~  Amyg_wave1_rs.c + waking_wave1.c + Age_wave2.c +  IDENT_SUBTYPE + sex.c + ICV_wave2_rs.c + scanner_wave2
waking_wave2.c ~ Amyg_wave1_rs.c + waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE + sex.c 

# Regression paths at T1 
Amyg_wave1_rs.c ~ Age_wave1.c + IDENT_SUBTYPE  +  sex.c + ICV_wave1_rs.c 
waking_wave1.c ~ Age_wave1.c + IDENT_SUBTYPE +  sex.c 

# covariance
waking_wave1.c ~~ Amyg_wave1_rs.c
waking_wave2.c ~~ Amyg_wave2_rs.c

# time 1 - time 2 things
ICV_wave1_rs.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Age_wave2.c 

# correlations at each time point 
Age_wave1.c ~ ICV_wave1_rs.c 
Age_wave2.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Amyg_wave1_rs.c 
Age_wave2.c ~ Amyg_wave2_rs.c
Age_wave1.c ~ waking_wave1.c
Age_wave2.c ~ waking_wave2.c
#sex.c	~	Amyg_wave1_rs.c		

ICV_wave1_rs.c	~	Amyg_wave1_rs.c	
ICV_wave2_rs.c	~	Amyg_wave2_rs.c		

IDENT_SUBTYPE	~	Amyg_wave1_rs.c
#IDENT_SUBTYPE ~ sex.c 
IDENT_SUBTYPE ~ ICV_wave1_rs.c
'

amyg_mod1 <- sem(model, data=df_wide, fixed.x=F, missing='fiml')

fit_measures <- data.frame(fitMeasures(amyg_mod1)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "amyg_mod1"

amyg_mod1_fit_measures <- fit_measures %>%
  spread(metric, value)

summary(amyg_mod1, fit.measures=TRUE, standardized=TRUE, rsquare=TRUE)
#standardizedSolution(amyg_mod1)
#parameterEstimates(amyg_mod1) #just useful for confidence intervals.
```

## testing model without amyg1 --> cort 2
```{r }

model= 
'
#Regression paths cross-lagged
Amyg_wave2_rs.c ~  Amyg_wave1_rs.c + waking_wave1.c + Age_wave2.c +  IDENT_SUBTYPE + sex.c + ICV_wave2_rs.c + scanner_wave2 
waking_wave2.c ~  waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE + sex.c 

# Regression paths at T1 
Amyg_wave1_rs.c ~ Age_wave1.c + IDENT_SUBTYPE  +  sex.c + ICV_wave1_rs.c 
waking_wave1.c ~ Age_wave1.c + IDENT_SUBTYPE +  sex.c 

# covariance
waking_wave1.c ~~ Amyg_wave1_rs.c
waking_wave2.c ~~ Amyg_wave2_rs.c

# time 1 - time 2 things
ICV_wave1_rs.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Age_wave2.c 

# correlations at each time point 
Age_wave1.c ~ ICV_wave1_rs.c 
Age_wave2.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Amyg_wave1_rs.c 
Age_wave2.c ~ Amyg_wave2_rs.c
Age_wave1.c ~ waking_wave1.c
Age_wave2.c ~ waking_wave2.c
#sex.c	~	Amyg_wave1_rs.c		

ICV_wave1_rs.c	~	Amyg_wave1_rs.c	
ICV_wave2_rs.c	~	Amyg_wave2_rs.c		

IDENT_SUBTYPE	~	Amyg_wave1_rs.c
#IDENT_SUBTYPE ~ sex.c 
IDENT_SUBTYPE ~ ICV_wave1_rs.c
'

amyg_mod2 <- sem(model, data=df_wide, fixed.x=F, missing='fiml')

fit_measures <- data.frame(fitMeasures(amyg_mod2)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "amyg_mod2"

amyg_mod2_fit_measures <- fit_measures %>%
  spread(metric, value)
```


## Chi-square test of model fit 
including amyg wave1 to predict waking wave 2 significantly improves model fit!
```{r}
amyg_mod1_fit_measures$chisq # model with amyg has better (lower) chi-square
amyg_mod2_fit_measures$chisq # model without amyg has higher (worse) chi-square 

# is this difference significant?
diff_chi_sq <- abs(amyg_mod2_fit_measures$chisq - amyg_mod1_fit_measures$chisq)
diff_df <-  abs(amyg_mod2_fit_measures$df - amyg_mod1_fit_measures$df)


# yes!
chi_square_amyg_result <- chi_square_test(diff_chi_sq, diff_df)
chi_square_amyg_result
save(chi_square_amyg_result, file = "model_output/amyg_cross_lag_chi_sq_test.Rdata")

```



# save final models 

```{r}
save(amyg_mod1, file = "model_output/amyg_waking_cross_lagged_model_main_effects.Rdata")

save(hipp_mod1, file = "model_output/hipp_waking_cross_lagged_model_main_effects.Rdata")

```


