---
title: "QC check and removing outliers"
author: "Michelle.VanTieghem"
date: "July 5, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# README:
same QC procedure as original analysis, for subset of data (6-19)
1. get freesurferand aseg QC info together 
2. calculate inter rater reliablity for motion and aseg ratings 
3. get age distribution of all scans from comps + PIs
4. filter based on scan quality
5. once all excluded bad brains, check age-distribution and exclude too old/ too young comps.
6. calculate outliers witin that final sample. 

# final age filter
6.0 minimum to 19.0 maximum 
1 PI < 6 and 0 PIs > 19


 
```{r, include = F, warnings =F, echo = F}
source("../../0_R_analysis_setup_file.R")

```


# load data 
- not filtered at all yet for brain_age_yrs or anything. 
```{r, include = F}
load("../../../data/2_fs_data/1_all_freesurfer_v6_data_with_aseg_QC_info.RData")
nrow(fsdata) # not filtered at all yet! 
```


## keep only comps and PIs with brain data.
```{r}

fsdata <- subset(fsdata, !is.na(brain_age_yrs) &
                       (GROUP == 1 | GROUP == 0))

nrow(fsdata) # after filtering 
```

## Check whole brain volume measures 

```{r,warnings=FALSE}
fsdata$ICV <- fsdata$EstimatedTotalIntraCranialVol
fsdata$GMV <- fsdata$TotalGrayVol
# don't have whole brain volume from ASEG because no white matter estimates. .
```


# calculate metrics for QC
## check motion metrics 
```{r, warnings = F}

motion_vars <- fsdata[grepl("motion", names(fsdata))]
motion_vars <- motion_vars[,1:3 ]
class(motion_vars$motion_Michelle)
class(motion_vars$motion_Jasmin)

motion_vars$motion_Jasmin.n <- as.numeric(substr(motion_vars$motion_Jasmin, 1, 1))
motion_vars$motion_Marta.n <- as.numeric(substr(motion_vars$motion_Marta, 1, 1))
motion_vars$motion_Michelle.n <-as.numeric(substr(motion_vars$motion_Michelle, 1, 1))

range(motion_vars$motion_Jasmin.n, na.rm = T)
range(motion_vars$motion_Marta.n, na.rm = T)
range(motion_vars$motion_Michelle.n, na.rm = T)

# fix this one... 
#motion_vars$motion_Marta.n <- ifelse(motion_vars$motion_Marta.n == 34, 3.5, #as.numeric((motion_vars$motion_Marta.n)))

# now get the means and add to df
motion_vars.m <- as.matrix(motion_vars[,4:6])
fsdata$motion_ave_new <- rowMeans(motion_vars.m, na.rm = TRUE)
hist(fsdata$motion_ave_new)

```

## check motion rater reliability - excellent! 
```{r, warnings = F}
# correlation 
corr_motion <- with(motion_vars, cor.test(motion_Marta.n, motion_Michelle.n))
interclass_coef <- corr_motion[4]
# Intra-class correlatin
# http://personality-project.org/r/html/ICC.html
#library(psych)
#ICC(motion_vars.m)

# how many overlap (have 2 ratings)
N_both_MM_raters <- with(motion_vars, sum(ifelse(!is.na(motion_Marta.n) & 
                              !is.na(motion_Michelle.n), 1, 0), na.rm = T))
# all brains rated by both raters except for 1! 
N_both_MM_raters

# calculate percent raters agreement
N_motion_MM_raters_agree <- sum(with(motion_vars, ifelse(motion_Marta.n == motion_Michelle.n, 1, 0)), na.rm = TRUE)

# only agree 56% of the time... 
Motion_MM_inter_rater_reliability <- N_motion_MM_raters_agree/ N_both_MM_raters

####
# how many overlap (have 2 ratings)
N_both_MJ_raters <- with(motion_vars, sum(ifelse(!is.na(motion_Marta.n) & 
                              !is.na(motion_Jasmin.n), 1, 0), na.rm = T))
# all brains rated by both raters except for 1! 
N_both_MJ_raters

# calculate percent raters agreement
N_motion_MJ_raters_agree <- sum(with(motion_vars, ifelse(motion_Marta.n == motion_Jasmin.n, 1, 0)), na.rm = TRUE)

# only agree 57% of the time... 
Motion_MJ_inter_rater_reliability <- N_motion_MJ_raters_agree/ N_both_MJ_raters


```



## check aseg metrics 
```{r, warnings = F}
## check aseg metrics 
aseg_vars <- fsdata[grepl("ASEG", names(fsdata))]
names(aseg_vars)
aseg_vars <- cbind(aseg_vars[1], aseg_vars[3])
class(aseg_vars$ASEG_rating_Marta)
class(aseg_vars$ASEG_rating_Liliana)

aseg_vars$ASEG_rating_Liliana.n <- as.numeric(substr(aseg_vars$ASEG_rating_Liliana, 1, 1))
aseg_vars$ASEG_rating_Marta.n <- as.numeric(substr(aseg_vars$ASEG_rating_Marta, 1, 1))

head(aseg_vars$ASEG_rating_Liliana.n)
range(aseg_vars$ASEG_rating_Marta.n, na.rm = T)
range(aseg_vars$ASEG_rating_Liliana.n, na.rm = T)
## calculate averbrain_age_yrs aseg rating 
aseg_vars.m <- as.matrix(aseg_vars[,3:4])
fsdata$aseg_ave_rating <- rowMeans(aseg_vars.m, na.rm = TRUE)
hist(fsdata$aseg_ave_rating)

```

## check ASEG rater reliability - excellent! 
note this is only within the included subjects. 
```{r, warnings = F}
# correlation 
corr_aseg <- with(aseg_vars, cor.test(ASEG_rating_Marta.n, ASEG_rating_Liliana.n))
interclass_coef <- corr_aseg[4]

# intraclass correlation
#ICC(aseg_vars.m)

# how many overlap (have 2 ratings)
N_both_raters_aseg <- with(aseg_vars, sum(ifelse(!is.na(ASEG_rating_Marta.n) & 
                              !is.na(ASEG_rating_Liliana.n), 1, 0), na.rm = T))
# all brains rated by both raters except for 1! 


# calculate percent raters agreement
N_aseg_raters_agree <- sum(with(aseg_vars, ifelse(ASEG_rating_Liliana.n == ASEG_rating_Marta.n, 1, 0)), na.rm = TRUE)

# 82 % rater reliablity. 
ASEG_inter_rater_reliability <- N_aseg_raters_agree/ N_both_raters_aseg


table_motion_info <- rbind(N_both_MM_raters, N_motion_MM_raters_agree, Motion_MM_inter_rater_reliability, 
                           N_both_MJ_raters, N_motion_MJ_raters_agree, Motion_MJ_inter_rater_reliability, 
                           N_both_raters_aseg, N_aseg_raters_agree, ASEG_inter_rater_reliability)

save(table_motion_info, file = "tables/Interrater_reliability_freesurfer_QC_assess.Rdata")
```



## exclude subjets with no brain data & check whether all brains have motion assessment 
```{r, warnings = F}
fsdata$brain_check <- as.factor(with(fsdata, 
                               ifelse(is.na(TotalGrayVol) & !is.na(motion_ave_new), 
                                      "missing_volume", 
                                ifelse(!is.na(TotalGrayVol) & is.na(motion_ave_new), 
                                       "missing_motion", 
                                       ifelse(is.na(TotalGrayVol) & is.na(motion_ave_new), 
                                              "no_brain_data","good")))))
summary(fsdata$brain_check)

missing_motion_index <- which(fsdata$brain_check == "missing_motion")
missing_motion <- subset(fsdata, brain_check == "missing_motion")
missing_motion$SUBJECTID_long
# SB066 should be excluded

fsdata1 <- subset(fsdata, brain_check != "missing_motion")

N_all_COMP_scans <- summary(fsdata1$GROUP)[1]
N_all_PI_scans <- summary(fsdata1$GROUP)[2]

```

# Get sample info on all scanned subjects 
## Add group info 
```{r}

fsdata1$GROUP <- as.factor(ifelse(fsdata1$GROUP == 1, "PI", "COMP"))
summary(fsdata1$GROUP)
fsdata1$sex <- fsdata1$DEM_3_GENDER_CHILD
summary(as.factor(fsdata1$sex))
```


## find out how many young PIS vs. comps 
for all subjects - before excluding anyone!! 
```{r}

fsdata1$brain_age_yrs_bin <- as.factor(ifelse(fsdata1$brain_age_yrs < 7,"4-6.9", ifelse(fsdata1$brain_age_yrs < 10, "7-9.9", 
                          ifelse(fsdata1$brain_age_yrs < 13,"10-12.9", ifelse(fsdata1$brain_age_yrs < 16, "13-15.9", 
                                                                    ifelse( fsdata1$brain_age_yrs >= 16, "16-20", "help"))))))

fsdata1$brain_age_yrs_bin <- factor(fsdata1$brain_age_yrs_bin, levels = c("4-6.9", "7-9.9", "10-12.9", "13-15.9", "16-20"))
fsdata1$brain_age_yrs_Factor <- as.factor(trunc(fsdata1$brain_age_yrs))

brain_age_yrs_bin_groups <- fsdata1 %>%
  group_by(GROUP, brain_age_yrs_bin) %>%
  dplyr::summarize(n = n(),
            mean_brain_age_yrs= mean(brain_age_yrs),
            min_brain_age_yrs = min(brain_age_yrs),
            max_brain_age_yrs = max(brain_age_yrs))

brain_age_yrs_bin_groups
```

## plot group bins
```{r}
brain_age_yrs_bin_groups <- fsdata1 %>%
  group_by(GROUP, brain_age_yrs_Factor) %>%
  dplyr::summarize(n = n(),
            mean_brain_age_yrs= mean(brain_age_yrs), 
            max_brain_age_yrs = max(brain_age_yrs),
            min_brain_age_yrs = min(brain_age_yrs))


ggplot(brain_age_yrs_bin_groups, aes(x = brain_age_yrs_Factor, y = n, fill = GROUP)) +
  geom_bar(stat = "identity") + theme_classic() + ylab("N subjects")

```

## get min brain_age_yrs for reference later
```{r}
fsdata1$index_wave <- as.numeric(as.character(fsdata1$index_wave))
fsdata1$GROUP.n <- as.numeric(as.character(fsdata1$GROUP))
sum(is.na(fsdata1$SUBJECTID.y))
fsdata1$IDENT_SUBID<- fsdata1$SUBJECTID.y

# get min brain_age_yrs for each subject. 
getinfo <- fsdata1 %>%
  group_by(IDENT_SUBID) %>%
  dplyr::summarize(GROUP = mean(GROUP.n),
            sex = mean(sex),
            first_visit = min(index_wave), 
            last_visit=max(index_wave), 
            max_brain_age_yrs = max(brain_age_yrs),
            min_brain_age_yrs = min(brain_age_yrs))


#now get info for each subject included in dataset.
Indiv_subject_table <- getinfo %>%
  group_by(GROUP) %>%
  dplyr::summarize(N= n(),
            mean_min_brain_age_yrs = mean(min_brain_age_yrs),  
            proportion_female = mean(sex))



# now add this info to the long-dataset for future reference! 
for (i in 1:nrow(fsdata1)){
  sub <- fsdata1$IDENT_SUBID[i]
  fsdata1$first_visit[i] <- getinfo$first_visit[getinfo$IDENT_SUBID == sub]
  fsdata1$last_visit[i] <- getinfo$last_visit[getinfo$IDENT_SUBID == sub]
  fsdata1$min_brain_age_yrs[i] <- getinfo$min_brain_age_yrs[getinfo$IDENT_SUBID== sub]
  fsdata1$max_brain_age_yrs[i] <- getinfo$max_brain_age_yrs[getinfo$IDENT_SUBID== sub]
}
```

# Exclude subjects based on QC metrics
## exclude subjects with motion ratings > 3
```{r, warnings = F}
nrow(fsdata1)
# first get data on who they are


fsdata1$exclude_fs_motion <- ifelse(fsdata1$motion_ave_new > 3, 1, 0)
N_excluded_motion <- sum(fsdata1$exclude_fs_motion, na.rm = T)
N_excluded_motion_PI <-  sum(fsdata1$exclude_fs_motion[fsdata1$GROUP == "PI"], na.rm = T)
N_excluded_motion_COMP <-  sum(fsdata1$exclude_fs_motion[fsdata1$GROUP == "COMP"], na.rm = T)


# not excluding ANY PIs < 6 years old here! 
fsdata1 %>% group_by(GROUP, exclude_fs_motion) %>%
  dplyr::summarize(n = n(), 
            mean_brain_age_yrs = mean(brain_age_yrs), 
            max_brain_age_yrs = max(brain_age_yrs),
            min_brain_age_yrs = min(brain_age_yrs))

                          
# exclude subjects marked 1 
fsdata2 <- subset(fsdata1, exclude_fs_motion == 0 | is.na(exclude_fs_motion))
N_brain_data_pts_usable_motion <- nrow(fsdata2)

```


## exclude others marked "exclude" (from freesurer proc MVT) if any 
```{r, warnings = F}
# we forgot to mark SB001_fu1 as exclude - data is corrupted
fsdata2$inclusion[fsdata2$SUBJECTID_long == "SB001_fu1"] <- "EXCLUDE"


summary(as.factor(fsdata2$inclusion))
# mark those 4 subjects and exclude them.
fsdata2$exclude_fs_other <- ifelse(fsdata2$inclusion == "EXCLUDE", 1, 0)

N_excluded_other <- sum(fsdata2$exclude_fs_other)
N_excluded_other_PI <- sum(fsdata2$exclude_fs_other[fsdata2$GROUP == "PI"])
N_excluded_other_COMP <- sum(fsdata2$exclude_fs_other[fsdata2$GROUP == "COMP"])

# get info on these subjects 
# not excluding ANY PIs < 6 years old here! 
fsdata2 %>% group_by(GROUP, exclude_fs_other) %>%
  dplyr::summarize(n = n(), 
            mean_brain_age_yrs = mean(brain_age_yrs), 
            max_brain_age_yrs = max(brain_age_yrs),
            min_brain_age_yrs = min(brain_age_yrs))

# now exclude subjects marked. 
fsdata3 <- subset(fsdata2, inclusion != "EXCLUDE")
nrow(fsdata3)
```

## of the brains remaining, how many have bad aseg rating ? 
```{r, warnings = F}
# 4 subjects with scores greater than 3.
fsdata3$excluded_aseg <-  ifelse(fsdata3$aseg_ave_rating > 3, 1,0)
N_excluded_aseg_PI <- sum(fsdata3$excluded_aseg[fsdata3$GROUP == "PI"], na.rm = TRUE)
N_excluded_aseg_COMP <- sum(fsdata3$excluded_aseg[fsdata3$GROUP == "COMP"], na.rm = TRUE)

# keep subjects with rating of 3 or lower. 
fsdata4 <- subset(fsdata3, aseg_ave_rating <= 3)
nrow(fsdata4)

```

## make average amyg and hipp vars 
```{r}

fsdata4$Amyg_ave <- (fsdata4$Left.Amygdala + fsdata4$Right.Amygdala) / 2 
fsdata4$Hipp_ave <- (fsdata4$Left.Hippocampus + fsdata4$Right.Hippocampus) / 2 
#Left.Hippocampus
fsdata4$Amyg_ratio_ICV <- fsdata4$Amyg_ave / fsdata4$ICV
fsdata4$Amyg_ratio_GMV <- fsdata4$Amyg_ave / fsdata4$GMV
```


## check brain_age_yrs distributions of the entire sample ...
```{r}

fsdata4$brain_age_yrs_bin <- as.factor(ifelse(fsdata4$brain_age_yrs < 7,"4-6.9", ifelse(fsdata4$brain_age_yrs < 10, "7-9.9", 
                          ifelse(fsdata4$brain_age_yrs < 13,"10-12.9", ifelse(fsdata4$brain_age_yrs < 16, "13-15.9", 
                                                                    ifelse( fsdata4$brain_age_yrs < 19, "16-19", "older 19"))))))

fsdata4$brain_age_yrs_bin <- factor(fsdata4$brain_age_yrs_bin, levels = c("4-6.9", "7-9.9", "10-12.9", "13-15.9", "16-19", "older 19"))

fsdata4$brain_age_yrs_Factor <- as.factor(trunc(fsdata4$brain_age_yrs))
brain_age_yrs_bin_groups <- fsdata4 %>%
  group_by(GROUP,brain_age_yrs_bin) %>%
  dplyr::summarize(n = n(),
            mean_brain_age_yrs= mean(brain_age_yrs), 
            max_brain_age_yrs = max(brain_age_yrs),
            min_brain_age_yrs = min(brain_age_yrs))

brain_age_yrs_bin_groups

```




### excluding bottom of brain_age_yrs-distribution in order to even-out distribution between groups 
 want at least 3 per group at the brain_age_yrs-extremes. 
```{r}
fsdata5 <- subset(fsdata4, brain_age_yrs >= 6 & brain_age_yrs <= 19) 
excluded_for_age_PI <- subset(fsdata4, (brain_age_yrs < 6 | brain_age_yrs > 19) & GROUP == "PI") 
N_excluded_brain_age_yrs_PI <- nrow(excluded_for_age_PI)

excluded_for_age_COMP <- subset(fsdata4, (brain_age_yrs < 6 | brain_age_yrs > 19 ) & GROUP == "COMP") 
N_excluded_brain_age_yrs_COMP <- nrow(excluded_for_age_COMP)


N_sample_pre_outliers <- nrow(fsdata5)
```



## plot the new  brain_age_yrs distribution 
still very uneven! 
```{r}
brain_age_yrs_bin_groups <- fsdata5 %>%
  group_by(GROUP, brain_age_yrs_Factor) %>%
  dplyr::summarize(n = n(),
            mean_brain_age_yrs= mean(brain_age_yrs), 
            max_brain_age_yrs = max(brain_age_yrs),
            min_brain_age_yrs = min(brain_age_yrs))


ggplot(brain_age_yrs_bin_groups, aes(x = brain_age_yrs_Factor, y = n, fill = GROUP)) +
  geom_bar(stat = "identity") + theme_classic() + ylab("N subjects")
```


# now check for outliers by TP and group. 
note: outliers by wave were already excluded, but this is a subsample... 
```{r}

fsdata5 <- fsdata5[
  with(fsdata5, order(GROUP, index_wave)),
]

# select variables you want outliers for
var_list <- fsdata5 %>% dplyr::select(Amyg_ave, Hipp_ave, ICV, GMV)
var_list <- names(var_list)
# make an empty dataframe to store new stuff into 
outlier_list <- NA
# run through loop of each variable 
for (var in var_list){
    print (var)
  # use dplyr magic to get all of the info you need.
  outlier_groups <- fsdata4 %>% 
    # group by GROUP and TIMEPOINT.
      dplyr::select (var, GROUP, index_wave) %>% 
         group_by(GROUP, index_wave) %>% # sort data by group and tp 
         summarize_all(funs(mean(., na.rm = T), sd(., na.rm = T)))         
    # get the means, SD, upper and lower limit (3 SD above or below mean) for 
   outlier_groups$upper <- outlier_groups$mean + 3*outlier_groups$sd 
   outlier_groups$lower <-  outlier_groups$mean - 3*outlier_groups$sd 
   
   # restart these here, because they will generated anew for each variable.

    outlier_list <- data.frame()
# for each grouping in outlier_groups, find outliers and mark them. 
for (i in 1:nrow(outlier_groups)){
  print(i)
  # which group is this? PI or comp
  group <- outlier_groups$GROUP[i]
  tp <- outlier_groups$index_wave[i]
  # set the max and minimum values of cortisol that we will use as outlier threshold
  lower_limits <- outlier_groups$lower[i]
  upper_limits <- outlier_groups$upper[i]
  # get all of the original data by group & wave
  data_chunk <- subset(fsdata5,  GROUP== group & index_wave == tp)
 # make sure you're only calculating outliers for the correct variable 
  variable <- data_chunk %>% dplyr::select(var)
   # mark the outliers as 1
  mark_outliers <- data.frame(ifelse(variable[1] >= upper_limits
                          | variable <= lower_limits, 1, 0))
   # make sure this name matches the variable name (e.g. L hipp)
  names(mark_outliers) <- paste0(var, "_outliers")
  # save the outlier 1s and 0s into a new column 
  outlier_list <- rbind(outlier_list, mark_outliers)
  # make sure the column name matches the variable name (e.g. L hipp)
  names(outlier_list) <- paste0(var, "_outliers")
} 
  print ("loop break")
  # add the new column of outliers for each variable (e.g. L hipp) to the orig dataframe. 
   fsdata5 <- cbind(fsdata5, outlier_list) 
   
}
```

## count outliers by group and wave
```{r}
  
# use dplyr magic to get all of the info you need.
sample_outlier_table <- fsdata5 %>%    
  
         group_by(GROUP, index_wave) %>% # sort data by group and wave 
         dplyr::summarize(N= n(), 
                   Mean_amyg = mean(Amyg_ave), 
                   SD_amyg=sd(Amyg_ave),
           amyg_outliers = sum(Amyg_ave_outliers),
             Mean_hipp = mean(Hipp_ave), 
                   SD_hipp =sd(Hipp_ave),
           hipp_outliers = sum(Hipp_ave_outliers))


sample_outlier_table

save(sample_outlier_table, file = "tables/outliers_fsdata_growthchart_ages_6_19.Rdata")
```




## remove outliers from dataset 
```{r, warnings = F}
fsdata5$gender <- with(fsdata5, as.factor(ifelse(sex == 0, "Male", "Female")))

fsdata6 <- subset(fsdata5, Amyg_ave_outliers !=1 
                     & Hipp_ave_outliers != 1 & ICV_outliers != 1)
nrow(fsdata5)
nrow(fsdata6) # final sample

```

# final sample

## make table of all excluded subjects and reasons 
```{r}


N_PI_final <- summary(fsdata6$GROUP)[2]
N_COMP_final <- summary(fsdata6$GROUP)[1]

N_amyg_outliers_PI <- sum(fsdata5$Amyg_ave_outliers[fsdata5$GROUP == "PI"])
N_amyg_outliers_COMP <- sum(fsdata5$Amyg_ave_outliers[fsdata5$GROUP == "COMP"])

N_hipp_outliers_PI <- sum(fsdata5$Hipp_ave_outliers[fsdata5$GROUP == "PI"])
N_hipp_outliers_COMP <- sum(fsdata5$Hipp_ave_outliers[fsdata5$GROUP == "COMP"])

PI_excluded_scans <- rbind(N_all_PI_scans,N_excluded_motion_PI,N_excluded_other_PI, N_excluded_aseg_PI, N_excluded_brain_age_yrs_PI, N_PI_final) 
COMP_excluded_scans <- rbind(N_all_COMP_scans,N_excluded_motion_COMP,N_excluded_other_COMP, N_excluded_aseg_COMP, N_excluded_brain_age_yrs_COMP, N_COMP_final) 
                    
exclusion_table_fsdata_growthchart_ages_6_19 <- cbind(PI_excluded_scans, COMP_excluded_scans)

rownames(exclusion_table_fsdata_growthchart_ages_6_19 ) <- c("all scans", 
                                                             "excluded_motion","excluded_other", 
                                                             "excluded_aseg", "excluded_age", "final N scans")

save(exclusion_table_fsdata_growthchart_ages_6_19, file = "tables/exclusion_table_fsdata_growthchart_ages_6_19.Rdata")
```

## get demographics ! 
```{r}

table <- fsdata6 %>%
  group_by(GROUP, index_wave) %>%
  dplyr::summarize(N= n(), 
            mean_brain_age_yrs = mean(brain_age_yrs), 
            sd_brain_age_yrs = sd(brain_age_yrs),
            min_brain_age_yrs = min(brain_age_yrs), 
            max_brain_age_yrs = max(brain_age_yrs), 
            proportion_female = mean(sex))

table

save(table, file = "tables/demographics_fsdata_growthchart_ages_6_19.Rdata")

```


# age graph for final sample

## plot age across waves for each group 
```{r}
# get min age for each subject. 
min_age <- fsdata6 %>%
  group_by(IDENT_SUBID) %>%
  dplyr::summarize(min_age = min(brain_age_yrs))
for (i in 1:nrow(fsdata6)){
  sub <-fsdata6$IDENT_SUBID[i]
  fsdata6$min_age[i] <- min_age$min_age[min_age$IDENT_SUBID== sub]
}

#hist(as.numeric(cort_merge7.no.outliers$min_age))
```

```{r}

my_colors <-  scale_color_manual(values=c("#0006CC", "#CC0000"))
my_colors2 <- scale_fill_manual(values=c("#0006CC", "#CC0000")) 
AgebyParticipant_Brain <-ggplot(fsdata6,
                         aes(colour=GROUP,
                             y=fct_reorder(IDENT_SUBID, min_age), x=brain_age_yrs))+
  geom_line(aes(group = IDENT_SUBID), size=.7, alpha = 0.7) + 
  ylab("Participants") + xlab("Age") +
  geom_point(size=1,  alpha = 0.7) + xlim(3, 20) +
  theme_classic() + my_colors + facet_grid (~ GROUP) +
  theme(axis.line = element_line(colour = "black"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position="none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) + 
   scale_x_continuous(breaks = seq(6,19, by = 3), limits=c(6, 19)) 


pdf(file = "figures/Manuscript_Figure2_AgebyParticipant_brain_6to19.pdf", width = 6, height =4)
AgebyParticipant_Brain
dev.off()

AgebyParticipant_Brain
```


# save data
```{r}
save(fsdata6, file="../../../data/2_fs_data/3_cleaned_fsdata_for_growth_chart_brain_age_yrs6_19.Rdata")
```
