---
title: 'Checking confound variables: motion, aseg, scanner confound, ICV/GMV'
author: "Michelle VanTieghem"
date: "January 23, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# Readme:
following same procedure for descriptives, with subset of scans (6-19)
 
```{r, include = F, warnings =F, echo = F}
source("../../0_R_analysis_setup_file.R")

```

# load data
```{r}
load("../../../data/2_fs_data/3_cleaned_fsdata_for_growth_chart_brain_age_yrs6_19.Rdata")

```


# check correlations among variables 
## hemisphere correlations 
```{r, warnings = F}
with(fsdata6, cor.test(Right.Hippocampus, Left.Hippocampus))
with(fsdata6, cor.test(Right.Amygdala, Left.Amygdala))
cor.test(fsdata6$Amyg_ave, fsdata6$Hipp_ave)
```

## CONFOUND VARIABLES
### make correlation matrix of amyg vol, ICV, GBV, motion, and aseg 
```{r, warnings = F}

my_data <- data.frame(with(fsdata6, cbind(brain_age_yrs, motion_ave_new, aseg_ave_rating, 
                                 ICV, GMV, Amyg_ave,Hipp_ave)))
names(my_data) <- c("brain_age_yrs",  "Motion", "Segmentation","ICV", "GMV", "Amygdala", "Hippocampus")
cor.table <- cor(my_data)
good_table <- round(cor.table, 2)

save(good_table, file = "tables/confound_variable_correlation_matrix.Rdata")
```

## get significance for each... 
```{r}
# make an empty table.
compiled_correlations <- data.frame(x = NA, y = NA, pvalue = NA)

# for each column in mydata...
for (i in 1:ncol(my_data)){
  print (i)
  var1 <- my_data[i]
  var1_name <- names(var1)
     var1_test <- as.matrix(var1)
  # get every other column in the dataframe...
  for (j in 1:ncol(my_data)){
    print (j)
     var2 <-my_data[j]
     var2_name <- names(var2)
     var2_test <- as.matrix(var2)
     # run the correlation and save the p-value...
     p_value <- cor.test(var1_test, var2_test)[3]
     # keep the p-value 
    row <- cbind(x=var1_name, y=var2_name, pvalue = p_value)
    compiled_correlations <- rbind(compiled_correlations, row)

  }
}
compiled_correlations$sig_level <- ifelse(compiled_correlations$pvalue < 0.001, '***', 
                                          ifelse(compiled_correlations$pvalue < 0.01, '**',
                                                 ifelse(compiled_correlations$pvalue < 0.05, '*', " ")))

compiled_correlations$x <- unlist(compiled_correlations$x)
compiled_correlations$y <- unlist(compiled_correlations$y)
compiled_correlations$pvalue <- unlist(compiled_correlations$pvalue)
head(compiled_correlations)
save(compiled_correlations, file = "tables/confound_variable_correlations_with_pvalues.Rdata")
```


### t-test table for categorical vars
```{r}

# sex, group, scanner 
fsdata6$scanner_confound <- ifelse(fsdata6$index_wave == 3, 1, 0)

my_dataX <- data.frame(with(fsdata6, cbind(sex, GROUP, scanner_confound)))
my_dataY <- data.frame(with(fsdata6, cbind(motion_ave_new, aseg_ave_rating, 
                                 ICV, GMV, Amyg_ave,Hipp_ave)))

# make an empty table.
compiled_ttests <- data.frame(x = NA, y = NA, tvalue = NA, df = NA, pvalue = NA)

# for each column in mydata...
for (i in 1:ncol(my_dataX)){
  print (i)
  var1 <- my_dataX[i]
  var1_name <- names(var1)
     var1_test <- as.matrix(var1)
  # get every other column in the dataframe...
  for (j in 1:ncol(my_dataY)){
    print (j)
     var2 <-my_dataY[j]
     var2_name <- names(var2)
     var2_test <- as.matrix(var2)
     # run the correlation and save the p-value...
     t_value <- t.test(var2_test ~ var1_test)[1]
     d_f <- t.test(var2_test ~ var1_test)[2]
     p_value <- t.test(var2_test ~ var1_test)[3]
     # keep the p-value 
    row <- cbind(x=var1_name, y=var2_name, tvalue = t_value, df = d_f, pvalue = p_value)
    compiled_ttests <- rbind(compiled_ttests, row)

  }
}
compiled_ttests$sig_level <- ifelse(compiled_ttests$pvalue < 0.001, '***', 
                                          ifelse(compiled_ttests$pvalue < 0.01, '**',
                                                 ifelse(compiled_ttests$pvalue < 0.05, '*', " ")))
head(compiled_ttests)
nrow(compiled_ttests)
compiled_ttests$x <- unlist(compiled_ttests$x)
compiled_ttests$y <- unlist(compiled_ttests$y)
compiled_ttests$tvalue <- unlist(compiled_ttests$tvalue)
compiled_ttests$df <- unlist(compiled_ttests$df)
compiled_ttests$pvalue <- unlist(compiled_ttests$pvalue)

save(compiled_ttests, file = "tables/confound_variable_t-tests_with_pvalues.Rdata")
```

# plots 
## whole brain measures
### plot ICV by group and brain_age_yrs 
PIs show faster decrease (aceelerated dev?) in ICV and GMV. 
ICV peaks later than GMV.
```{r,warnings=FALSE}

 ggplot(fsdata6, aes(x = brain_age_yrs, y = ICV, color = GROUP)) +
  geom_point( ) + theme_classic() + facet_grid(~ gender) +
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm') + 
  stat_smooth() 

```


### plot GMV by group and brain_age_yrs 
```{r,warnings=FALSE}

 ggplot(fsdata6, aes(x = brain_age_yrs, y = GMV, color = GROUP)) +
  geom_point( ) + theme_classic() + facet_grid(~gender)+
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm') + 
  stat_smooth() 

```


## Motion 
### plot motion with brain_age_yrs
PIs have less motion!! and then overall decline with brain_age_yrs. 
```{r, warnings = F}
 ggplot(fsdata6, aes(x =brain_age_yrs, y = motion_ave_new, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm') + 
  stat_smooth() 

```

### plot motion with GMV 
no association here! 
```{r, warnings = F}

RA <- ggplot(fsdata6, aes(x =motion_ave_new, y = GMV, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm') + 
  stat_smooth() 
RA

```

### plot motion with ICV 
```{r, warnings = F}

RA <- ggplot(fsdata6, aes(x =motion_ave_new, y = ICV, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm') + 
  stat_smooth() 
RA

```

### plot motion with amyg vol
no relationship 
```{r, warnings = F}
 ggplot(fsdata6, aes(x =motion_ave_new, y = Left.Amygdala, 
                             color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm')+ 
  stat_smooth() 

```

### plot motion with R amyg vol
no relationship 
```{r, warnings = F}
 ggplot(fsdata6, aes(x =motion_ave_new, y = Right.Amygdala, 
                             color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm')+ 
  stat_smooth() 

```

### plot motion with ASEG ratings 
not necessarily correlated. 
```{r, warnings = F}
 ggplot(fsdata6, aes(x =motion_ave_new, y = aseg_ave_rating, 
                             color = GROUP)) +
  geom_jitter( alpha = 0.5) + theme_classic() +
   stat_smooth(method = 'lm')
 # geom_line(aes(group = IDENT_SUBID), alpha = 0.5) 

```

## Segmentation quality
### plot aseg ratings with brain_age_yrs
no relationship. 
```{r, warnings = F}
 ggplot(fsdata6, aes(x = brain_age_yrs, y = aseg_ave_rating, 
                             color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm')

```

### plot aseg ratings with amyg. 
no relationship. 
```{r, warnings = F}
 ggplot(fsdata6, aes(x = aseg_ave_rating, y = Left.Amygdala, 
                             color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm')

```

### plot aseg ratings with R amyg. 
no relationship. 
```{r, warnings = F}
 ggplot(fsdata6, aes(x = aseg_ave_rating, y =Right.Amygdala, 
                             color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm')

```


# test covariates for impt confounds 
```{r, warnings = F}

hist(fsdata6$brain_age_yrs)
hist(fsdata6$ICV)
hist(fsdata6$GMV)
```

## make scanner confound variable
- nstead of index wave 
```{r, warnings = F}

fsdata6$scanner_confound <- with(fsdata6, as.factor(ifelse(index_wave == 1 | index_wave == 2, "scanner1", "scanner2")))
summary(fsdata6$scanner_confound)
```

## are there volume differences by scanner?
### ICV - yes
GMV is *smaller* for scanner 2
```{r}

 ggplot(fsdata6, aes(x =scanner_confound, y = ICV, 
                             color = GROUP)) +
  geom_boxplot( ) + theme_classic() 

```

### GMV 
GMV is *smaller* for scanner 2
```{r}

 ggplot(fsdata6, aes(x =scanner_confound, y = GMV, 
                             color = GROUP)) +
  geom_boxplot( ) + theme_classic() 

```

### Amyg  by scanner
subcortical volume is *larger* even though ICV/GMV is *smaller* 
```{r}
 ggplot(fsdata6, aes(x =scanner_confound, y =Right.Amygdala, 
                             color = GROUP)) +
  geom_boxplot( ) + theme_classic() 

```

### Hipp by scanner
subcortical volume is *larger* on scanner 2, even though ICV/GMV is *smaller* 
```{r}
 ggplot(fsdata6, aes(x =scanner_confound, y =Right.Hippocampus, 
                             color = GROUP)) +
  geom_boxplot( ) + theme_classic() 

```

### do we need to worry about data quality influencing volume measures?
no, but ICV & GMV influence amyg. 
```{r, warnings = F}
pairs(my_data[,1:4])
pairs(my_data[, 3:6])
```

### how strongly correlated are WBV and Amyg volume? 
quite highly correlated!! 
```{r, warnings = F}
with(fsdata6, cor.test(GMV, Right.Amygdala))
with(fsdata6, cor.test(GMV, Left.Amygdala))
```

### how strongly correlated are Hipp volume and Amyg volume? 
quite highly correlated!! 
```{r, warnings = F}
with(fsdata6, cor.test(Amyg_ave, Hipp_ave))
```

## group differences in confound regressors 
### GROUP and GMV 
main effect of GROUP: PIs have smaller 
 no effect of brain_age_yrs x group 
```{r, warnings = F}
summary(fsdata6$sex)
fsdata6$index_wave <- as.factor(fsdata6$index_wave)
library(nlme)

GMV_test <- lme(GMV ~ brain_age_yrs* GROUP + gender + 
                  index_wave, 
                random = ~ 1 | IDENT_SUBID, data = fsdata6)

anova(GMV_test)

```

huge effect of brain_age_yrs, group, sex and wave.
```{r}
GMV_test <- lme(GMV ~ brain_age_yrs + GROUP+ gender + 
                  index_wave, 
                random = ~ 1 | IDENT_SUBID, data = fsdata6)

anova(GMV_test)
```

### GROUP and ICV 
no  brain_age_yrs x group 
```{r, warnings = F}

ICV_test <- lme(ICV ~ brain_age_yrs*GROUP + gender+ 
                  index_wave, random = ~ 1 | IDENT_SUBID, data = fsdata6)

anova(ICV_test)
```

main effect of brain_age_yrs, group, gender, wave
```{r, warnings = F}

ICV_test <- lme(ICV ~ brain_age_yrs + GROUP + gender+ 
                  index_wave, random = ~ 1 | IDENT_SUBID, data = fsdata6)

anova(ICV_test)
```

### GROUP and motion
definitely brain_age_yrs effects, and wave effects.
*PI have LESS MOTION than COMPARISONS*. 
```{r, warnings = F}

motion_test <- lme(motion_ave_new ~ brain_age_yrs + GROUP + gender+ 
                  index_wave, random = ~ 1 | IDENT_SUBID, data = fsdata6)

anova(motion_test)
plot(effect("GROUP", motion_test))
```

### GROUP and aseg quality. 
nope!
```{r, warnings = F}

aseg_test <- lme(aseg_ave_rating ~ brain_age_yrs + GROUP + gender+ 
                  index_wave, random = ~ 1 | IDENT_SUBID, data = fsdata6)

anova(aseg_test)
plot(effect("GROUP", aseg_test))
```


