---
title: "Supplement: Growth chart for Hippocampus volume ages 6-19"
author: "Michelle.VanTieghem"
date: "October 16, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---


# Readme:
following same procedure as original analysis, with subset of data (6-19)


```{r, include = F, warnings =F}
source("../../0_R_analysis_setup_file.R")

```

# load data

```{r}

load("../../../data/2_fs_data/3_cleaned_fsdata_for_growth_chart_brain_age_yrs6_19.Rdata")
fsdata6$scanner_confound <-with(fsdata6, ifelse(index_wave == 3, 0, 1))
```

# Hippocampus VOLUMES
## plots of raw data for left hipp
```{r, warnings = F}
 ggplot(fsdata6, aes(x =brain_age_yrs, y = Left.Hippocampus, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) +  stat_smooth() 
```


## test differences between hemispheres? 
### if not, averbrain_age_yrs them and use for subsequent analyses 
```{r, warnings = F}
RA <- ggplot(fsdata6, aes(x = Right.Hippocampus, y = Left.Hippocampus, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm') 
RA
```


## Bilateral Hippocampus 
probably quadratic in comps. 
```{r, warnings = F}
ggplot(fsdata6, aes(x =brain_age_yrs, y = Hipp_ave, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth() 

```

# fix up variables
## rescale variables 
for coefficients to be interpretable
```{r}

fsdata6$Hipp_ave <- fsdata6$Hipp_ave/100
# NOTE: ICV NEEDS TO BE RESCALED FOR SIMILAR SD RANGE AND INTERPRETABLE COEFFICIENTS
fsdata6$ICV <- fsdata6$ICV/100000

```


## mean center stuff

```{r}
fsdata6$brain_age_yrs.c <- fsdata6$brain_age_yrs - mean(fsdata6$brain_age_yrs)
fsdata6$sex.c <- fsdata6$sex - mean(fsdata6$sex)
# main effect of age =squared 
fsdata6$age_squared <- fsdata6$brain_age_yrs^2
fsdata6$age_squared.c <- fsdata6$age_squared - mean(fsdata6$age_squared)
fsdata6$ICV.c <- fsdata6$ICV - mean(fsdata6$ICV)
fsdata6$meds.c <- fsdata6$ICV - mean(fsdata6$ICV)
fsdata6$motion_ave_new.c  <- fsdata6$motion_ave_new - mean(fsdata6$motion_ave_new)
fsdata6$scanner_confound.c <- fsdata6$scanner_confound - mean(fsdata6$scanner_confound)

```


# AGE MODELS FOR Hippocampus
do brain_age_yrs and group interact to predict change in Hippocampus? 
using lmer, mixed effects - random intercepts for subject only. 

## linear model
interaction point around 12
```{r, warnings = F}
# same for ICV. 
hipp_linear <- lmer(Hipp_ave ~ brain_age_yrs *GROUP + sex.c + ICV.c 
                     + motion_ave_new.c + 
                 scanner_confound.c +(1 | IDENT_SUBID), data = subset(fsdata6, brain_age_yrs < 20))


hipp_linear_model <- anova(hipp_linear)
summary(hipp_linear)
save(hipp_linear_model, file = "model_output/hipp_linear_model_results.Rdata")

AIC(hipp_linear)
```



### graph model with raw data 
```{r, warnings = F}

#  graph the model and confirm the post-hoc tests!! 
effect_hipp <- as.data.frame(effect("brain_age_yrs:GROUP", hipp_linear, confint=TRUE, xlevels = list(brain_age_yrs=c(min(fsdata6$brain_age_yrs), min(fsdata6$brain_age_yrs[fsdata6$GROUP == "PI"]), mean(fsdata6$brain_age_yrs),
  max(fsdata6$brain_age_yrs[fsdata6$GROUP == "PI"]),
  max(fsdata6$brain_age_yrs[fsdata6$GROUP == "COMP"])))))
# remove oldest and youngest PI age ranges. 
effect_hipp <- effect_hipp %>%
  filter(! (GROUP == "PI" & brain_age_yrs == min(brain_age_yrs[GROUP == "COMP"]))
         & !(GROUP == "PI" & brain_age_yrs == max(brain_age_yrs[GROUP == "COMP"])))

hipp_linear_plot_with_raw <- ggplot(effect_hipp, 
                                    aes(x = brain_age_yrs, y = fit, color = GROUP)) +
  geom_ribbon(aes(ymin=lower, ymax=upper, 
                  group = GROUP, fill = GROUP), alpha =0.2, color = NA) +
# fill = "grey70", color = "grey70") + 
   geom_line(size = 1) + facet_grid (~GROUP) + 
   theme_classic() + ylab(bquote('Hippocampus Volume '*(mm/100)^3*'')) + 
  xlab ("Age (years)") + my_colors + my_colors2 + 
   # add raw data points! 
   geom_point(data = fsdata6, aes(x = brain_age_yrs, y = Hipp_ave, color = GROUP),
              size = 1, alpha = 0.5) + 
   geom_line(data = fsdata6, aes(x = brain_age_yrs, y = Hipp_ave, group = IDENT_SUBID), alpha = 0.5) + scale_x_continuous(breaks = seq(4,21.5, by = 3)) + theme(legend.position="none")

#pdf (file = "figures/Hippocampus_linear_model_with_raw.#pdf", width = 6 , height = 4)
hipp_linear_plot_with_raw
#dev.off()

```


## main effects linear model 
```{r, warnings = F}
# same for ICV.
hipp_linear_main_effects <- lmer(Hipp_ave ~ brain_age_yrs + GROUP + sex.c + ICV.c 
                     + motion_ave_new.c + 
                 scanner_confound.c +(1 | IDENT_SUBID), data = subset(fsdata6, brain_age_yrs < 20))

summary(hipp_linear_main_effects)


save(hipp_linear_main_effects, file = "model_output/hipp_linear_main_effects_model_results.Rdata")

AIC(hipp_linear_main_effects)
```
