---
title: "merging freesurfer output, QC data, and master datasets"
author: "Michelle VanTieghem"
date: "January 23, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# Readme:
1) filter freesurfer data for only dev sample \
2) add demographics \
3) clean QC excel sheet \ 
4) combine with freesurfer data \


checked Jan 30, 2019

```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
```

# load data 
## raw freesurfer data
```{r}
fsdata <- read.csv("../../../data/2_fs_data/raw_data/v6_asegstats_all_compiled_Marta_02-01-2018.csv")
nrow(fsdata)
names(fsdata)[1] <- "SUBJECTID"
# mark all of the waves for each row of data
fsdata$WAVE <- ifelse(grepl("fu1", fsdata$SUBJECTID), "WAVE2", ifelse(grepl("fu2", fsdata$SUBJECTID), "WAVE3", "WAVE1"))

# save this "long formatted" SUBJECTID 
fsdata$SUBJECTID_long <- fsdata$SUBJECTID 

# convert into a long datasest by removing "fu1" "fu2" on SUBIDS
# ( only keep first 5 characters of subjectid)
fsdata$SUBJECTID <- substr(fsdata$SUBJECTID_long, 1, 5)

```

## get master data with demographics 
```{r, message = F}
master <- read_csv("../../../data/master/J.Data_Entry_Master_8_17_2017_fixed_fixed.csv")

# select demo variables we need 
master$SUBJECTID <- as.character(master$IDENT_SUBID)
sub <- master %>% 
  dplyr::select("SUBJECTID","IDENT_SUBTYPE", "SUBAGE_Session1", "SUBAGE_session2", "SUBAGE_session3", "SUBAGE_session4", "DEM_3_GENDER_CHILD") %>%
  rename(GROUP = IDENT_SUBTYPE)

## convert into a long dataset by wave, repeating session AGE
sub_long <- sub %>%
  gather(Session, brain_age_mos, SUBAGE_session2:SUBAGE_session4) %>%
  # make a long-version of this ID number. we will use this to merge with fsdata.
  # note: session 2 = baseline scan 
  # session 3 = scan at fu1 
  # session 4 = scan at fu2 
  mutate(SUBJECTID_long = ifelse(Session == "SUBAGE_session3", paste0(SUBJECTID, "_fu1"), ifelse(Session == "SUBAGE_session4", paste0(SUBJECTID, "_fu2"),SUBJECTID)))

```

## now, merge two long datasets by longID
```{r}
# use the long version of SUBJECTID, so that each wave is labeled as a row.
df <- merge(fsdata, sub_long, by= "SUBJECTID_long")
N_total_scans <- nrow(df) 

# exclude  adults and preschoolers
df_dev <- filter(df, GROUP != 3 & GROUP !=2)
```

## manually fix age for known error
```{r, results = F}
# one sscan has no AGE: SB019_fu2
sum(is.na(df_dev$brain_age_mos))

# fix SB019 ages manually.
# ages labeled incorrectly in master:
subSB019 <- sub %>%
  filter(SUBJECTID == "SB019")
subSB019 # these labels are supposed to be  SUBAGE session2  = first scan, SUBAGE session3 = fu1, SUBAGE session4 = fu2


# for this subject, SUBAGE session 2 = fu1, SUBAGE session 3 = fu2 
fsdata %>%
  filter(SUBJECTID == "SB019") %>%
  dplyr::select(SUBJECTID_long, SUBJECTID)

# change the ages to correct times.
df_dev$brain_age_mos[df_dev$SUBJECTID_long == "SB019_fu1"] <- subSB019$SUBAGE_session2
df_dev$brain_age_mos[df_dev$SUBJECTID_long == "SB019_fu2"] <- subSB019$SUBAGE_session3

# make age in years instead of months!
df_dev$brain_age_yrs <- df_dev$brain_age_mos/12

```

## Now add ASEG Quality Control info
excel sheet with motion and segmentation ratings 
```{r}

df.qc <- read_excel("../../../data/2_fs_data/raw_data/Freesurfer_Aseg_SB_organized_final_March30_2018.xlsx")
nrow(df.qc)

```

## clean dataset 
```{r}
# removing extra row in header
names(df.qc) <- df.qc[1,]
df.qc <- df.qc[2:nrow(df.qc), ]


## rename variables 
# which motion was calculated by michelle, jasmin marta or liliana.
names(df.qc)[1] <- "SUBJECTID_long"
names(df.qc)[10] <- "motion_Michelle"
names(df.qc)[11] <- "motion_Jasmin"
names(df.qc)[12] <- "motion_Marta"
names(df.qc)[13] <- "motion_average"
names(df.qc)[14] <- "inclusion"
names(df.qc)[15] <- "reason_exclude"
names(df.qc)[16] <- "ASEG_rating_Marta"
names(df.qc)[17] <- "GM_rating_Marta"
names(df.qc)[21] <- "ASEG_rating_Liliana"
names(df.qc)[22] <- "GM_rating_Liliana"

# remove junk columns
df.qc <- df.qc[,-c(2,3,4,5,6,7,8,9)]
```

## merge QC data and freesurfer volume data - wide format 
```{r}
nrow(df.qc) # 389 rows because several adult scans 
nrow(df_dev) # 330 for dev sample

# merrge them together
fsdata <- merge(df.qc, df_dev, by = c("SUBJECTID_long"))
nrow(fsdata) # 330 for dev sample
ncol(fsdata) # does not include master dataset

# check that subjectids are correctly formatted. 
length(unique(fsdata$SUBJECTID_long)) # number of scans
length(unique(fsdata$SUBJECTID.x)) # number of subjects

# make a subid that doesn't have fu1 / fu2 added
fsdata$IDENT_SUBID <- str_sub(fsdata$SUBJECTID_long, 1, 5)

```

## update wave variable (numbers 1,2,3) instead of fu1, fu2
```{r}
fsdata$index_wave <- as.factor(ifelse(grepl("fu1", fsdata$SUBJECTID_long), 2, 
                            ifelse(grepl("fu2", fsdata$SUBJECTID_long), 3, 1)))
summary(fsdata$index_wave)

sum(is.na(fsdata$brain_age_yrs))
```

# final data frame
## check age & gender distribution by wave and group
```{r}
Age_by_wave_table <- fsdata %>% 
  group_by(GROUP, WAVE) %>%
  dplyr::summarize (Nscans=n(), 
             min_age_by_wave= min(brain_age_yrs, na.rm = TRUE),
              max_age= max(brain_age_yrs,  na.rm = TRUE),
             mean_age=mean(brain_age_yrs,  na.rm = TRUE),  
             sd_age=sd(brain_age_yrs,  na.rm = TRUE), 
             Ratio_female=mean(DEM_3_GENDER_CHILD))
Age_by_wave_table

```
 

## save this data~
```{r, include = F}

save(fsdata, file = "../../../data/2_fs_data/1_all_freesurfer_v6_data_with_aseg_QC_info.RData")

```


