---
title: 'Checking confound variables: motion, aseg, scanner confound, ICV/GMV'
author: "Michelle VanTieghem"
date: "January 23, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# Readme:
1) checking correlations  of confound variables  \
2) checking categorical differences for sex, group, scanner \
3) plotting effects of confound variables \


```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
```

# load data 
```{r}
# removed outliers already
load("../../../data/2_fs_data/3_cleaned_fsdata_for_growth_chart_brain_age_yrs4_19.Rdata")

```



# check correlations among variables 
## hemisphere correlations 
```{r, warnings = F}
with(fsdata6, cor.test(Right.Hippocampus, Left.Hippocampus))
with(fsdata6, cor.test(Right.Amygdala, Left.Amygdala))
cor.test(fsdata6$Amyg_ave, fsdata6$Hipp_ave)
```

## continuous confound variables
make correlation matrix of amyg vol, ICV, GBV, motion, and aseg 
```{r, warnings = F}
# select variables 
my_data <- data.frame(with(fsdata6, cbind(brain_age_yrs, motion_ave_new, aseg_ave_rating, 
                                 ICV, GMV, Amyg_ave,Hipp_ave)))
names(my_data) <- c("brain_age_yrs",  "Motion", "Segmentation","ICV", "GMV", "Amygdala", "Hippocampus")
# get the correlations

cor.table <- cor(my_data)
good_table <- round(cor.table, 2)
good_table 
save(good_table, file = "tables/confound_variable_correlation_matrix.Rdata")
```

```{r}
# make an empty table.
compiled_correlations <- data.frame(x = NA, y = NA, pvalue = NA)

# for each column in mydata...
for (i in 1:ncol(my_data)){
  print (i)
  var1 <- my_data[i]
  var1_name <- names(var1)
     var1_test <- as.matrix(var1)
  # get every other column in the dataframe...
  for (j in 1:ncol(my_data)){
    print (j)
     var2 <-my_data[j]
     var2_name <- names(var2)
     var2_test <- as.matrix(var2)
     # run the correlation and save the p-value...
     p_value <- cor.test(var1_test, var2_test)[3]
     # keep the p-value 
    row <- cbind(x=var1_name, y=var2_name, pvalue = p_value)
    compiled_correlations <- rbind(compiled_correlations, row)

  }
}
compiled_correlations$sig_level <- ifelse(compiled_correlations$pvalue < 0.001, '***', 
                                          ifelse(compiled_correlations$pvalue < 0.01, '**',
                                                 ifelse(compiled_correlations$pvalue < 0.05, '*', " ")))

compiled_correlations$x <- unlist(compiled_correlations$x)
compiled_correlations$y <- unlist(compiled_correlations$y)
compiled_correlations$pvalue <- unlist(compiled_correlations$pvalue)
head(compiled_correlations)
save(compiled_correlations, file = "tables/confound_variable_correlations_with_pvalues.Rdata")
```


## t-test table for categorical vars
testing for group, sex, scanner effects on Age, motion, segmentation quality.
```{r}

fsdata_by_sub <-fsdata6 %>%
  group_by(IDENT_SUBID ) %>%
  dplyr::summarize(GROUP = GROUP[1], 
                   sex = sex[1],
                   ICV = mean(ICV, na.rm = T),
                   brain_age_yrs = mean(brain_age_yrs, na.rm = T),
                   motion_ave_new = mean(motion_ave_new, na.rm = T), 
                   aseg_ave_rating = mean(aseg_ave_rating, na.rm = T))

# set up x andy variables 
my_dataX <- data.frame(with(fsdata_by_sub, cbind(GROUP)))
my_dataY <- data.frame(with(fsdata_by_sub, cbind(sex, brain_age_yrs, motion_ave_new, aseg_ave_rating, 
                                 ICV)))

# make an empty table.
compiled_ttests <- data.frame(x = NA, y = NA, tvalue = NA, df = NA, pvalue = NA)

# set up loop for making t-tests 
# for each column in mydata...
for (i in 1:ncol(my_dataX)){
  print (i)
  var1 <- my_dataX[i]
  var1_name <- names(var1)
     var1_test <- as.matrix(var1)
  # get every other column in the dataframe...
  for (j in 1:ncol(my_dataY)){
   # print (j)
     var2 <-my_dataY[j]
     var2_name <- names(var2)
     var2_test <- as.matrix(var2)
     # run the correlation and save the p-value...
     t_value <- t.test(var2_test ~ var1_test)[1]
     d_f <- t.test(var2_test ~ var1_test)[2]
     p_value <- t.test(var2_test ~ var1_test)[3]
     # keep the p-value 
    row <- cbind(x=var1_name, y=var2_name, tvalue = t_value, df = d_f, pvalue = p_value)
    compiled_ttests <- rbind(compiled_ttests, row)

  }
}
compiled_ttests$sig_level <- ifelse(compiled_ttests$pvalue < 0.001, '***', 
                                          ifelse(compiled_ttests$pvalue < 0.01, '**',
                                                 ifelse(compiled_ttests$pvalue < 0.05, '*', " ")))
compiled_ttests$x <- unlist(compiled_ttests$x)
compiled_ttests$y <- unlist(compiled_ttests$y)
compiled_ttests$tvalue <- unlist(compiled_ttests$tvalue)
compiled_ttests$df <- unlist(compiled_ttests$df)
compiled_ttests$pvalue <- unlist(compiled_ttests$pvalue)
compiled_ttests_brain <- compiled_ttests
compiled_ttests_brain
save(compiled_ttests_brain, file = "tables/confound_variable_t-tests_with_pvalues.Rdata")
```


# Testing group differences on whole brain measures
while including effects of age, sex, scanner confound. 
## main effect of group on ICV: PI < COMPS
```{r}

ICV_main <- lmer(ICV ~ brain_age_yrs + GROUP + sex +  scanner_confound +  motion_ave_new +
                (1 | IDENT_SUBID), 
              data = fsdata6)
summary(ICV_main)
```

## no group x age interaction on ICV
```{r}

ICV_int <- lmer(ICV ~ brain_age_yrs * GROUP + sex +  scanner_confound + motion_ave_new +
                (1 | IDENT_SUBID), 
              data = fsdata6)
summary(ICV_int)
```

plotting grorup effect
```{r}
# plotting group and age effects even though no interaction to show that there is no interaction! 
effect_df <- data.frame(effect("brain_age_yrs:GROUP", ICV_main))
 
ICV_plot <- ggplot(data = effect_df, aes(x = brain_age_yrs, y = fit)) +
  geom_line(aes(color = GROUP) ) + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = GROUP), alpha = 0.2) +
  geom_point(data = fsdata6,aes(x = brain_age_yrs, y = ICV, color = GROUP), alpha = 0.2 ) + 
  geom_line(data = fsdata6, aes(x = brain_age_yrs, y = ICV, 
                                group = IDENT_SUBID, color = GROUP), alpha = 0.2) +
  theme_classic() + ylab (" ICV") + xlab ("Age (years)") +
  my_colors + my_colors2 
ICV_plot

ggsave(ICV_plot, file = "figures/ICV_by_age_and_group.pdf")
```

plotting scanner effect
```{r}
# plotting group and age effects even though no interaction to show that there is no interaction! 
effect_df <- data.frame(effect("scanner_confound", ICV_main)) 
 
motion_plot <- ggplot(data = effect_df, aes(x = scanner_confound, y = fit)) +
  geom_point(color = "red", size = 2) + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_jitter(data = fsdata6,
             aes(x = scanner_confound, y = ICV), alpha = 0.2, width = 0.1) + 
  theme_classic() + ylab ("ICV") + xlab ("Scanner") +
  my_colors + my_colors2 
motion_plot

ggsave(motion_plot, file = "figures/ICV_by_scanner.pdf")
```


## main effect of group on GMV: PI < COMPS
```{r}

GMV_main <- lmer(GMV ~ brain_age_yrs + GROUP + sex + scanner_confound + motion_ave_new +
                (1 | IDENT_SUBID), 
              data = fsdata6)
summary(GMV_main)
```

## no group x age interaction on GMV
```{r}

GMV_int <- lmer(GMV ~ brain_age_yrs * GROUP + sex + scanner_confound + motion_ave_new +
                (1 | IDENT_SUBID), 
              data = fsdata6)
summary(GMV_int)
```

plotting effect
```{r}
# plotting group and age effects even though no interaction to show that there is no interaction! 
effect_df <- data.frame(effect("brain_age_yrs:GROUP", GMV_main))
 
GMV_plot <- ggplot(data = effect_df, aes(x = brain_age_yrs, y = fit)) +
  geom_line(aes(color = GROUP) ) + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = GROUP), alpha = 0.2) +
  geom_point(data = fsdata6,aes(x = brain_age_yrs, y = GMV, color = GROUP), alpha = 0.2 ) + 
  geom_line(data = fsdata6, aes(x = brain_age_yrs, y = GMV, 
                                group = IDENT_SUBID, color = GROUP), alpha = 0.2) +
  theme_classic() + ylab (" GMV") + xlab ("Age (years)") +
  my_colors + my_colors2 
GMV_plot

ggsave(GMV_plot, file = "figures/GMV_by_age_and_group.pdf")
```


plotting scanner effect
```{r}
# plotting group and age effects even though no interaction to show that there is no interaction! 
effect_df <- data.frame(effect("scanner_confound", GMV_main)) 
 
motion_plot <- ggplot(data = effect_df, aes(x = scanner_confound, y = fit)) +
  geom_point(color = "red", size = 2) + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_jitter(data = fsdata6,
             aes(x = scanner_confound, y = GMV), alpha = 0.2, width = 0.1) + 
  theme_classic() + ylab ("GMV") + xlab ("Scanner") +
  my_colors + my_colors2 
motion_plot

ggsave(motion_plot, file = "figures/GMV_by_scanner.pdf")
```


# Motion by group 
when controlling for age, sex, scanner, aseg ratings 
## main effect is a trend: PI < COMPS 
```{r}

motion_main <- lmer(motion_ave_new ~ brain_age_yrs + GROUP + sex + scanner_confound + 
                   aseg_ave_rating + 
                (1 | IDENT_SUBID), 
              data = fsdata6)
summary(motion_main)

```

## no group x age interaction on motion
```{r}

motion_int <- lmer(motion_ave_new ~ brain_age_yrs * GROUP + sex + scanner_confound + aseg_ave_rating +
                (1 | IDENT_SUBID), 
              data = fsdata6)
summary(motion_int)
```

plotting effect
```{r}
# plotting group and age effects even though no interaction to show that there is no interaction! 
effect_df <- data.frame(effect("brain_age_yrs:GROUP", motion_main))
 
motion_plot <- ggplot(data = effect_df, aes(x = brain_age_yrs, y = fit)) +
  geom_line(aes(color = GROUP) ) + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = GROUP), alpha = 0.2) +
  geom_point(data = fsdata6,
             aes(x = brain_age_yrs, y = motion_ave_new, color = GROUP), alpha = 0.2 ) + 
  geom_line(data = fsdata6, aes(x = brain_age_yrs, y = motion_ave_new, 
                                group = IDENT_SUBID, color = GROUP), alpha = 0.2) +
  theme_classic() + ylab ("Average motion rating") + xlab ("Age (years)") +
  my_colors + my_colors2 
motion_plot

ggsave(motion_plot, file = "figures/motion_by_age_and_group.pdf")
```


plotting scanner effect
```{r}
# plotting group and age effects even though no interaction to show that there is no interaction! 
effect_df <- data.frame(effect("scanner_confound", motion_main)) 
 
motion_plot <- ggplot(data = effect_df, aes(x = scanner_confound, y = fit)) +
  geom_point(color = "red", size = 2) + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_jitter(data = fsdata6,
             aes(x = scanner_confound, y = motion_ave_new), alpha = 0.2, width = 0.1) + 
  theme_classic() + ylab ("Average motion rating") + xlab ("Scanner") +
  my_colors + my_colors2 
motion_plot

ggsave(motion_plot, file = "figures/motion_by_scanner.pdf")
```

## Segmentation quality by group 
no relationship. 

when controlling for age, sex, scanner, motion ratings 
## main effect is NS
```{r}

aseg_main <- lmer(aseg_ave_rating ~ brain_age_yrs + GROUP + sex + scanner_confound + 
                   motion_ave_new + 
                (1 | IDENT_SUBID), 
              data = fsdata6)
summary(aseg_main)

```

plotting effect
```{r}
# plotting group and age effects even though no interaction to show that there is no interaction! 
effect_df <- data.frame(effect("brain_age_yrs:GROUP", aseg_main))
 
aseg_plot <- ggplot(data = effect_df, aes(x = brain_age_yrs, y = fit)) +
  geom_line(aes(color = GROUP) ) + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = GROUP), alpha = 0.2) +
  geom_point(data = fsdata6,
             aes(x = brain_age_yrs, y = aseg_ave_rating, color = GROUP), alpha = 0.2 ) + 
  geom_line(data = fsdata6, aes(x = brain_age_yrs, y = aseg_ave_rating, 
                                group = IDENT_SUBID, color = GROUP), alpha = 0.2) +
  theme_classic() + ylab ("Average Segmenntation Rating") + xlab ("Age (years)") +
  my_colors + my_colors2 
aseg_plot

ggsave(aseg_plot, file = "figures/aseg_by_age_and_group.pdf")
```




plotting scanner effect
```{r}
# plotting group and age effects even though no interaction to show that there is no interaction! 
effect_df <- data.frame(effect("scanner_confound", aseg_main)) 
 
motion_plot <- ggplot(data = effect_df, aes(x = scanner_confound, y = fit)) +
  #geom_point(color = "red", size = 2) + 
  #geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_boxplot(data = fsdata6,
             aes(x = scanner_confound, y = aseg_ave_rating)) +
  geom_jitter(data = fsdata6,
             aes(x = scanner_confound, y = aseg_ave_rating), alpha = 0.2, width = 0.1) + 
  theme_classic() + ylab ("Average Segmentation Rating") + xlab ("Scanner") +
  my_colors + my_colors2 
motion_plot

ggsave(motion_plot, file = "figures/aseg_rating_by_scanner.pdf")
```
