---
title: "Growth chart for amygdala volume ages 4+"
author: "Michelle.VanTieghem"
date: "October 16, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# Readme:
Analysis of amygdala volume, comparing linear quadratic and piecewise models \


```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
keep_legend <- theme(legend.position= c(0.90, 0.10),  legend.key.height = unit(0.05, "cm"), legend.key.width = unit(0.05, "cm"), legend.background = element_rect(color = "black"))
```


```{r, include = F, echo = F}
# functions for post-hoc / simple slopes / conditional effects from mixed effects models. 

## simple slope function 
condslope.lmer.slopes <- function(x, z, c, y){
#  condslope.lmer( "age1","GROUP", 0, Amyg_piecewise_orig)
  # x = name of x variable
  # z = name of z variable (moderator)
  # c = conditional value of z
  # y = reg object
  # lmer model must be in this order x + z + x*z
# updated slightly by MVT on March 4, 2019 
# bcause tidy doesn't work. 
out <- summary(y)
xz <- paste(x, z, sep=":")
w0.intercept <- fixef(y)["(Intercept)"] + fixef(y)[z]*c
w1.slope <- fixef(y)[x] + fixef(y)[xz]*c 
#y.cond <- w0.intercept + w1.slope*xvalue
require(broom)
modtidy <- data.frame(summary(y)$coef)
modtidy$rownames <- rownames(modtidy)
coef2.var <- subset(modtidy, rownames == x)$Std..Error^2
coef4.var <- subset(modtidy, rownames == xz)$Std..Error^2
out.vcov <- vcov(y)
cond.se <- sqrt(coef2.var + (c) * (c) * coef4.var + 2 * (c) * out.vcov[x, xz])
t.val <- w1.slope/cond.se
p.val <- 2*(1-pt(abs(t.val), subset(modtidy, rownames == x)$df, lower.tail=T))
lower95 <- w1.slope-qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
upper95 <- w1.slope+qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
# z.out <- z.test(w1.slope, sigma.x= cond.se)
return(list(w0.intercept=round(w0.intercept, digits = 5), w1.slope=round(w1.slope, digits = 5),
            df = subset(modtidy, rownames == x)$df,
            t.val = round(t.val, digits = 5), p.val = round(p.val, digits = 3), 
            lower95 = round(lower95, digits = 5), 
            upper95 = round(upper95, digits = 5)))
}

# function for post-hoc group effects
condslope.lmer.pointestimates <- function(x, z, c, y){
#  condslope.lmer( "age1","GROUP", 0, Amyg_piecewise_orig)
  # x = name of x variable
  # z = name of z variable (moderator)
  # c = conditional value of z
  # y = reg object
  # lmer model must be in this order x + z + x*z
  
# updated slightly by MVT on March 4, 2019 
# bcause tidy doesn't work. 
out <- summary(y)
xz <- paste(x, z, sep=":")
w0.intercept <- fixef(y)["(Intercept)"] + fixef(y)[z]*c
w1.slope <- fixef(y)[x] + fixef(y)[xz]*c 
#y.cond <- w0.intercept + w1.slope*xvalue
require(broom)
modtidy <- data.frame(summary(y)$coef)
modtidy$rownames <- rownames(modtidy)
coef2.var <- subset(modtidy, rownames == x)$Std..Error^2
coef4.var <- subset(modtidy, rownames == xz)$Std..Error^2
out.vcov <- vcov(y)
cond.se <- sqrt(coef2.var + (c) * (c) * coef4.var + 2 * (c) * out.vcov[x, xz])
t.val <- w1.slope/cond.se
p.val <- 2*(1-pt(abs(t.val), subset(modtidy, rownames == x)$df, lower.tail=T))
lower95 <- w1.slope-qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
upper95 <- w1.slope+qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
# z.out <- z.test(w1.slope, sigma.x= cond.se)
return(list(w0.intercept=round(w0.intercept, digits = 5), w1.slope=round(w1.slope, digits = 5),
            df = subset(modtidy, rownames == x)$df,
            t.val = round(t.val, digits = 5), p.val = round(p.val, digits = 3), 
            lower95 = round(lower95, digits = 5), 
            upper95 = round(upper95, digits = 5)))
}


```


# load data
outliers already removed! 
```{r, include = F, warnings =F}
load("../../../data/2_fs_data/3_cleaned_fsdata_for_growth_chart_brain_age_yrs4_19.Rdata")
nrow(fsdata6)
fsdata6$scanner_confound <- with(fsdata6, ifelse(index_wave == 3, 0, 1))
```

# AMYGDALA VOLUMES
## plots of raw data for left Amyg
```{r, warnings = F}
 ggplot(fsdata6, aes(x =brain_age_yrs, y = Left.Amygdala, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) +  stat_smooth()  + 
  my_colors + my_colors2
```


## test differences between hemispheres? 
```{r, warnings = F}
RA <- ggplot(fsdata6, aes(x = Right.Amygdala, y = Left.Amygdala, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm')  + 
  my_colors + my_colors2 
RA

Left_right_amyg_cor <- cor.test(fsdata6$Right.Amygdala, fsdata6$Left.Amygdala)
Left_right_hipp_cor <- cor.test(fsdata6$Right.Hippocampus, fsdata6$Left.Hippocampus)
save(Left_right_amyg_cor, file = "model_output/cor_hemispheres_amyg.Rdata")
save(Left_right_hipp_cor, file = "model_output/cor_hemispheres_hipp.Rdata")

```


## Bilateral amygdala 
```{r, warnings = F}
ggplot(fsdata6, aes(x =brain_age_yrs, y = Amyg_ave, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth()  + 
  my_colors + my_colors2 

```

# data cleaninng
## rescale variables 
for coefficients to be interpretable
```{r}

fsdata6$Amyg_ave <- fsdata6$Amyg_ave/100
### NOTE: ICV SCALED EVEN MORE SO THAT COEFFICIENTS ARE INTERPRETABLE/ SD IN SIMILAR RANGE.
fsdata6$ICV <- fsdata6$ICV/100000

hist(fsdata6$Amyg_ave)

```


## mean center variables

```{r}
fsdata6$brain_age_yrs.c <- fsdata6$brain_age_yrs - mean(fsdata6$brain_age_yrs)
fsdata6$sex.c <- fsdata6$sex - mean(fsdata6$sex)
# main effect of age =squared 
fsdata6$age_squared <- fsdata6$brain_age_yrs^2
fsdata6$age_squared.c <- fsdata6$age_squared - mean(fsdata6$age_squared)
fsdata6$ICV.c <- fsdata6$ICV - mean(fsdata6$ICV)
fsdata6$meds.c <- fsdata6$ICV - mean(fsdata6$ICV)
fsdata6$motion_ave_new.c  <- fsdata6$motion_ave_new - mean(fsdata6$motion_ave_new)
fsdata6$scanner_confound.c <- fsdata6$scanner_confound - mean(fsdata6$scanner_confound)

```


# AGE MODELS FOR AMYGDALA
does age and group interact to predict change in amygdala? 
using lmer, mixed effects - random intercepts for subject only. 

## 0) main effect 

```{r, warnings = F}
# same for ICV. 
amyg_linear <- lmer(Amyg_ave ~ GROUP + brain_age_yrs.c + sex.c + ICV.c 
                     + motion_ave_new.c + 
                 scanner_confound.c +(1 | IDENT_SUBID), data = fsdata6)

summary(amyg_linear)

save(amyg_linear, file = "model_output/Amyg_main_effect_model_results.Rdata")
```

## 1) linear model of group x age interaction


```{r, warnings = F}
# same for ICV. 
amyg_linear <- lmer(Amyg_ave ~ GROUP *brain_age_yrs.c + sex.c + ICV.c 
                     + motion_ave_new.c + 
                 scanner_confound.c +(1 | IDENT_SUBID), data = fsdata6)

summary(amyg_linear)

save(amyg_linear, file = "model_output/Amyg_linear_model_results.Rdata")
```


### post-hoc at specific ages 
```{r}
amyg_sd_age <- sd(fsdata6$brain_age_yrs)
amyg_mean_Age <- mean(fsdata6$brain_age_yrs)
## GETTING GROUP DIFFERENCES, BASED ON AGE 
group_diff_2agelwr <- data.frame(level = "2SD lower", age = amyg_mean_Age  - 2*amyg_sd_age, condslope.lmer.pointestimates("GROUPPI", "brain_age_yrs", amyg_mean_Age -2*amyg_sd_age, amyg_linear))
group_diff_1.5agelwr <- data.frame(level = "1.5 SD lower", age = amyg_mean_Age  - 1.5*amyg_sd_age, condslope.lmer.pointestimates("GROUPPI", "brain_age_yrs", amyg_mean_Age -1.5*amyg_sd_age, amyg_linear))


group_diff_1agelwr <- data.frame(level = "1SD lower", age = amyg_mean_Age - amyg_sd_age, condslope.lmer.pointestimates("GROUPPI", "brain_age_yrs", amyg_mean_Age -1*amyg_sd_age, amyg_linear))
group_diff_mean_age <- data.frame(level = "mean", age = amyg_mean_Age , condslope.lmer.pointestimates("GROUPPI", "brain_age_yrs",  amyg_mean_Age, amyg_linear))
group_diff_1ageupr <- data.frame(level = "1SD upper", age = amyg_mean_Age +amyg_sd_age, condslope.lmer.pointestimates("GROUPPI", "brain_age_yrs", amyg_mean_Age+ 1*amyg_sd_age, amyg_linear))
group_diff_1.5ageupr <- data.frame(level = "1.5 SD upper", age = amyg_mean_Age + 1.5*amyg_sd_age, condslope.lmer.pointestimates("GROUPPI", "brain_age_yrs", amyg_mean_Age + 1.5*amyg_sd_age, amyg_linear))

group_diff_2ageupr <- data.frame(level = "2SD upper", age = amyg_mean_Age  + 2*amyg_sd_age, condslope.lmer.pointestimates("GROUPPI", "brain_age_yrs", amyg_mean_Age + 2*amyg_sd_age, amyg_linear))

posthoc_table <- rbind(group_diff_2agelwr, group_diff_1.5agelwr, group_diff_1agelwr, group_diff_mean_age,  group_diff_1ageupr, group_diff_1.5ageupr, group_diff_2ageupr)

posthoc_table <- posthoc_table %>%
  dplyr::select(-w0.intercept)
names(posthoc_table) <- c("Level", "Age (years)", "Estimate", "df", "t.value", "p.value", "lower.95%.CI", "upper.95.CI%")
posthoc_table
save(posthoc_table, file = "model_output/amyg_linear_posthoc_group_diff.Rdata")
```


### graph model with raw data 
```{r, warnings = F}

age_list <- seq(4,19, by = 0.5)

#  graph the model and confirm the post-hoc tests!! 
effect_amyg <- as.data.frame(effect("GROUP:brain_age_yrs", amyg_linear, confint=TRUE, xlevels = list(brain_age_yrs= age_list)))

amyg_linear_plot_with_raw <- ggplot(effect_amyg, 
                                    aes(x = brain_age_yrs, y = fit, color = GROUP)) +
    geom_line(size = 1) + #facet_grid (~GROUP) + 
    theme_classic() + ylab(bquote('Amygdala Volume '*(mm/100)^3*'')) + 
    xlab ("Age (years)") + my_colors + my_colors2 + 
    # add raw data points! 
    geom_point(data = fsdata6, aes(x = brain_age_yrs, y = Amyg_ave, color = GROUP),
              size = 1, alpha = 0.2) + 
    geom_line(data = fsdata6, aes(x = brain_age_yrs, y = Amyg_ave, group = IDENT_SUBID), alpha = 0.1) +
     scale_x_continuous(breaks = seq(4,20.5, by = 3), limits=c(4, 20.5), expand = c(.02,0)) +
 scale_y_continuous(breaks = seq (10, 28, by = 3), limits = c(10, 23), expand = expand_scale(mult = c(0, .1))) +  theme(legend.position="none")

pdf (file = "figures/amygdala_linear_model_overlapping.pdf", width = 6 , height = 4)
amyg_linear_plot_with_raw
dev.off()

pdf (file = "figures/amygdala_linear_model_overlapping_CI.pdf", width = 6 , height = 4)
amyg_linear_plot_with_raw + geom_ribbon(aes(ymin=lower, ymax=upper, 
                  group = GROUP, fill = GROUP), alpha =0.1, color = NA) 
dev.off()

```


## 2) quadratic model: NS
```{r}

amyg_quad <- lmer(Amyg_ave ~ (brain_age_yrs + I(brain_age_yrs^2)) * GROUP +
                sex.c + ICV.c + motion_ave_new.c + scanner_confound.c +
                (1 | IDENT_SUBID), data = fsdata6)
summary(amyg_quad)
amyg_quad_model <- anova(amyg_quad)
save(amyg_quad_model, file = "model_output/Amyg_quadratic_model_results.Rdata")

```

### plot with facet by GROUP and raw data 
```{r, warnings = F}

effect_agesq <- as.data.frame(effect("I(brain_age_yrs^2):GROUP", amyg_quad, confint=TRUE, xlevels = list(brain_age_yrs= age_list)))

### plot starts here. 
quadratic_plot <- ggplot(effect_agesq, aes(x = brain_age_yrs, y = fit, 
                                         color = GROUP )) +
  # plot the lines! 
  theme_classic() + my_colors + my_colors2 +  
   geom_line( aes(group = GROUP, color = GROUP), size =1)  + 
  ylab(bquote('Amygdala Volume '*(mm/100)^3*'')) + xlab("Age (years)")+ theme_classic () + 
   theme(legend.position="none") + 
  #raw data 
   geom_point(data = fsdata6, aes(x= brain_age_yrs, y= Amyg_ave,  
                                color = GROUP), size=1, alpha = 0.2) + 
  geom_line(data = fsdata6,aes(x= brain_age_yrs, y= Amyg_ave,  
                                color = GROUP, group = IDENT_SUBID), alpha = 0.2) +
     scale_x_continuous(breaks = seq(4,20.5, by = 3), limits=c(4, 20.5), expand = c(.02,0)) +
 scale_y_continuous(breaks = seq (10, 28, by = 3), limits = c(10, 23), expand = expand_scale(mult = c(0, .1))) 

pdf(file = "figures/amygdala_quadratic_plot_with_raw.pdf", width = 6, height = 4)
quadratic_plot
dev.off()

# add the CI ribbon...
pdf(file = "figures/amygdala_quadratic_plot_with_raw_CI.pdf", width = 6, height = 4)
quadratic_plot +  geom_ribbon(aes(ymin=lower, ymax=upper, fill = GROUP,
                                    group = GROUP), color = NA, alpha=0.1) 
dev.off() 

# facet groups 
pdf(file = "figures/amygdala_quadratic_plot_with_raw_CI_facet.pdf", width = 6, height = 4)
quadratic_plot +  geom_ribbon(aes(ymin=lower, ymax=upper, fill = GROUP,
                                    group = GROUP), color = NA, alpha=0.1)  + facet_grid(~GROUP)
dev.off() 
  
```


## 3) original piecewise model
### run optimizaton of breakpoint
```{r, warnings = F, message= F, results = F}
#Basis functions
b1 <- function(x, bp) ifelse(x < bp,x, bp)
b2 <- function(x, bp) ifelse(x < bp, bp, x)

# Wrapper for Mixed effects model with variable break point
run_models <- function(bp){
  # this is the formula for your model 
  mod <- lmer(Amyg_ave ~ 
                GROUP*b1(brain_age_yrs, bp)+ 
                GROUP*b2(brain_age_yrs, bp) +
                sex + ICV+ motion_ave_new + scanner_confound +
                (1 | IDENT_SUBID), data = fsdata6)
  deviance(mod, REML = F) # you're minimizing this value 
}

# set the range of breakpoints (age range) you'd allow. 
search.range <- c(min(fsdata6$brain_age_yrs)+0.5,
                  max(fsdata6$brain_age_yrs)-0.5)
# find the optimal model  based on breakpoint. 
mod.opt <- optimize(run_models, interval = search.range)
bp <- mod.opt$minimum # 9.4

# get a confidence interval of your break point. 
CI.bp <- function(bp, tgt)
{
  run_models(bp) - tgt
}
# get info for confidence interval for break point! 
tgt <- mod.opt$objective + qchisq(0.95,1)
lb95 <- uniroot(CI.bp, lower=search.range[1], upper=bp, tgt=tgt)
ub95 <- uniroot(CI.bp, lower=bp, upper=20, tgt=tgt)
bp_lower <-lb95$root
bp_upper <- ub95$root

# make a table to save
optimize_piecewise_bp <- cbind(bp, bp_lower, bp_upper)
optimize_piecewise_bp
save(optimize_piecewise_bp, file = "model_output/Amyg_optimize_piecewise_orig_breakpoint.Rdata")
```

## run final model, with optimal breakpoint
note: with age centered at bp
```{r}
# ages centered at bp
fsdata6$age1 <- b1(fsdata6$brain_age_yrs, bp) - bp
fsdata6$age2 <- b2(fsdata6$brain_age_yrs, bp) - bp
hist(fsdata6$age1) # age1 is zero when > 9.4 yrs
hist(fsdata6$age2) # age2 is zero when < 9.4  yrs


fsdata6$index_wave.c <-fsdata6$index_wave - mean(fsdata6$index_wave)
Amyg_piecewise_orig <- lmer(Amyg_ave ~ 
                GROUP*age1+ 
                GROUP*age2 + 
                sex.c + ICV.c + motion_ave_new.c + scanner_confound +
                (1 | IDENT_SUBID), data = fsdata6)
summary(Amyg_piecewise_orig ) 

# intercept of model at 10.96 = level for comparisons at breakpoint.
save(Amyg_piecewise_orig, file= "model_output/Amyg_piecewise_orig_results.Rdata")

```


## Calculate ICC for amygdala 
(repeated measures within sub)
```{r, warnings = F}
#version 1: basic model 
mod <- lmer(Amyg_ave ~ (1 | IDENT_SUBID), data = fsdata6)
icc(mod)
ICC_amyg <- icc(mod)[1]
save(ICC_amyg, file = "model_output/ICC_for_amygdala.Rdata")

#version 2: from final piecewise mod
ICC_amyg_full_piecewise_model <- icc(Amyg_piecewise_orig)[1]
ICC_amyg_full_piecewise_model 
save(ICC_amyg_full_piecewise_model, file = "model_output/ICC_for_amygdala_full_piecewise_model.Rdata")

```

### plot model fit with raw data

```{r}
# make specific new data for COMP age effect.
# only include age ranges for both groups! 
stuff <- data.frame(age1 = c(min(fsdata6$age1[fsdata6$GROUP == "PI"]),
                             -3, -2.5, -2, -1, -0.001, rep(0, 10)),
                    age2 = c(rep(0, 6), .001, 0.5, 1,2,3,4,5, 6, 7, 
                             max(fsdata6$age2[fsdata6$GROUP == "PI"])), 
                    GROUP = c(rep("COMP", 16), rep("PI", 16)), 
                sex.c = 0, ICV.c= 0, motion_ave_new.c = 0, scanner_confound = 0)
nrow(stuff)

# specify design matrix with fixed effects and new df
fit.mat <- model.matrix(~ GROUP*age1+ 
                GROUP*age2 +
                sex.c + ICV.c + motion_ave_new.c + scanner_confound,
                data = stuff) 
cis <- diag(fit.mat %*% tcrossprod(vcov(Amyg_piecewise_orig), fit.mat))

# predict y values and lwr/upr bounds using model object and new data
stuff$amyg_Y <- predict(Amyg_piecewise_orig, stuff, re.form = NA)
stuff$amyg_lwr <- stuff$amyg_Y-1.96*sqrt(cis)
stuff$amyg_upr <- stuff$amyg_Y+1.96*sqrt(cis)

# clean up the tables for plotting
stuffage1 <- stuff %>%
  filter(age1 < 0) %>%
  mutate(age1 = age1+ bp) %>%
  dplyr::select(-age2) %>%
  rename(brain_age_yrs = age1)

stuffage2 <- stuff %>%
  filter(age2 > 0) %>%
  mutate(age2 = age2 + bp) %>%
  dplyr::select(-age1) %>%
  rename(brain_age_yrs = age2)
```


```{r}
# plot! 
amyg_piecewise_plot <- ggplot(data = stuffage1, aes(x = brain_age_yrs, y = amyg_Y, color = GROUP)) +  geom_line(size = 1) + geom_line(data = stuffage2, aes(x = brain_age_yrs, y = amyg_Y, color = GROUP), size = 1) + 
  my_colors + my_colors2 + theme_classic() +
 ylab(bquote('Amygdala Volume '*(mm/100)^3*'')) + xlab("Age (years)") + 
  geom_point(data = fsdata6,
             aes(x= brain_age_yrs, y= Amyg_ave, color = GROUP), size=1, alpha = 0.2)+ 
  geom_line(data = fsdata6,
            aes(x = brain_age_yrs, y = Amyg_ave, 
                color = GROUP, group = IDENT_SUBID), alpha = 0.2) + 
  scale_y_continuous(breaks = seq (10, 20, by = 2), limits = c(10, 20), expand = expand_scale(mult = c(0, .1))) + 
   scale_x_continuous(breaks = seq(4,21.1, by = 2), limits=c(4, 21.1), expand = c(.02,0)) +
   theme(legend.position= c(0.10, 0.90),  legend.key.height = unit(0.1, "cm"),
        legend.key.width = unit(0.1, "cm"), legend.box.background = element_rect(color = "black"), legend.title = element_blank(), legend.margin=margin(c(2,5,2,2)))


pdf("figures/amygdala_piecewise_model_plot_overlapped.pdf", width = 9, height = 6)
amyg_piecewise_plot
dev.off()

amyg_piecewise_plot <- amyg_piecewise_plot + geom_ribbon(data = stuffage1, aes(ymin = amyg_lwr, ymax = amyg_upr, fill = GROUP), alpha = 0.1, color = NA) +
  geom_ribbon(data = stuffage2, aes(ymin = amyg_lwr, ymax = amyg_upr, fill = GROUP), alpha = 0.1, color = NA) 

pdf("figures/amygdala_piecewise_model_plot_overlapped_CI.pdf", width = 6, height = 4)
amyg_piecewise_plot
dev.off()
```


SAVING TO POWERPT 
```{r}

ppt_test2 <- read_pptx() %>%
 add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with_vg(code = print(amyg_piecewise_plot)) %>%
  print(target = "../../../manuscript/figures/powerpoint_vector_figures/Feb_2020/amygdala_piecewise_model_plot_overlapped_CI.pptx")
  
```


## significance plot of group differences across age
```{r}
# centered at 9.4 = 0
bp = 9.4
age.search.range <- seq(from = min(fsdata6$brain_age_yrs[fsdata6$GROUP == "PI"]) - bp, to =  max(fsdata6$brain_age_yrs[fsdata6$GROUP == "PI"]) - bp, by = 0.25)

amyg_SS_table <- data.frame() 
for (age in age.search.range){
  if(age < 0){ # 0 = breakpoint of age 9 
    ss_test <- data.frame(Age = age + bp, condslope.lmer.pointestimates("GROUPPI", "age1", age, Amyg_piecewise_orig))
   # print(ss_test)
  } else {
    ss_test <- data.frame(Age = age + bp, condslope.lmer.pointestimates("GROUPPI", "age2", age, Amyg_piecewise_orig))
  }
  amyg_SS_table <- rbind(amyg_SS_table, ss_test)
}

save(amyg_SS_table, file = "model_output/Amyg_piecewise_SS_posthoc_group_diff.Rdata")
#names(amyg_SS_table)

# age cut-offs
PI_greater_than_comp <- amyg_SS_table %>%
  filter(p.val < 0.05 & w1.slope > 0) %>%
  dplyr::summarize(min_age = min(Age), 
            max_age = max(Age)) %>%
  mutate(comparison = "PI > COMP")

PI_smaller_than_comp <- amyg_SS_table %>%
  filter(p.val < 0.05 & w1.slope < 0) %>%
  dplyr::summarize(min_age = min(Age), 
            max_age = max(Age)) %>%
  mutate(comparison = "PI < COMP")

amyg_SS_agerange_table <- rbind(PI_greater_than_comp, PI_smaller_than_comp)
amyg_SS_agerange_table
save(amyg_SS_agerange_table, file = "model_output/Amyg_piecewise_orig_posthoc_ageranges_of_group_diff.Rdata")
```


```{r}
amyg_SS_plot <- ggplot(amyg_SS_table, aes(x = Age, y = w1.slope)) + 
  geom_line() + geom_ribbon(aes(ymin = lower95, ymax = upper95), alpha = 0.2) + 
  geom_hline(aes (yintercept = 0), linetype = "dashed") + ylab("PI < COMP      PI > COMP") +
 theme_classic() + ylim(-5,5) + xlab("Age (years)") +
   scale_x_continuous(breaks = seq(4,21.1, by = 2), limits=c(4, 21.1), expand = c(.02,0)) 


amyg_SS_plot
pdf("figures/amyg_piecewise_model_SS_significance_plot.pdf", width =6, height = 4)
amyg_SS_plot
dev.off()

```

SAVING TO PWRPT
```{r}

grid.arrange(amyg_piecewise_plot,amyg_SS_plot)

ppt_test2 <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with_vg(code = print(grid.arrange(amyg_piecewise_plot,amyg_SS_plot))) %>%
  print(target = "../../../manuscript/figures/powerpoint_vector_figures/Feb_2020/stacked_amygdala_piecewise_model_plot_PI_age_range.pptx")
  
```

## GETTING SLOPES FOR AGE EFFECTS, PER GROUP
```{r}
Amyg_piecewise_orig2 <- lmer(Amyg_ave ~  age1 + age2 + GROUP + age1:GROUP + age2:GROUP +
                sex.c + ICV.c + motion_ave_new.c + scanner_confound +
                (1 | IDENT_SUBID), data = fsdata6)
#summary(Amyg_piecewise_orig2)
comp_slope_age1 <- data.frame(group = "COMP", condslope.lmer.slopes( "age1","GROUPPI", 0, Amyg_piecewise_orig2 ))
PI_slope_age1 <- data.frame(group = "PI", condslope.lmer.slopes("age1", "GROUPPI", 1, Amyg_piecewise_orig2 ))
comp_slope_age2 <- data.frame(group = "COMP", condslope.lmer.slopes( "age2","GROUPPI", 0, Amyg_piecewise_orig2 ))
PI_slope_age2 <- data.frame(group = "PI", condslope.lmer.slopes("age2", "GROUPPI", 1, Amyg_piecewise_orig2 ))

amyg_posthoc_slopes <- rbind(comp_slope_age1, PI_slope_age1, comp_slope_age2, PI_slope_age2)

names(amyg_posthoc_slopes) <- c("GROUP", "Breakpoint Intercept", "Estimated Slope", "df", "t.value", "p.value", "lower", "upper")
amyg_posthoc_slopes$'Age range' <- c("age1", "age1", "age2", "age2")
amyg_posthoc_slopes
save(amyg_posthoc_slopes, file = "model_output/Amyg_piecewise_orig_posthoc_age_slopes.Rdata")
```


### how many PI data points are ages < 9.4 
```{r}
table_ages_by_group <- fsdata6 %>%
  mutate(bp_age_groups = as.factor(ifelse(brain_age_yrs > 6.5, "age >= 6.5", "age < 6.5"))) %>%
  group_by(GROUP, bp_age_groups) %>%
  
  dplyr::summarize(n_tp = n(), 
            n_subs = length(unique(IDENT_SUBID)))
table_ages_by_group

save(table_ages_by_group, file = "tables/N_data_points_by_group_and_agebp.Rdata")
```


### double check group differences > 9.4
```{r}
Amyg_adolescent_group_mod <- lmer(Amyg_ave ~ GROUP + brain_age_yrs.c +
                sex.c + ICV.c + motion_ave_new.c + scanner_confound +
                (1 | IDENT_SUBID), data = subset(fsdata6, brain_age_yrs > bp))
summary(Amyg_adolescent_group_mod)
save(Amyg_adolescent_group_mod, file = "model_output/amyg_group_diff_adolescents_model_results.Rdata")
```

# model comparisons 
conflicting results depending on whether use anova() or AIC() 
that means anova() is calculating AIC() differently.... 
```{r}

anova(amyg_linear, amyg_quad) # big improvement by adding quadratic
anova(amyg_quad, Amyg_piecewise_orig) 

AIC_table <- AIC(amyg_linear, amyg_quad, Amyg_piecewise_orig)
BIC_table <- BIC(amyg_linear, amyg_quad, Amyg_piecewise_orig)

# get these into nicer table formrat
model_comparison_amyg <- data.frame(cbind(AIC_table, BIC_table$BIC))
names(model_comparison_amyg) <- c("DF", "AIC", "BIC")
rownames(model_comparison_amyg) <- c("Linear", "Quadratic", "Piecewise")
model_comparison_amyg$model <- rownames(model_comparison_amyg)
save(model_comparison_amyg, file = "model_output/model_comparisons_amyg.Rdata")

model_comparison_amyg
```


