---
title: "supplemental results for cross-lag models"
author: "Michelle.VanTieghem"
date: "Jan 23, 2019"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# Readme
supplemental analyses for cross-lag SEM \
Group interactions in cross-lag paths 


get useful packages
```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
```

# prep 
## load data 
```{r}

# this one compiled more subjects with 2 waves of data. 
load("../../../data/5_SEM_cross_lag_data/cleaned2_all_2wave_cort_brain_wide_format_ages4to19.Rdata")
nrow(df_wide)
```

## function for chi-square test later
```{r}
 # is p < 0.05? test with the chi-square distribution for this DF
chi_square_test <- function(chi_sq_diff, df_diff) {
  if (chi_sq_diff > qchisq(.95, df=df_diff)){
    p  = '< 0.05'
  } else{
    p = '> 0.05' 
  }
  return(list(chi_square_difference = chi_sq_diff, df_difference = df_diff, p.value = p))
}

```

## load main models to compare later
```{r}


load("../manuscript/model_output/hipp_waking_cross_lagged_model_main_effects.Rdata")

fit_measures <- data.frame(fitMeasures(hipp_mod1)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "hipp_mod1"

hipp_mod1_fit_measures <- fit_measures %>%
  spread(metric, value)

```

```{r}

load("../manuscript/model_output/amyg_waking_cross_lagged_model_main_effects.Rdata")

fit_measures <- data.frame(fitMeasures(amyg_mod1)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "amyg_mod1"

amyg_mod1_fit_measures <- fit_measures %>%
  spread(metric, value)

```

# 1) Hippocampus models

## GROUP interaction with cort 
### make interaction regressors 
```{r}
# here the comps are coded as 0, PIs as 1 
# how do I interpret the direction of this interaction? 

df_wide$Hipp_wave1_rs.cXIDENT_SUBTYPE <- with(df_wide, Hipp_wave1_rs.c*IDENT_SUBTYPE)
df_wide$waking_wave1.cXIDENT_SUBTYPE <- with(df_wide, waking_wave1.c*IDENT_SUBTYPE)
df_wide$age_wave1.cXIDENT_SUBTYPE <- with(df_wide, Age_wave1.c*IDENT_SUBTYPE)
# here the comps are coded as 0, PIs as 1 
# how do I interpret the direction of this interaction? 
df_wide$IDENT_SUBTYPE_COMP1 <- ifelse(df_wide$IDENT_SUBTYPE== 0, 1, 0)

df_wide$Hipp_wave1_rs.cXIDENT_SUBTYPE1 <- with(df_wide, Hipp_wave1_rs.c*IDENT_SUBTYPE_COMP1)
df_wide$waking_wave1.cXIDENT_SUBTYPE1 <- with(df_wide, waking_wave1.c*IDENT_SUBTYPE_COMP1)
```


could not include all of the paths for the model to converge
fit statistics are not very good! 
```{r }
model= 
'
#Regression
Hipp_wave2_rs.c ~ waking_wave1.cXIDENT_SUBTYPE + Hipp_wave1_rs.c + waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE  + ICV_wave2_rs.c + sex.c + scanner_wave2
waking_wave2.c ~ Hipp_wave1_rs.cXIDENT_SUBTYPE + Hipp_wave1_rs.c + waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE  + sex.c 

Hipp_wave1_rs.c ~ Age_wave1.c + IDENT_SUBTYPE  + ICV_wave1_rs.c + sex.c 
waking_wave1.c ~ Age_wave1.c + IDENT_SUBTYPE + sex.c 

# covariance
waking_wave1.c ~~ Hipp_wave1_rs.c
waking_wave2.c ~~ Hipp_wave2_rs.c

# time 1 - time 2 things
ICV_wave1_rs.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Age_wave2.c 

# correlations at each time point 
Age_wave1.c ~ ICV_wave1_rs.c 
Age_wave2.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Hipp_wave1_rs.c 
Age_wave2.c ~ Hipp_wave2_rs.c
Age_wave1.c ~ waking_wave1.c
Age_wave2.c ~ waking_wave2.c
#sex.c	~	Hipp_wave1_rs.c		

#ICV_wave1_rs.c	~	Hipp_wave1_rs.c	
#ICV_wave2_rs.c	~	Hipp_wave2_rs.c		

#IDENT_SUBTYPE	~	Hipp_wave1_rs.c
#IDENT_SUBTYPE ~ sex.c 
#IDENT_SUBTYPE ~ ICV_wave1_rs.c
'

# run the model adn print output
hipp_mod1_group <- sem(model, data=df_wide, fixed.x=F, missing='fiml')
summary(hipp_mod1_group , fit.measures=TRUE, standardized=TRUE, rsquare=TRUE)[1]

fit_measures <- data.frame(fitMeasures(hipp_mod1_group)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "hipp_mod1_group"

hipp_mod1_group_fit_measures <- fit_measures %>%
  spread(metric, value)

save(hipp_mod1_group, file = "model_output/hipp_waking_cross_lagged_model_group_int.Rdata")

```


## comparison of hipp models
```{r}

hipp_model_comparisons <- rbind(hipp_mod1_fit_measures, hipp_mod1_group_fit_measures)

save(hipp_model_comparisons, file = "model_output/hipp_waking_model_fit_indices.Rdata")
```

## testing whether adding group interaction is better or worse model fit
```{r}
hipp_mod1_fit_measures$chisq # model without scanner has better (lower) chi-square
hipp_mod1_group_fit_measures$chisq # model with scanner has higher chi-square 

# is this difference significant?
diff_chi_sq <- abs(hipp_mod1_fit_measures$chisq - hipp_mod1_group_fit_measures$chisq)
diff_df <-  abs(hipp_mod1_fit_measures$df - hipp_mod1_group_fit_measures$df)

# yes!
chi_square_hipp_group_result <- chi_square_test(diff_chi_sq, diff_df)
chi_square_hipp_group_result
save(chi_square_hipp_group_result, file = "model_output/chi_square_hipp_group_result.Rdata")
```

# 2) amygdala models 

## GROUP interaction with cort 
### make interaction regressors 
```{r}
# here the comps are coded as 0, PIs as 1 
# how do I interpret the direction of this interaction? 

df_wide$Amyg_wave1_rs.cXIDENT_SUBTYPE <- with(df_wide, Amyg_wave1_rs.c*IDENT_SUBTYPE)
df_wide$waking_wave1.cXIDENT_SUBTYPE <- with(df_wide, waking_wave1.c*IDENT_SUBTYPE)
df_wide$age_wave1.cXIDENT_SUBTYPE <- with(df_wide, Age_wave1.c*IDENT_SUBTYPE)
# here the comps are coded as 0, PIs as 1 
# how do I interpret the direction of this interaction? 
df_wide$IDENT_SUBTYPE_COMP1 <- ifelse(df_wide$IDENT_SUBTYPE== 0, 1, 0)

df_wide$Amyg_wave1_rs.cXIDENT_SUBTYPE1 <- with(df_wide, Amyg_wave1_rs.c*IDENT_SUBTYPE_COMP1)
df_wide$waking_wave1.cXIDENT_SUBTYPE1 <- with(df_wide, waking_wave1.c*IDENT_SUBTYPE_COMP1)
```

Also would not converge without omitting additional paths
note: had to remove a coefficient to converge, so removed #Age_wave2.c ~ waking_wave2.c
```{r }
model= 
'
#Regression
Amyg_wave2_rs.c ~ waking_wave1.cXIDENT_SUBTYPE + Amyg_wave1_rs.c + waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE  + ICV_wave2_rs.c + sex.c + scanner_wave2
waking_wave2.c ~ Amyg_wave1_rs.cXIDENT_SUBTYPE + Amyg_wave1_rs.c + waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE  +  sex.c 

Amyg_wave1_rs.c ~ Age_wave1.c + IDENT_SUBTYPE  + ICV_wave1_rs.c + sex.c 
waking_wave1.c ~ Age_wave1.c + IDENT_SUBTYPE + sex.c 

# covariance
waking_wave1.c ~~ Amyg_wave1_rs.c
waking_wave2.c ~~ Amyg_wave2_rs.c

# time 1 - time 2 things
ICV_wave1_rs.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Age_wave2.c 

# correlations at each time point 
Age_wave1.c ~ ICV_wave1_rs.c 
Age_wave2.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Amyg_wave1_rs.c 
Age_wave2.c ~ Amyg_wave2_rs.c
Age_wave1.c ~ waking_wave1.c
#Age_wave2.c ~ waking_wave2.c
#sex.c	~	Amyg_wave1_rs.c		

#ICV_wave1_rs.c	~	Amyg_wave1_rs.c	
#ICV_wave2_rs.c	~	Amyg_wave2_rs.c		

#IDENT_SUBTYPE	~	Amyg_wave1_rs.c
#IDENT_SUBTYPE ~ sex.c 
#IDENT_SUBTYPE ~ ICV_wave1_rs.c

'

amyg_mod1_group <- sem(model, data=df_wide, fixed.x=F, missing='fiml')
fit_measures <- data.frame(fitMeasures(hipp_mod1_group)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures)
fit_measures$model <- "amyg_mod1_group"

amyg_mod1_group_fit_measures <- fit_measures %>%
  spread(metric, value)

save(amyg_mod1_group, file = "model_output/amyg_waking_cross_lagged_model_group_int.Rdata")
summary(amyg_mod1_group)
```

## comparison of Amyg models

```{r}

amyg_model_comparisons <- rbind(amyg_mod1_fit_measures, amyg_mod1_group_fit_measures)
save(amyg_model_comparisons, file = "model_output/amyg_waking_model_fit_indices.Rdata")

```


## testing whether adding group interaction is better or worse model fit
```{r}
amyg_mod1_fit_measures$chisq # model without scanner has better (lower) chi-square
amyg_mod1_group_fit_measures$chisq # model with scanner has higher chi-square 

# is this difference significant?
diff_chi_sq <- abs(amyg_mod1_fit_measures$chisq - amyg_mod1_group_fit_measures$chisq)
diff_df <-  abs(amyg_mod1_fit_measures$df - amyg_mod1_group_fit_measures$df)

# yes!
chi_square_amyg_group_result <- chi_square_test(diff_chi_sq, diff_df)
chi_square_amyg_group_result
save(chi_square_amyg_group_result, file = "model_output/chi_square_amyg_group_result.Rdata")
```


# 3) make figures of sig cross-lag paths
## regression for each path 
```{r}

hipp1_waking2 <- lm(waking_wave2.c ~ Hipp_wave1_rs.c + waking_wave1.c + 
     Age_wave1.c + IDENT_SUBTYPE + ICV_wave1_rs.c + sex.c + scanner_wave2, data = df_wide)

hipp1_waking2_results <- summary(hipp1_waking2)
hipp1_waking2_results
save(hipp1_waking2_results, file= "model_output/hipp1_waking2_lm_path_results.Rdata")
```


```{r}

amyg1_waking2 <- lm(waking_wave2.c ~ Amyg_wave1_rs.c + waking_wave1.c + 
     Age_wave1.c + IDENT_SUBTYPE + ICV_wave1_rs.c + sex.c +
     scanner_wave2, data = df_wide)

amyg1_waking2_results <- summary(amyg1_waking2)
amyg1_waking2_results
save(amyg1_waking2_results, file= "model_output/amyg1_waking2_lm_path_results.Rdata")

```

## plots effect of hipp 
```{r}
hipp_effect <- data.frame(effect( "Hipp_wave1_rs.c",hipp1_waking2))

hipp1_waking2_effect_plot <- ggplot(hipp_effect, aes(x = Hipp_wave1_rs.c, y = fit)) +
 geom_ribbon(aes(ymin=lower, ymax=upper), alpha =0.2, color = NA) +
   geom_line(size = 1) + 
   theme_classic() + my_colors + my_colors2 + ylab ("Waking Cortisol T2") + 
   xlab(bquote('Hippocampal Volume T1 '*(mm/100)^3*''))

hipp1_waking2_effect_raw <- hipp1_waking2_effect_plot + 
  geom_point(data = df_wide, aes(x = Hipp_wave1_rs.c, y = waking_wave2.c, color = GROUP), 
             alpha = 0.5, size = 1) +theme(legend.position=c(0.85, 0.15), legend.box.background = element_rect(colour = "black"), legend.title = element_blank(), legend.margin=margin(c(1,5,1,1)))


#pdf(file = "figures/Hipp1_waking2_raw_CI.pdf", width = 4, height = 4) 
hipp1_waking2_effect_raw
#dev.off() 

```


## plot effect of amyg
```{r}

amyg_effect <- as.data.frame(effect("Amyg_wave1_rs.c", amyg1_waking2, confint=TRUE))

amyg1_waking2_effect_plot <- ggplot(amyg_effect, aes(x = Amyg_wave1_rs.c, y = fit)) +
 geom_ribbon(aes(ymin=lower, ymax=upper), alpha =0.2, color = NA) +
   geom_line(size = 1) + 
   theme_classic() + my_colors + my_colors2 + ylab ("Waking Cortisol T2") + 
   xlab(bquote('Amygdala Volume T1 '*(mm/100)^3*''))

amyg1_waking2_effect_raw <- amyg1_waking2_effect_plot + 
  geom_point(data = df_wide, aes(x = Amyg_wave1_rs.c, y = waking_wave2.c, color = GROUP), alpha = 0.5, size = 1) +  theme(legend.position="None") + ylab ("") + xlim(-4, 4)

#pdf(file = "figures/Amyg1_waking2_raw_CI.pdf", width = 4, height = 4) 
amyg1_waking2_effect_raw  
#dev.off() 

```

### combine amyg and hipp plots for supplemental material
```{r}
library(gridExtra)
pdf(file = "figures/Amyg1_Hipp1_waking2_panel_pretty.pdf", width = 6, height = 4)
grid.arrange(hipp1_waking2_effect_raw, amyg1_waking2_effect_raw, ncol = 2)
dev.off()
```
