---
title: making long master dataset with cort & structural brain imaging & symptoms 
author: "Michelle.VanTieghem"
date: "June 12, 2018"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# Notes 
##  making long-dataset of mutlilevel all variables \
combining cort data  and brain data after removing outliers.
 
## updates Oct 3, 2018 \
made a second version that includes all longitudinal cortisol data  \
so instead of 1 row per wave per subject (with summary cort data) \
all cort days and times of day have their own row \
and everything else is repeated! \

## updated Jan 11, 2019 \
To include all brain data ages 4-19 for this version. \

## updated Feb 18, 2019  \
To include all data ages 4-21 \
made RCADS_age - based on wave completed. \
made combiend age variable - across cort, brain, RCADS  \
created age chart for paper.

```{r, include = F, warnings =F}
source("../0_R_analysis_setup_file.R")

```


# load data file for longitudinal cort: 1 row per data point 
```{r}

load("../../data/1_NEW_cort_and_T_marta_version_data/Cortisol_Ready_to_analyze.Rdata")
nrow(cort_merge8)
```

## summarize cortisol across days - so just one waking / slope value per wave. 
cort_merge9 has 1 row per wave per subject
```{r}
N_cort_check <- length(unique(cort_merge8$IDENT_SUBID))
N_cort_check

cort_merge8$GROUP.n <- as.numeric(ifelse(cort_merge8$GROUP == "COMP", 0, 1))
cort_merge9 <- cort_merge8 %>% 
  group_by(IDENT_SUBID, index_wave) %>%
  summarise(GROUP = mean(GROUP.n), 
            waking = mean(cort_values_fixed_no_outliers[time_of_day == "waking"],na.rm =T),
            evening = mean(cort_values_fixed_no_outliers[time_of_day == "evening"],na.rm =T),
            slope = mean(cort_values_fixed[time_of_day == "waking"],na.rm =T) - 
            mean(cort_values_fixed_no_outliers[time_of_day == "evening"],na.rm =T), 
            testos_normed = mean(testos_cleaned_log_wave_ave_gender_normed, na.rm = T),
            testos = mean(testos_cleaned_log_wave_ave, na.rm = T), 
            corrected_cort_age_mos = mean(corrected_cort_age_mos, na.rm = T),
            corrected_cort_age_yrs = mean(corrected_cort_age_yrs, na.rm = T),
            meds_by_wave = mean(as.numeric(meds_by_wave), na.rm = T), 
            batch = mean(batch, na.rm = T)) 
# FIXED BATCH TYPO!!!!
sum(is.na(cort_merge9$batch))
sum(is.na(cort_merge9$corrected_cort_age_yrs))

# check N per wave. 
cort_merge9 %>% 
  group_by(GROUP, index_wave) %>%
  dplyr::summarize(n= n() )
nrow(cort_merge9)
# make sure you didn't lose any subjects in this process! 
N_cort_new <- length(unique(cort_merge9$IDENT_SUBID))
identical(N_cort_check, N_cort_new)


#head(cort_merge9)
```


## make a SUBJECTID_long variable to use for merging later
```{r}
# refactor this so it matches fsdata6
cort_merge9$GROUP <- ifelse(cort_merge9$GROUP == 0, "COMP", "PI")
# make this for mergeing. 
cort_merge9$SUBJECTID_long <- ifelse(cort_merge9$index_wave == 2, 
                                      paste0(cort_merge9$IDENT_SUBID, "_fu1"),
                               ifelse(cort_merge9$index_wave == 3, 
                                      paste0(cort_merge9$IDENT_SUBID, "_fu2"), 
                                      as.character(cort_merge9$IDENT_SUBID)))

length(unique(cort_merge9$IDENT_SUBID))
length(unique(cort_merge9$SUBJECTID_long)) 
nrow(cort_merge9)
#head(cort_merge9)
```


# add freesurfer data
note: this is already cleaned, and only ages 6-19.
fsfdata6 has 1 row per subject per wave
```{r}
load("../../data/2_fs_data/3_cleaned_fsdata_for_growth_chart_brain_age_yrs4_19.Rdata")
nrow(fsdata6)
range(fsdata6$brain_age_yrs) 
range(cort_merge9$corrected_cort_age_yrs) 
```

## merge cort dataset with brain dataset 
note: using "type = full" to keep all data with brains, even if they don't have cort.
this is just basically for later, calculating who does and doesn't have what! 
merged_df still has 1 row per subject per wave 
```{r}

# MERGING ALL TOGETHER AND KEEPING ALL ! 
merged_df <- merge(cort_merge9, fsdata6, by = c("SUBJECTID_long", "IDENT_SUBID", "index_wave", "GROUP"), all = T)
nrow(merged_df)
#head(merged_df)

```


# fix formatting & add variables 
```{r}
# chek everyone has a subid.
sum(is.na(merged_df$IDENT_SUBID))
nrow(merged_df)

# check that wave got encoded corretly. 
summary(merged_df$index_wave)

```

## add more variables you need from master! 
only getting partial master data, stuff i will need.
```{r}

master <- read.csv("../../data/master/J.Data_Entry_Master_8_17_2017_fixed_fixed.csv")

adopt_vars <- names(master)[grepl("ADOPT", names(master)) |grepl("IAI", names(master)) ]
med_vars <- names(master)[grepl("MEDS", names(master))]
RCADS_vars <- names(master)[grepl("RCADS", names(master))]
CBCL_vars <- names(master)[grepl("CBCL_4", names(master))]
genotype <-  names(master)[grepl("BDNF", names(master))]
demo <- names(master)[grepl("DEM", names(master))]
ident <- names(master)[grepl("IDENT", names(master))]
ident <- ident[1:6]
session <- names(master)[grepl("ession", names(master))]
pub_vars <- names(master)[grepl("PDS", names(master))]
master2 <- master %>%
  dplyr::select(WASI_FSIQ,  ident, demo,session, adopt_vars, med_vars, RCADS_vars, CBCL_vars, genotype, pub_vars)

# add in these variables into long format! 
# note not using "full" option here, because don't want anyone included 
# in the dataset who doesn't have at least ONE VARIABLE of interest.
merged_final <- merge(merged_df, master2, by = "IDENT_SUBID")
nrow(merged_final)
#head(merged_final)
nrow(merged_df)
summary(as.factor(merged_final$IDENT_SUBTYPE))
# for now, keep everyone in here so we ca check
```


## add symptoms in long -repeated format! 
add columns to long dataset that consist of repeated symptoms at each wave
```{r, warnings = F, include = F }
sublist <- unique(merged_final$IDENT_SUBID)
length(sublist)
#for each subject{
for (subject in sublist){
######### WAVE 1
 merged_final$CBCL_tot_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 1] <-
  merged_final$CBCL_4_TOTPROB_T[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
 
  merged_final$CBCL_tot_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 1] <-
  merged_final$CBCL_4_TOTPROB_TOT[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
 
  
  merged_final$CBCL_int_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 1] <-
  merged_final$CBCL_4_INTERNPROB_T[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
  
    merged_final$CBCL_int_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 1] <-
  merged_final$CBCL_4_INTERNPROB_TOT[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
 
   merged_final$CBCL_ext_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 1] <-
  merged_final$CBCL_4_EXTERNPROB_T[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
   
   
   merged_final$CBCL_ext_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 1] <-
  merged_final$CBCL_4_EXTERNPROB_TOT[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
   
    merged_final$RCADS_int_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 1] <-
  merged_final$J.RCADS_Total_Score_T_new[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
    
    
    merged_final$RCADS_int_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 1] <-
  merged_final$J.RCADS_Total_Score_raw_new[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
 
     merged_final$RCADS_anx_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 1] <-
  merged_final$J.RCADS_Anxiety_Total_T_new[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
     
     
     merged_final$RCADS_anx_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 1] <-
  merged_final$J.RCADS_Anxiety_Total_raw_new[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]

          merged_final$RCADS_MDD_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 1] <-
  merged_final$J.RCADS_MDD_T_new[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]
          
          
          merged_final$RCADS_MDD_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 1] <-
  merged_final$J.RCADS_MDD_raw_new[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 1]

     
     ####### WAVE 2
      merged_final$CBCL_tot_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 2] <-
  merged_final$CBCL_4_TOTPROB_T_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
      
      
      merged_final$CBCL_tot_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 2] <-
  merged_final$CBCL_4_TOTPROB_TOT_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
 
  merged_final$CBCL_int_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 2] <-
  merged_final$CBCL_4_INTERNPROB_T_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
  
  
  merged_final$CBCL_int_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 2] <-
  merged_final$CBCL_4_INTERNPROB_TOT_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
 
   merged_final$CBCL_ext_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 2] <-
  merged_final$CBCL_4_EXTERNPROB_T_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
   
  
   merged_final$CBCL_ext_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 2] <-
  merged_final$CBCL_4_EXTERNPROB_TOT_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
   
    merged_final$RCADS_int_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 2] <-
  merged_final$J.RCADS_Total_Score_t_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
    
    
    merged_final$RCADS_int_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 2] <-
  merged_final$J.RCADS_Total_Score_raw_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
 
 
     merged_final$RCADS_anx_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 2] <-
  merged_final$J.RCADS_Anxiety_Total_t_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
     
     
     merged_final$RCADS_anx_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 2] <-
  merged_final$J.RCADS_Anxiety_Total_raw_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]

        merged_final$RCADS_MDD_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 2] <-
  merged_final$J.RCADS_MDD_T_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]
        
        
        merged_final$RCADS_MDD_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 2] <-
  merged_final$J.RCADS_MDD_raw_FU1[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 2]

    ######### WAVE 3
      merged_final$CBCL_tot_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 3] <-
  merged_final$CBCL_4_TOTPROB_T_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
      
      
      merged_final$CBCL_tot_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 3] <-
  merged_final$CBCL_4_TOTPROB_TOT_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
 
  merged_final$CBCL_int_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 3] <-
  merged_final$CBCL_4_INTERNPROB_T_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
  
  
  merged_final$CBCL_int_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 3] <-
  merged_final$CBCL_4_INTERNPROB_TOT_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
 
   merged_final$CBCL_ext_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 3] <-
  merged_final$CBCL_4_EXTERNPROB_T_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
   
   
   merged_final$CBCL_ext_raw_long[merged_final$IDENT_SUBID == subject &   merged_final$index_wave == 3] <-
  merged_final$CBCL_4_EXTERNPROB_TOT_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
   
    merged_final$RCADS_int_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 3] <-
  merged_final$J.RCADS_Total_Score_t_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
 
    merged_final$RCADS_int_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 3] <-
  merged_final$J.RCADS_Total_Score_raw_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
    
     merged_final$RCADS_anx_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 3] <-
  merged_final$J.RCADS_Anxiety_Total_t_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
     
     
     merged_final$RCADS_anx_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 3] <-
  merged_final$J.RCADS_Anxiety_Total_raw_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]

       merged_final$RCADS_MDD_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 3] <-
  merged_final$J.RCADS_MDD_T_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]
       
       
       merged_final$RCADS_MDD_raw_long[merged_final$IDENT_SUBID == subject &  merged_final$index_wave == 3] <-
  merged_final$J.RCADS_MDD_raw_FU2[merged_final$IDENT_SUBID == subject & merged_final$index_wave == 3]

     }

check_vars <- merged_final[names(merged_final)[grepl("long", names(merged_final))]]
#head(check_vars)
```


## add puberty scores at each wave

```{r}
# Calculating mean Petersen development score! # skip number 5 for females 
# http://www.sleepforscience.org/contentmgr/showdetails.php/id/91

# calculate means first. 
merged_final$PDS_female_mean <- ifelse(merged_final$index_wave == 1, ((merged_final$PPDS_F_1 + merged_final$PPDS_F_2 + merged_final$PPDS_F_3 + merged_final$PPDS_F_4 +merged_final$PPDS_F_5)/5),
                                ifelse(merged_final$index_wave == 2, ((merged_final$PPDS_F_1_FU1 + merged_final$PPDS_F_2_FU1 + merged_final$PPDS_F_3_FU1 + merged_final$PPDS_F_4_FU1 +merged_final$PPDS_F_5_FU1)/5),
                                  ifelse(merged_final$index_wave == 3, ((merged_final$PPDS_F_1_FU2 + merged_final$PPDS_F_2_FU2 + merged_final$PPDS_F_3_FU2 + merged_final$PPDS_F_4_FU1 +merged_final$PPDS_F_5_FU2)/5), NA)))

merged_final$PDS_male_mean <- ifelse(merged_final$index_wave == 1, ((merged_final$PPDS_M_1 + merged_final$PPDS_M_2 + merged_final$PPDS_M_3 + merged_final$PPDS_M_4 +merged_final$PPDS_M_5)/5),
                                ifelse(merged_final$index_wave == 2, ((merged_final$PPDS_M_1_FU1 + merged_final$PPDS_M_2_FU1 + merged_final$PPDS_M_3_FU1 + merged_final$PPDS_M_4_FU1 +merged_final$PPDS_M_5_FU1)/5),
                                  ifelse(merged_final$index_wave == 3, ((merged_final$PPDS_M_1_FU2 + merged_final$PPDS_M_2_FU2 + merged_final$PPDS_M_3_FU2 + merged_final$PPDS_M_4_FU1 +merged_final$PPDS_M_5_FU2)/5), NA)))

merged_final$PDS_mean <- ifelse(merged_final$DEM_3_GENDER_CHILD.y == 0, merged_final$PDS_male_mean, merged_final$PDS_female_mean)

merged_final$has_puberty_baseline <-ifelse(!is.na(merged_final$PPDS_M_5) | !is.na(merged_final$PPDS_F_5), 1, 0)
merged_final$has_puberty_FU1 <- ifelse(!is.na(merged_final$PPDS_M_5_FU1) | !is.na(merged_final$PPDS_F_5_FU1), 1, 0)
merged_final$has_puberty_FU2 <- ifelse(!is.na(merged_final$PPDS_M_5_FU2) | !is.na(merged_final$PPDS_F_5_FU2), 1, 0)
# lots of NAs...
sum(is.na(merged_final$PDS_mean))
 # we have very little puberty data!!
sum(!is.na(merged_final$PDS_mean))

nrow(merged_final)
summary(merged_final$PDS_mean[merged_final$index_wave==1]) # sooo many missing from wave1 !!!! 
summary(merged_final$PDS_mean[merged_final$index_wave==2])
summary(merged_final$PDS_mean[merged_final$index_wave==3])
```

## calculate an age for RCADS.
```{r}
# there are subjects missing cort
sum(is.na(merged_final$corrected_cort_age_yrs))
sum(is.na(merged_final$waking))

# there are subjects missing brain.
sum(is.na(merged_final$brain_age_yrs))
sum(is.na(merged_final$ICV))

#  there are subjects missing RCADS. 
sum(is.na(merged_final$RCADS_int_long))

# get the RCADS age, based on sessions. 
# RCADS baseline would have been collected at session 1 (behavior), not session 2 (first scan)
# RCADS fu1 collected at session3 (scan fu1) and RCADS fu2 collected at session 4 (scan fu2)
merged_final$RCADS_age <- with(merged_final, 
                               ifelse((!is.na(RCADS_int_long) & index_wave == 1), 
                                 SUBAGE_Session1.y, 
                                 ifelse((!is.na(RCADS_int_long) & index_wave == 2),
                                 SUBAGE_session3,
                                 ifelse((!is.na(RCADS_int_long) & index_wave == 3),
                                 SUBAGE_session4, NA))))
sum(is.na(merged_final$RCADS_age))

```

## check age ranges across brain, cort, and RCADS 
just doing some sanity checks.
```{r}
### check age ranges 
range(merged_final$RCADS_age/12, na.rm = T) 
# this is because RCADS was collected before first scan/ cortisol 

range(merged_final$brain_age_yrs[merged_final$GROUP == "PI"], na.rm = T)
range(merged_final$brain_age_yrs[merged_final$GROUP == "COMP"], na.rm = T)

range(merged_final$corrected_cort_age_yrs[merged_final$GROUP == "PI"], na.rm = T)
range(merged_final$corrected_cort_age_yrs[merged_final$GROUP == "COMP"], na.rm = T)

# make an 'age' variable, that takes the ages available for a given time point. 
monster_SB <- merged_final %>%
  mutate(age = ifelse( !is.na(corrected_cort_age_yrs), corrected_cort_age_yrs, 
                      ifelse(!is.na(brain_age_yrs), brain_age_yrs,
                             ifelse(!is.na(RCADS_age), RCADS_age/12,NA))))

monster_SB %>%
  group_by(GROUP) %>%
  dplyr::summarize(min_cort_age = min(corrected_cort_age_yrs, na.rm = T), 
            min_brain_age = min(brain_age_yrs, na.rm = T), 
            min_age = min(age, na.rm = T))
range(monster_SB$corrected_cort_age_yrs[merged_final$GROUP == "PI"], na.rm = T)
range(merged_final$corrected_cort_age_yrs[merged_final$GROUP == "COMP"], na.rm = T)

```

## save data 
this has one row per wave per subject
```{r}

save(monster_SB, file= "../../data/3_ALL_long_cort_symptoms_brain_structure_all_ages_2019-11-12.Rdata")

```



