---
title: "Demographics for entire sample"
author: "Michelle.VanTieghem"
date: " November 13, 2019"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# Ntes
making demographic tables and age chart

```{r, include = F, warnings =F}
source("../0_R_analysis_setup_file.R")

```

# load data 
combined long dataset of all variables. 
```{r}
load("../../data/3_ALL_long_cort_symptoms_brain_structure_all_ages_2019-11-12.Rdata")

```

## filter out non-PI or comps 
```{r}
monster_SB <- subset(monster_SB, IDENT_SUBTYPE == 0 | IDENT_SUBTYPE == 1)
nrow(monster_SB)
monster_SB$GROUP <- as.factor(ifelse(monster_SB$IDENT_SUBTYPE == 0, "COMP", "PI"))
summary(monster_SB$GROUP)

monster_SB$GROUP.c <- ifelse(monster_SB$IDENT_SUBTYPE == 0, -.5, .5)
range(monster_SB$age)

```

## keep data points with CORT OR BRAIN
double checking, no data omitted here.
```{r}
demo_df <- monster_SB  %>% 
 filter(!is.na(Hipp_ave) | !is.na(waking))
nrow(demo_df)

# should be 240 subjects with cort 
check <- monster_SB %>%
  filter(!is.na(waking))
length(unique(check$IDENT_SUBID))
```


# make age- sampling graph 
*use subjects who have cort or brain - and then symptoms subset will be smaller bc excluding ppl who are missing rCADS.
## get min age for each subject. 
```{r}
min_age <- demo_df %>%
  group_by(IDENT_SUBID) %>%
  dplyr::summarize(min_age = min(age))
for (i in 1:nrow(demo_df)){
  sub <- demo_df$IDENT_SUBID[i]
  demo_df$min_age[i] <- min_age$min_age[min_age$IDENT_SUBID== sub]
}

demo_df <- demo_df %>%
  mutate(IDENT_SUBID.f = fct_reorder(IDENT_SUBID, min_age))
min(demo_df$min_age)
```


## graph
```{r}
AgebyParticipant_cortbrain <-ggplot(demo_df,
                         aes(colour=GROUP,
                             y=IDENT_SUBID.f, x= age))+
  geom_line(aes(group = IDENT_SUBID), size=.7, alpha = 0.7) + 
  ylab("Participants") + xlab("Age") +
  geom_point(size=1,  alpha = 0.7) + 
  theme_classic() + my_colors + facet_grid (~ GROUP) +
  theme(axis.line = element_line(colour = "black"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position="none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) + 
   scale_x_continuous(breaks = seq(4,19, by = 3), limits=c(3.8, 19)) 


pdf(file = "tables/Manuscript_Figure_AgebyParticipant_all_data_cort_brain_4to19.pdf", width = 6, height =4)
AgebyParticipant_cortbrain
dev.off()

AgebyParticipant_cortbrain
```

SAVE AS PWRPT
```{r}
#ppt_test2 <- read_pptx() %>%
 # add_slide(layout = "Title and Content", master = "Office Theme") %>%
  #ph_with_vg(code = print(AgebyParticipant_cortbrain)) %>%
  #print(target = "../../manuscript/figures_for_PNAS/powerpoint_vector_figures/Age_sampling.pptx")
```

# make demographic tables for manuscript
## table 1: all subjects, and including PI-specific adoption info
```{r}

demo_df <- demo_df %>%
  # to get 1 row per subject, use their min age row
  filter(age == min_age) %>%
  mutate(sex =  ifelse(!is.na(DEM_3_GENDER_CHILD.x), DEM_3_GENDER_CHILD.x, 
                       DEM_3_GENDER_CHILD.y), 
    country_origin = ifelse(GROUP == "COMP", NA,
                     ifelse(IAI_1A_COUNTRYADOPT == "Russia" |IAI_1A_COUNTRYADOPT == "Kazakhstan" |
                        IAI_1A_COUNTRYADOPT == "Russia" | IAI_1A_COUNTRYADOPT == "Ukraine" | IAI_1A_COUNTRYADOPT == "Azerbaijan" |IAI_1A_COUNTRYADOPT == "Slovak Republic" | IAI_1A_COUNTRYADOPT == "Hungary" |
  IAI_1A_COUNTRYADOPT == "Romania" | IAI_1A_COUNTRYADOPT == "Belarus", "Eastern Europe", 
  ifelse(IAI_1A_COUNTRYADOPT == "Guatamala" | IAI_1A_COUNTRYADOPT ==  "India", "other",
         ifelse(IAI_1A_COUNTRYADOPT == " ", "unknown", "Asia")))))
summary(as.factor(demo_df$country_origin))
```

## make actual table 
```{r}

demo_table1 <- demo_df %>%
  group_by(GROUP) %>%
  dplyr::summarize( N = n(), 
             N_female = sum(sex),
             perc_female = round(sum(sex)/n(), 2)*100,
             mean_age = round(mean(age, na.rm =T),2),
             median_age = round(median(age, na.rm = T), 2),
             sd_age = round(sd(age, na.rm = T),2),
             min_age = round(min(age, na.rm = T),2),
             max_age = round(max(age, na.rm = T),2),
             mean_IQ = round(mean(WASI_FSIQ, na.rm = T), 2), 
             sd_IQ = round(sd(WASI_FSIQ, na.rm = T), 2),
             min_IQ = round(min(WASI_FSIQ, na.rm = T), 2),
             max_IQ = round(max(WASI_FSIQ, na.rm = T), 2),
             mean_age_placed = round(mean(IAI_3_AGEORPH, na.rm = T), 2),
             median_age_placed = round(median(IAI_3_AGEORPH, na.rm = T), 2),
             sd_age_placed = round(sd(IAI_3_AGEORPH, na.rm= T), 2),
             min_age_placed = round(min(IAI_3_AGEORPH, na.rm= T), 2),
             max_age_placed = round(max(IAI_3_AGEORPH, na.rm= T), 2),
      
             mean_age_adopted = round(mean(IAI_7_AGEADOPT, na.rm = T), 2),
             median_age_adopted = round(median(IAI_7_AGEADOPT, na.rm = T), 2),

             sd_age_adopted = round(sd(IAI_7_AGEADOPT, na.rm= T), 2),
             min_age_adopted = round(min(IAI_7_AGEADOPT, na.rm= T), 2),
             max_age_adopted = round(max(IAI_7_AGEADOPT, na.rm= T), 2),
             N_asian_origin = sum(ifelse(country_origin == "Asia", 1, 0), na.rm = T),
            N_eastern_europ_origin = sum(ifelse(country_origin == "Eastern Europe", 1, 0), na.rm = T),
            N_other = sum(ifelse(country_origin == "other", 1, 0), na.rm = T), 
            N_unknown = sum(ifelse(country_origin == "unknown", 1, 0), na.rm = T),
             Perc_Asian_origin = round(sum(ifelse(country_origin == "Asia", 1, 0))/n(),2)*100,
             Perc_Eastern_europ_origin = round(sum(ifelse(country_origin == "Eastern Europe", 1, 0))/n(),2)*100,
             Perc_other = round(sum(ifelse(country_origin == "other", 1, 0))/n(),2)*100,
              Perc_unknown = round(sum(ifelse(country_origin == "unknown", 1, 0))/n(),2)*100)
# flip rows so that PIs are on top.
demo_table <- rbind(demo_table1[2,], demo_table1[1,])

## save 
save(demo_table, file = "tables/new_demo_table_by_subj.Rdata")
```

## add t-tests 
```{r}
# SEX 
sex.ttest <- t.test(sex ~ GROUP, data = demo_df)
# AGE 
age.ttest <- t.test(min_age ~ GROUP, data = demo_df)
# IQ 
IQ.ttest <- t.test(WASI_FSIQ ~ GROUP, data = demo_df)


save(sex.ttest, age.ttest, IQ.ttest, file = "tables/demo_ttest_group_diff_entier_sample.rda")
```



## table 2: breakdown by wave of cort, and brain, by WAVE, side by side panels.
```{r, include = F, eval = F}
demo_waking <- demo_df %>%
  filter(!is.na(waking)) %>%
  group_by(GROUP, index_wave) %>%
  dplyr::summarize(N = n(), 
            N_female = sum(sex),
             perc_female = round(sum(sex)/n(), 2)*100,
             mean_age = round(mean(age, na.rm =T),2),
             sd_age = round(sd(age, na.rm = T),2), 
            min_age = round(min(age, na.rm = T), 2), 
            max_age = round(max(age, na.rm = T), 2))
demo_waking

demo_brain <- demo_df %>%
    filter(!is.na(Hipp_ave)) %>%
   group_by(GROUP, index_wave) %>%
   dplyr::summarize(N = n(), 
             N_female = sum(sex),
             perc_female = round(sum(sex)/n(), 2)*100,
             mean_age = round(mean(age, na.rm =T),2),
             sd_age = round(sd(age, na.rm = T),2), 
            min_age = round(min(age, na.rm = T), 2), 
            max_age = round(max(age, na.rm = T), 2))

demo_brain
cort_brain_demo_by_wave <- cbind(demo_waking, demo_brain) %>%
  dplyr::select(-GROUP1, -index_wave1)
save(cort_brain_demo_by_wave, file = "tables/new_cort_brain_demo_by_wave.Rdata")
```



sanity check: some subjects without cort did have brain? YES.
these are cumulative data points in long format across waves

```{r}

demo_check <- demo_df %>%
  filter(is.na(waking) & !is.na(Hipp_ave)) %>%
  group_by(GROUP) %>%
  dplyr::summarize(N = n())

demo_check
```


sanity check: some subjects without brain did have cort? YES
these are cumulative data points in long format across waves
```{R}

demo_check <- demo_df %>%
  filter(!is.na(waking) & is.na(Hipp_ave)) %>%
  group_by(GROUP) %>%
  dplyr::summarize(N = n())

demo_check
```