---
title: "Cleaning data for symptom analysis"
author: "Michelle VanTieghem"
date: "July 25, 2018"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# README
prep data for analysis with RCADS symptoms scores


```{r, include = F, echo = F, message = F, warning = F}
source("../../0_R_analysis_setup_file.R")
```

# 1) prep data 
## load data 
```{r}

load("../../../data/3_ALL_long_cort_symptoms_brain_structure_all_ages_2019-11-12.Rdata")
```

## filter out non-PI or comps 
```{r}
# double checking only comps & PIS 
monster_SB <- subset(monster_SB, IDENT_SUBTYPE == 0 | IDENT_SUBTYPE == 1)
nrow(monster_SB) # 452 

# renaming group variable
monster_SB$GROUP <- as.factor(ifelse(monster_SB$IDENT_SUBTYPE == 0, "COMP", "PI"))
summary(monster_SB$GROUP)

# making group centered variable.
monster_SB$GROUP.c <- ifelse(monster_SB$IDENT_SUBTYPE == 0, -.5, .5)
summary(monster_SB$GROUP.c)

```


## set sample with no missing cort or brain or symptoms
this data will be used for symptom analysis. 
```{r}
monster_SB_cleaned <- monster_SB %>%
  filter(!is.na(Hipp_ave) & !is.na(waking) & !is.na(RCADS_int_long))

nrow(monster_SB_cleaned) 
```



# 2) check for outliers by TP and group. 
note: outliers by wave were already excluded, but this is a subsample... 
we are labeling these outliers as "outliers_mixed_effects_sample"
```{r, echo = F}
monster_SB_cleaned <- monster_SB_cleaned[
  with(monster_SB_cleaned, order(GROUP, index_wave)),]

# select variables you want outliers for
var_list <- monster_SB_cleaned %>% 
  dplyr::select(Amyg_ave, Hipp_ave, waking, slope, RCADS_int_long, ICV)
var_list <- names(var_list)

# make an empty dataframe to store new stuff into 
outlier_list <- NA
# run through loop of each variable 
for (var in var_list){
    print (var)
  #var <- "slope"
  # use dplyr magic to get all of the info you need.
  outlier_groups <- monster_SB_cleaned %>% 
    # group by GROUP and TIMEPOINT.
      dplyr::select (var, GROUP, index_wave) %>% 
         group_by(GROUP, index_wave) %>% # sort data by group and tp 
         summarize_all(funs(mean(., na.rm = T), sd(., na.rm = T)))         
    # get the means, SD, upper and lower limit (3 SD above or below mean) for 
  outlier_groups$upper <- outlier_groups$mean + 3*outlier_groups$sd 
  outlier_groups$lower <-  outlier_groups$mean - 3*outlier_groups$sd 
   
   # restart these here, because they will generated anew for each variable.

    outlier_list <- data.frame()
# for each grouping in outlier_groups, find outliers and mark them. 
for (i in 1:nrow(outlier_groups)){
  print(i)
  # which group is this? PI or comp
  group <- outlier_groups$GROUP[i]
  tp <- outlier_groups$index_wave[i]
  # set the max and minimum values of cortisol that we will use as outlier threshold
  lower_limits <- outlier_groups$lower[i]
  upper_limits <- outlier_groups$upper[i]
  # get all of the original data by group & wave
  data_chunk <- subset(monster_SB_cleaned,  GROUP== group & index_wave == tp)
 # make sure you're only calculating outliers for the correct variable 
  variable <- data_chunk %>% dplyr::select(var)
   # mark the outliers as 1
  mark_outliers <- data.frame(ifelse(variable >= upper_limits
                          | variable<= lower_limits, 1, 0))
   # make sure this name matches the variable name (e.g. L hipp)
  names(mark_outliers) <- paste0(var, "_outliers_mixed_effects_sample")
  # save the outlier 1s and 0s into a new column 
  outlier_list <- rbind(outlier_list, mark_outliers)
  # make sure the column name matches the variable name (e.g. L hipp)
  names(outlier_list) <- paste0(var, "_outliers_mixed_effects_sample")
} 
  print ("loop break")
  # add the new column of outliers for each variable (e.g. L hipp) to the orig dataframe. 
   monster_SB_cleaned <- cbind(monster_SB_cleaned, outlier_list) 
   
}
```


## get info on these outliers 
```{r}
# get info on who they are and what wave they are 
mixed_effects_outlier_table <- monster_SB_cleaned %>% 
  group_by(IDENT_SUBTYPE, index_wave) %>%
 dplyr::summarize(N_exclude_amyg = sum(Amyg_ave_outliers_mixed_effects_sample, na.rm = T), 
            N_exclude_hipp = sum(Hipp_ave_outliers_mixed_effects_sample, na.rm  = T), 
            N_exclude_slope = sum(slope_outliers_mixed_effects_sample, na.rm = T), 
            N_exclude_waking = sum(waking_outliers_mixed_effects_sample, na.rm = T), 
            N_exclude_RCADS_int = sum(RCADS_int_long_outliers_mixed_effects_sample, na.rm = T),
            N_exclude_ICV = sum(ICV_outliers_mixed_effects_sample, na.rm=T))
mixed_effects_outlier_table
save(mixed_effects_outlier_table, file = "tables/outlier_table_mixed_effects_cort_brain_symptoms_ages4to19.Rdata")
```

## mark these outliers as NAs.
```{r}
monster_SB_cleaned$Amyg_ave_cleaned <- monster_SB_cleaned$Amyg_ave
monster_SB_cleaned$Amyg_ave_cleaned[monster_SB_cleaned$Amyg_ave_outliers_mixed_effects_sample == 1] <- NA

monster_SB_cleaned$Hipp_ave_cleaned <- monster_SB_cleaned$Hipp_ave
monster_SB_cleaned$Hipp_ave_cleaned[monster_SB_cleaned$Hipp_ave_outliers_mixed_effects_sample == 1] <- NA

monster_SB_cleaned$waking_cleaned <- monster_SB_cleaned$waking
monster_SB_cleaned$waking_cleaned[monster_SB_cleaned$waking_outliers_mixed_effects_sample == 1] <- NA


monster_SB_cleaned$slope_cleaned <- monster_SB_cleaned$slope
monster_SB_cleaned$slope_cleaned[monster_SB_cleaned$slope_outliers_mixed_effects_sample == 1] <- NA

monster_SB_cleaned$RCADS_int_long_cleaned <- monster_SB_cleaned$RCADS_int_long
monster_SB_cleaned$RCADS_int_long_cleaned[monster_SB_cleaned$RCADS_int_long_outliers_mixed_effects_sample == 1] <- NA


monster_SB_cleaned$ICV_cleaned <- monster_SB_cleaned$ICV
monster_SB_cleaned$ICV_cleaned[monster_SB_cleaned$ICV_outliers_mixed_effects_sample == 1] <- NA

```


# 3) get cort/brain/symptoms sample demographics 

set age for this analysis: to RCADS outcome
may be slight variations in RCADS, cort and brain ages.
```{r}
# for this analysis, outcome is RCADS. set ages to be = when RCADS collected.
monster_SB_cleaned <- monster_SB_cleaned %>%
  mutate(age = ifelse(!is.na(RCADS_age), RCADS_age/12, 
          ifelse( !is.na(corrected_cort_age_yrs), corrected_cort_age_yrs, 
                      ifelse(!is.na(brain_age_yrs), brain_age_yrs, NA ))), 
         Age.c = age - mean(age), 
         scanner_confound = ifelse(index_wave == 3, 1, 0), 
         GROUP.c = ifelse(GROUP == "PI", 0.5, -0.5))

```

use updated age for your demographics table. 
```{r}
cort_hipp_dem_table_by_wave <- monster_SB_cleaned %>%
  group_by(GROUP, index_wave) %>%
  dplyr::summarize(n= n() , 
            mean_age = mean(age, na.rm = T), 
            sd_age = sd(age, na.rm = T),
            min_age = min(age, na.rm = T),
            max_age = max(age, na.rm = T),
            N_female = sum(DEM_3_GENDER_CHILD.x, na.rm = T),
            prop_female = mean(DEM_3_GENDER_CHILD.x, na.rm = T))

cort_hipp_dem_table_by_wave

save(cort_hipp_dem_table_by_wave, file = "tables/Demo_cort_hipp_symptoms_mixed_effects_sample_by_wave_ages4to19.Rdata")

```

## get the total N per group.. 
```{r}
N_total_symptoms <- length(unique(monster_SB_cleaned$IDENT_SUBID))
N_PI_symptoms <- length(unique(monster_SB_cleaned$IDENT_SUBID[monster_SB_cleaned$GROUP == "PI"]))
N_COMP_symptoms <- length(unique(monster_SB_cleaned$IDENT_SUBID[monster_SB_cleaned$GROUP == "COMP"]))



cort_hipp_dem_table <- data.frame(group = c("all", "PI", "COMP"), n = c(N_total_symptoms, N_PI_symptoms, N_COMP_symptoms))
cort_hipp_dem_table 

save(cort_hipp_dem_table, file = "tables/Demo_cort_hipp_symptoms_mixed_effects_sample_overall_ages4to19.Rdata")


```


# 4) mean-centering for analysis 
```{r}
# center necessary variables.
monster_SB_cleaned$GROUP.c <- monster_SB_cleaned$IDENT_SUBTYPE - mean(monster_SB_cleaned$IDENT_SUBTYPE)
monster_SB_cleaned$sex.c <- monster_SB_cleaned$sex - mean(monster_SB_cleaned$sex, na.rm = T)
monster_SB_cleaned$scanner_confound <- with(monster_SB_cleaned, ifelse(index_wave == 3, 1, 0))

# rescale variables 
monster_SB_cleaned$Amyg_ave_cleaned_rs <- monster_SB_cleaned$Amyg_ave_cleaned/100
monster_SB_cleaned$Hipp_ave_cleaned_rs <- monster_SB_cleaned$Hipp_ave_cleaned/100
#**IMPT: ICV SCALED EVEN MORE - because tiny coefficients < 0.001 that are significant effects.
monster_SB_cleaned$ICV_cleaned_rs <- monster_SB_cleaned$ICV_cleaned/ 100000

# then center the re-scaled variables 
monster_SB_cleaned$Amyg_ave_cleaned_rs.c <- monster_SB_cleaned$Amyg_ave_cleaned_rs - mean(monster_SB_cleaned$Amyg_ave_cleaned_rs, na.rm = T)
monster_SB_cleaned$Hipp_ave_cleaned_rs.c <- monster_SB_cleaned$Hipp_ave_cleaned_rs - mean(monster_SB_cleaned$Hipp_ave_cleaned_rs, na.rm = T)
monster_SB_cleaned$ICV_cleaned_rs.c <- monster_SB_cleaned$ICV_cleaned_rs - mean(monster_SB_cleaned$ICV_cleaned_rs, na.rm = T)
monster_SB_cleaned$waking_cleaned.c <- monster_SB_cleaned$waking_cleaned - mean(monster_SB_cleaned$waking_cleaned, na.rm = T)
```

SAVE THIS DATA
```{r}
save(monster_SB_cleaned,file= "../../../data/4_all_data_mixed_effects/CLEANED_ALL_long_cort_symptoms_brain_structure_all_ages_2019-11-12.Rdata")

```
