---
title: "Supplemental analyses for cortisol"
author: "Michelle.VanTieghem"
date: "May 12, 2019"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---



```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")

## funtion for age breakpoint splitting
b1 <- function(x, bp) ifelse(x < bp,x, bp)
b2 <- function(x, bp) ifelse(x < bp, bp, x) 

```

functions needed for these analyses
```{r, include = F, echo = F}

######### lmer Version #######
condslope.lmer.slopes <- function(x, z, c, y){
#  condslope.lmer( "age1","GROUP", 0, Amyg_piecewise_orig)
  # x = name of x variable
  # z = name of z variable (moderator)
  # c = conditional value of z
  # y = reg object
  # lmer model must be in this order x + z + x*z
# updated slightly by MVT on March 4, 2019 
# bcause tidy doesn't work. 
out <- summary(y)
xz <- paste(x, z, sep=":")
w0.intercept <- fixef(y)["(Intercept)"] + fixef(y)[z]*c
w1.slope <- fixef(y)[x] + fixef(y)[xz]*c 
#y.cond <- w0.intercept + w1.slope*xvalue
require(broom)
modtidy <- data.frame(summary(y)$coef)
modtidy$rownames <- rownames(modtidy)
coef2.var <- subset(modtidy, rownames == x)$Std..Error^2
coef4.var <- subset(modtidy, rownames == xz)$Std..Error^2
out.vcov <- vcov(y)
cond.se <- sqrt(coef2.var + (c) * (c) * coef4.var + 2 * (c) * out.vcov[x, xz])
t.val <- w1.slope/cond.se
p.val <- 2*(1-pt(abs(t.val), subset(modtidy, rownames == x)$df, lower.tail=T))
lower95 <- w1.slope-qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
upper95 <- w1.slope+qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
# z.out <- z.test(w1.slope, sigma.x= cond.se)
return(list(w0.intercept=round(w0.intercept, digits = 2), w1.slope=round(w1.slope, digits = 2),
            df = subset(modtidy, rownames == x)$df,
            t.val = round(t.val, digits = 2), p.val = round(p.val, digits = 3), 
            lower95 = round(lower95, digits = 2), 
            upper95 = round(upper95, digits = 2)))
}


condslope.lmer.pointestimates <- function(x, z, c, y){
#  condslope.lmer( "age1","GROUP", 0, Amyg_piecewise_orig)
  # x = name of x variable
  # z = name of z variable (moderator)
  # c = conditional value of z
  # y = reg object
  # lmer model must be in this order x + z + x*z
  
# updated slightly by MVT on March 4, 2019 
# bcause tidy doesn't work. 
out <- summary(y)
xz <- paste(x, z, sep=":")
w0.intercept <- fixef(y)["(Intercept)"] + fixef(y)[z]*c
w1.slope <- fixef(y)[x] + fixef(y)[xz]*c 
#y.cond <- w0.intercept + w1.slope*xvalue
require(broom)
modtidy <- data.frame(summary(y)$coef)
modtidy$rownames <- rownames(modtidy)
coef2.var <- subset(modtidy, rownames == x)$Std..Error^2
coef4.var <- subset(modtidy, rownames == xz)$Std..Error^2
out.vcov <- vcov(y)
cond.se <- sqrt(coef2.var + (c) * (c) * coef4.var + 2 * (c) * out.vcov[x, xz])
t.val <- w1.slope/cond.se
p.val <- 2*(1-pt(abs(t.val), subset(modtidy, rownames == x)$df, lower.tail=T))
lower95 <- w1.slope-qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
upper95 <- w1.slope+qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
# z.out <- z.test(w1.slope, sigma.x= cond.se)
return(list(w0.intercept=round(w0.intercept, digits = 2), w1.slope=round(w1.slope, digits = 2),
            df = subset(modtidy, rownames == x)$df,
            t.val = round(t.val, digits = 2), p.val = round(p.val, digits = 3), 
            lower95 = round(lower95, digits = 2), 
            upper95 = round(upper95, digits = 2)))
}


```

# load data 
```{r}
load("../../../data/1_cort_data/Cortisol_Ready_to_analyze.Rdata")

# check this is full 240 N
length(unique(cort_merge8$IDENT_SUBID))
```

## clean data
```{r}
# relabel time of day
cort_merge8$time_of_day <- as.factor(with(cort_merge8,
                                   ifelse(time_of_day == "waking", "Waking", "Evening")))

summary(cort_merge8$time_of_day)

# make some more centered variables
cort_merge8$index_day.c <- as.numeric(as.character(cort_merge8$index_day))
summary(cort_merge8$index_day.c)
cort_merge8$meds.c <- as.numeric(as.character(cort_merge8$meds)) - mean(as.numeric(as.character(cort_merge8$meds)), na.rm = T)

# simplifying name of cort stuff 
cort_merge8$cort_use <- cort_merge8$cort_values_fixed_no_outliers 

cort_merge8$corrected_cort_age_yrs.c <- cort_merge8$corrected_cort_age_yrs - mean(cort_merge8$corrected_cort_age_yrs, na.rm = T)

# make a usable dataset
data_use <- subset(cort_merge8, !is.na(cort_use) & 
                                                  !is.na(batch.c) & !is.na(meds.c) & !is.na(index_wave.c) & !is.na(index_day.c))
```


## make summary dataset for plotting 
this is so that we can plot the raw data, but plot means across both days instead of plotting 2 days per subj (per time of day, per wave...)
```{r}
cort_ave_days_m <- data_use %>%
  group_by(IDENT_SUBID, index_wave, time_of_day) %>%
  # this will provide average cort by each wave, and time of day...
  #aka average across 2 days if that subject has 2 days of data!
  dplyr::summarize(n_cort_samples = n(), 
            cort_use = mean(cort_use, na.rm = T),
            PDS_mean = mean(PDS_mean),
            corrected_cort_age_yrs = mean(corrected_cort_age_yrs), 
            GROUP.n = mean(GROUP.n), 
            sex = mean(sex))
nrow(cort_ave_days_m)
cort_ave_days_m$GROUP <- as.factor(ifelse(cort_ave_days_m$GROUP.n == 0, "COMP", "PI"))

```


# 1)  check retention stats 
## prep data 
```{r, warning = F, message = F}
# get everyone's baseline info 
baseline_demo <- cort_merge8 %>% 
    filter((corrected_cort_age_yrs == min_age) & (index20.n == min_index20)) %>%
   select(IDENT_SUBID, GROUP, corrected_cort_age_yrs, sex, J.RCADS_Total_Score_T_new)
nrow(baseline_demo)

# get number f waves per subject
follow_up_waves <- cort_merge8 %>%
  group_by(IDENT_SUBID, index_wave) %>%
# first just find out who has cort for a given wave
  dplyr::summarize(cort_exists = mean(cort_values_fixed_no_outliers, na.rm = T)) %>%
  # then make a wide df so we have each wave of cort as a column, 
  spread(index_wave, cort_exists) %>%
  rename(wave1 = '1', wave2 = '2', wave3 = '3') %>%
  mutate(wave1 = ifelse(!is.na(wave1), 1, 0), 
         wave2 = ifelse(!is.na(wave2), 1, 0), 
         wave3 = ifelse(!is.na(wave3), 1, 0)) %>%
  # now add counts of how many waves each subject has 
  mutate(total_waves = as.factor((wave1 + wave2 + wave3)))

# check that therer is a single row per subj
identical(nrow(follow_up_waves), length(unique(follow_up_waves$IDENT_SUBID)))

# this provides N for subs with 1,2,3 waves.
summary(as.factor(follow_up_waves$total_waves))

# combine with baseline demo
retention_Df <- merge(baseline_demo, follow_up_waves, by = "IDENT_SUBID") %>%
  mutate(GROUP = ifelse(GROUP == "PI", 1, 0))

summary(retention_Df$GROUP)
```

## ANOVA to test difference between baseline, 2 waves, 3 waves 

```{r}

# testing age differernce
summary(age_test <- lm(corrected_cort_age_yrs ~ total_waves, data = retention_Df))
Anova_age_test <- data.frame(Anova(age_test, type = "III"))

# testing sex difference
summary(sex_test <- glm (sex ~ total_waves,
                         family = binomial, 
                         data = retention_Df))
Anova_sex_test <- data.frame(Anova(sex_test, type = "III"))

# trend difference...
summary(group_test <- glm(GROUP ~ total_waves, 
                  family = binomial, 
                  data = retention_Df))
Anova_group_test <- data.frame(Anova(group_test, type = "III"))

save(Anova_age_test, Anova_sex_test, Anova_group_test, file = "model_output/Retention_anova_tests.rda")
```

## T-test for baseline-only vs. follow-up 
```{r}

retention_Df$one_vs_more_waves <- as.factor(ifelse(retention_Df$total_waves == 1, 1, 2))
summary(retention_Df$one_vs_more_waves)

age_test <- lm(corrected_cort_age_yrs ~ one_vs_more_waves, data = retention_Df)
summary(age_test)

# testing sex difference
sex_test <- lm (sex ~ one_vs_more_waves, data = retention_Df)
summary(sex_test)

# SIG difference!
group_test <- glm(GROUP ~ one_vs_more_waves, 
                  family = binomial, 
                  data = retention_Df)
summary(group_test)

# more liekly that PIs came back 
effect_df <- data.frame(effect("one_vs_more_waves", group_test))
ggplot(effect_df, aes(x = one_vs_more_waves, y = fit)) +
  geom_bar(stat = "identity", alpha = 0.5) + ylim(0,1) + ylab("GROUP") + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.5)
```


# 2) SEX EFFECTS 

## set up data
```{r}
waking_cort_df <- data_use %>% 
  filter(time_of_day == "Waking") %>%
  mutate(sex.f = as.factor(sex))
summary(waking_cort_df$sex.f)
```

## using the same breakpoint as orig waking model 
```{r}
load("../manuscript/model_output/waking_optimize_piecewise_breakpoint.Rdata")
bp <- optimize_piecewise_bp[1]
# center these ages on the bp
# changing this so that age1 and age2 are centered at breakpoint.
waking_cort_df$age1 <- b1(waking_cort_df$corrected_cort_age_yrs, bp) -bp 
waking_cort_df$age2 <- b2(waking_cort_df$corrected_cort_age_yrs, bp) -bp
```

## how male PI data points > 13 (breakpoint): only 6! 
cannot split sexes by age 13 so do not continue with piecewise 
```{r}
cort_sex_age_group_table <- waking_cort_df %>%
  filter( index20.n == min_index20) %>%
  mutate(age_group = ifelse(corrected_cort_age_yrs < 13, "age1", "age2")) %>%
  group_by(GROUP, age_group, sex) %>%
  dplyr::summarize(n=n())
  
cort_sex_age_group_table
save(cort_sex_age_group_table, file = "tables/sex_breakdown_by_age_group.Rdata")
```


## linear sex x group: NS

```{r}
sex_waking_linear <- lmer(cort_use ~ 
                           GROUP*sex.f + corrected_cort_age_yrs + sex.f + 
                index_wave.c + index_day.c + batch.c + meds.c + 
                     + (1 | IDENT_SUBID), 
                      data = waking_cort_df)
summary(sex_waking_linear)
save(sex_waking_linear, file = "model_output/sexXgroup_waking_linear_model_results.Rdata")
```


## linear sex x AGE:  SIG.
upshoot in cortisol is driven by FEMALES

```{r}

agexsex_waking_linear <- lmer(cort_use ~ 
                           GROUP + corrected_cort_age_yrs.c*sex.f +
                index_wave.c + index_day.c + batch.c + meds.c + 
                     + (1 | IDENT_SUBID), 
                      data = waking_cort_df)
summary(agexsex_waking_linear)
save(agexsex_waking_linear, file = "model_output/sexXage_waking_linear_model_results.Rdata")
```

### plot the effect
```{r}
plot(effect("corrected_cort_age_yrs.c:sex.f", agexsex_waking_linear))
```

### post-hoc of sex x age effects
```{r}
## GETTING SLOPES FOR AGE EFFECTS, PER GROUP
male_slope_age <- data.frame(level = "male", condslope.lmer.slopes( "corrected_cort_age_yrs.c","sex.f1", 0, agexsex_waking_linear))
female_slope_age <- data.frame(level =  "female", condslope.lmer.slopes("corrected_cort_age_yrs.c", "sex.f1", 1, agexsex_waking_linear))


## GETTING SEX DIFF IN CORT AT DIFF AGES 
agexsex_waking_linear <- lmer(cort_use ~ 
                           GROUP + sex.f*corrected_cort_age_yrs.c +
                index_wave.c + index_day.c + batch.c + meds.c + 
                     + (1 | IDENT_SUBID), 
                      data = waking_cort_df)
age_mean <- mean(waking_cort_df$corrected_cort_age_yrs)
age_sd <- sd(waking_cort_df$corrected_cort_age_yrs)

sex_diff_adol <- data.frame(level =  as.character(round(age_sd + age_mean),2), condslope.lmer.pointestimates("sex.f1","corrected_cort_age_yrs.c", age_sd, agexsex_waking_linear))

sex_diff_child <- data.frame(level = as.character(round(age_mean - age_sd),2), condslope.lmer.pointestimates( "sex.f1", "corrected_cort_age_yrs.c", -age_sd, agexsex_waking_linear))

sex_waking_posthoc_slopes <- rbind(male_slope_age, female_slope_age, sex_diff_adol, sex_diff_child)

sex_waking_posthoc_slopes <- sex_waking_posthoc_slopes %>%
  dplyr::select(-w0.intercept)
names(sex_waking_posthoc_slopes) <- c("GROUP", "Estimate", "df", "t.value", "p.value", "lower.95%.CI", "upper.95.CI%")
sex_waking_posthoc_slopes
save(sex_waking_posthoc_slopes, file = "model_output/sexXage_waking_linear_posthoc_age_slopes.Rdata")
```

# 3) Age of adoption (continuous, logged)

## set up  variables, automatically excluding COMPS
```{r}
adopt_vars_cort_df <- waking_cort_df %>%
  mutate(duration_adversity = IAI_7_AGEADOPT - IAI_3_AGEORPH) %>%
  mutate(duration_split = as.factor(ifelse(duration_adversity > median(duration_adversity, na.rm = T), "later", "early"))) %>%
           filter(GROUP == "PI")
summary(adopt_vars_cort_df$GROUP)
```

## make wider df with 1 row per subject per wave for plotting
```{r}
adopt_vars_wide_by_wave <- adopt_vars_cort_df %>%
  filter( GROUP == "PI") %>%
  group_by(IDENT_SUBID, index_wave) %>%
  dplyr::summarize(IAI_3_AGEORPH = mean(IAI_3_AGEORPH, na.rm = T),
           IAI_7_AGEADOPT = mean(IAI_7_AGEADOPT, na.rm = T),
           duration_adversity = mean(duration_adversity, na.rm = T),
           corrected_cort_age_yrs = mean(corrected_cort_age_yrs, na.rm = T),
           cort_use = mean(cort_use, na.rm = T), 
           sex = mean(DEM_3_GENDER_CHILD), 
           batch = mean(batch)) %>%
  mutate(age_adopt_log = log(IAI_7_AGEADOPT + 0.4),
         age_adopt_sqrt = sqrt(IAI_7_AGEADOPT))

# number of rows 
nrow(adopt_vars_wide_by_wave) 
# number of subjects
length(unique(adopt_vars_wide_by_wave$IDENT_SUBID))

```

## make wide df with 1 subject per row for descriptives 
```{r}
#first you need to make a 1 row per subj df
adopt_vars_wide <- adopt_vars_cort_df %>%
  filter( GROUP == "PI") %>%
  group_by(IDENT_SUBID) %>%
  dplyr::summarize(IAI_3_AGEORPH = mean(IAI_3_AGEORPH, na.rm = T),
           IAI_7_AGEADOPT = mean(IAI_7_AGEADOPT, na.rm = T),
           duration_adversity = mean(duration_adversity, na.rm = T),
           corrected_cort_age_yrs = mean(corrected_cort_age_yrs, na.rm = T),
           cort_use = mean(cort_use, na.rm = T), 
           sex = mean(DEM_3_GENDER_CHILD), 
           batch = mean(batch)) %>%
  mutate(age_adopt_log = log(IAI_7_AGEADOPT + 0.4),
         age_adopt_sqrt = sqrt(IAI_7_AGEADOPT))
nrow(adopt_vars_wide) # check this = # of unique subjects in df? yes
length(unique(adopt_vars_wide$IDENT_SUBID))

```

## get descriptives
```{r}
adoption_waking_table <- adopt_vars_wide %>%
  # issue is that this has multiple columns per subj. 
  dplyr::summarize(mean_age_inst = mean(IAI_3_AGEORPH, na.rm = T),
            sd_age_inst = sd(IAI_3_AGEORPH, na.rm = T), 
            min_age_inst = min(IAI_3_AGEORPH, na.rm = T),
            max_age_inst = max(IAI_3_AGEORPH, na.rm = T),
            mean_age_adopt = mean(IAI_7_AGEADOPT, na.rm = T), 
            sd_age_adopt = sd(IAI_7_AGEADOPT, na.rm = T), 
            min_age_adopt = min(IAI_7_AGEADOPT, na.rm = T),
            max_age_adopt = max(IAI_7_AGEADOPT, na.rm = T),
            mean_duration = mean(duration_adversity, na.rm = T), 
            sd_duration = sd(duration_adversity, na.rm = T), 
            min_duration = min(duration_adversity, na.rm = T),
            max_duration = max(duration_adversity, na.rm = T))

adoption_waking_table
save(adoption_waking_table, file = "tables/adoption_descriptives_table_cort_sample.Rdata")

N_for_adoption_waking_table <- adopt_vars_wide %>%
            dplyr::summarize (N = n(), 
               has_age_inst = sum(!is.na(IAI_3_AGEORPH)),
            has_age_adopt = sum(!is.na(IAI_7_AGEADOPT)),
            has_duration = sum(!is.na(duration_adversity)),
            missing_age_inst = sum(is.na(IAI_3_AGEORPH)),
            missing_age_adopt = sum(is.na(IAI_7_AGEADOPT)),
            missing_duration = sum(is.na(duration_adversity)))
N_for_adoption_waking_table
save(N_for_adoption_waking_table, file = "tables/adoption_N_missing_table_cort_sample.Rdata")
```

## remove subjects with no adoption data 
```{r}
# using age orph as the exclusion variable, because there is 1 subject who has age ofo adoption but no other info and that's not very helpful! 
adopt_vars_wide <- adopt_vars_wide %>%
  filter(!is.na(IAI_3_AGEORPH))

adopt_vars_cort_df <- adopt_vars_cort_df %>%
  filter(!is.na(IAI_3_AGEORPH))

sum(is.na(adopt_vars_cort_df$IAI_7_AGEADOPT))
length(unique(adopt_vars_cort_df$IDENT_SUBID))
```


## correlations between adoption variables and age/sex
relevant: age and batch are correlated.
```{r}

cor_table_adoption <- adopt_vars_wide %>%
 dplyr:: select(IAI_3_AGEORPH, IAI_7_AGEADOPT,corrected_cort_age_yrs, sex, batch) %>%
  na.omit(.) %>%
  correlation
cor_table_adoption_formatted <- summary(cor_table_adoption)
cor_table_adoption_formatted 
save(cor_table_adoption_formatted, file = "tables/adoption_corr_table_beautiful_formatting.Rdata")
```

## check cortisol for outliers in PI sub-sample
```{r}
hist(adopt_vars_cort_df$cort_use)

# get means and SDs by wave
cort_table <- adopt_vars_cort_df %>%
  group_by(index_wave) %>%
  dplyr::summarize(mean = mean(cort_use), 
            sd = sd(cort_use), 
            upper = mean + 3*sd, 
            lower = mean - 3*sd)

# calculate outliers by wave 
adopt_vars_cort_df <- adopt_vars_cort_df %>%
  group_by(index_wave) %>%
  mutate(cort_outlier = ifelse (index_wave == 1 & cort_use > cort_table$upper[cort_table$index_wave == 1], 1, 
               ifelse(index_wave == 2 & cort_use > cort_table$upper[cort_table$index_wave == 2], 1, 
                ifelse(index_wave == 3 & cort_use > cort_table$upper[cort_table$index_wave == 3], 1, 0 )))) 

# count outliers 
cort_outlier_table <- adopt_vars_cort_df %>%
  dplyr::summarize(n_outlier = sum(cort_outlier))

cort_outlier_table

# exclude outliers 
adopt_vars_df <- adopt_vars_cort_df %>% 
  filter(cort_outlier != 1)
```

## check age of placement: log transform
```{r}
hist(adopt_vars_cort_df$IAI_3_AGEORPH)

#log-transforming... doesn't quite help but better
shapiro.test(adopt_vars_cort_df$IAI_3_AGEORPH)
shapiro.test(log(adopt_vars_cort_df$IAI_3_AGEORPH + 0.4))

# log transform, but change 0 to 0.4 so we don't have negative values.
adopt_vars_cort_df$age_placed_log <- log(adopt_vars_cort_df$IAI_3_AGEORPH  + 0.4)
hist(adopt_vars_cort_df$age_placed_log)

```

## check age of adoption: log transform
```{r}
hist(adopt_vars_cort_df$IAI_7_AGEADOPT)

#log-transforming... doesn't quite help but better
shapiro.test(adopt_vars_cort_df$IAI_7_AGEADOPT)
shapiro.test(log(adopt_vars_cort_df$IAI_7_AGEADOPT))

shapiro.test(sqrt(adopt_vars_cort_df$IAI_7_AGEADOPT))

# log transform, but change 0 to 0.4 so we don't have negative values.
adopt_vars_cort_df <- adopt_vars_cort_df %>%
  mutate(age_adopt_log = log(IAI_7_AGEADOPT + 0.4), 
         age_adopt_sqrt = sqrt(IAI_7_AGEADOPT))

hist(adopt_vars_cort_df$age_adopt_log)
hist(adopt_vars_cort_df$age_adopt_sqrt)

```

## check for outliers in age adopted
there are no outliers after log-transformation! 
```{r}

# if the value is > 3 SD from the mean, then change value to 3SD from mean.
adopt_vars_cort_df <- adopt_vars_cort_df %>%
  mutate(age_adopt_log_outlier = ifelse(age_adopt_log > 3*sd(age_adopt_log, na.rm = T) 
                                        + mean(age_adopt_log, na.rm = T),1,0))

summary(adopt_vars_cort_df$age_adopt_log_outlier)
```

## linear models with all data 
### main effect of log age of adoption: SIG.
```{r}
logageadopt_linear_mod <- lmer(cort_use ~ 
                          age_adopt_log + corrected_cort_age_yrs.c + 
                           index_wave.c + index_day.c + batch.c + 
                    meds.c + sex.centered + #age_placed_log 
                     + (1 + index_wave.c | IDENT_SUBID), 
                      data = adopt_vars_cort_df)
summary(logageadopt_linear_mod)

save(logageadopt_linear_mod, file = "model_output/logageadopt_linear_model_results.Rdata")
```

### main effect of log age of adoption WITH age placed: SIG
```{r}
logageadopt_logageplaced_linear_mod <- lmer(cort_use ~ 
                          age_adopt_log + corrected_cort_age_yrs.c + 
                           index_wave.c + index_day.c + batch.c + 
                    meds.c + sex.centered + age_placed_log 
                     + (1 + index_wave.c | IDENT_SUBID), 
                      data = adopt_vars_cort_df)
summary(logageadopt_logageplaced_linear_mod)

save(logageadopt_logageplaced_linear_mod, file = "model_output/logageadopt_logageplaced_linear_model_results.Rdata")
```

### plot of these new significant effects...
```{r}

effect_df <- data.frame(effect("age_adopt_log", logageadopt_logageplaced_linear_mod))

effect_plot <- ggplot(data = effect_df, aes(x = age_adopt_log, y = fit)) + 
  geom_line() + geom_ribbon(aes(ymin = lower, ymax = upper), fill = dark_blue, alpha = 0.2) + 
  geom_point(data = adopt_vars_wide_by_wave, aes(x = age_adopt_log, y = cort_use), 
             color = dark_blue, alpha = 0.5, width = 0.05)  + 
  geom_line(data = adopt_vars_wide_by_wave, aes(x = age_adopt_log, y = cort_use, 
                  group = IDENT_SUBID), color = dark_blue, alpha = 0.2) +
  ylab("Waking cortisol") + xlab ("Age Adopted (log-transformed)")
effect_plot

ggsave(effect_plot, file = "figures/waking_cort_by_age_adopt_log.pdf", width = 6, height = 4)
```

### main effect of sqrt age of adoption: TREND.
```{r}
sqrtageadopt_linear_mod <- lmer(cort_use ~ age_adopt_sqrt + corrected_cort_age_yrs.c + 
                           index_wave.c + index_day.c + batch.c + 
                     sex.centered + age_placed_log + meds.c +
                     + (1 + index_wave.c | IDENT_SUBID), 
                      data = adopt_vars_cort_df)
summary(sqrtageadopt_linear_mod)

save(sqrtageadopt_linear_mod, file = "model_output/sqrtageadopt_linear_model_results.Rdata")
```

```{r}

effect_df <- data.frame(effect("age_adopt_sqrt", sqrtageadopt_linear_mod))

effect_plot <- ggplot(data = effect_df, aes(x = age_adopt_sqrt, y = fit)) + 
  geom_line() + geom_ribbon(aes(ymin = lower, ymax = upper), fill = dark_blue, alpha = 0.2) + 
  geom_jitter(data = adopt_vars_wide_by_wave, aes(x = age_adopt_sqrt, y = cort_use), 
             color = dark_blue, alpha = 0.5, width = 0.05)  + 
  geom_line(data = adopt_vars_wide_by_wave, aes(x = age_adopt_sqrt, y = cort_use, 
                  group = IDENT_SUBID), color = dark_blue, alpha = 0.2) +
  ylab("Waking cortisol") + xlab ("Age Adopted (sqrt-transformed)")
effect_plot

ggsave(effect_plot, file = "figures/waking_cort_by_age_adopt_sqrt.pdf", width = 6, height = 4)
```

### main effect of RAW age of adoption: TREND.
```{r}
ageadopt_linear_mod <- lmer(cort_use ~ IAI_7_AGEADOPT + corrected_cort_age_yrs.c + 
                           index_wave.c + index_day.c + batch.c + 
                    meds.c + sex.centered + #age_placed_log 
                     + (1 + index_wave.c | IDENT_SUBID), 
                      data = adopt_vars_cort_df)
summary(ageadopt_linear_mod)

save(ageadopt_linear_mod, file = "model_output/ageadopt_linear_model_results.Rdata")
```

### main effect of RAW age of adoption with RAW age placed: SIG.
```{r}
ageadopt_linear_mod <- lmer(cort_use ~ IAI_7_AGEADOPT + corrected_cort_age_yrs.c + 
                           index_wave.c + index_day.c + batch.c + 
                    meds.c + sex.centered +IAI_3_AGEORPH +
                     + (1 + index_wave.c | IDENT_SUBID), 
                      data = adopt_vars_cort_df)
summary(ageadopt_linear_mod)

save(ageadopt_linear_mod, file = "model_output/ageadopt_linear_model_results.Rdata")
```

```{r}

effect_df <- data.frame(effect("IAI_7_AGEADOPT", ageadopt_linear_mod))

effect_plot <- ggplot(data = effect_df, aes(x = IAI_7_AGEADOPT, y = fit)) + 
  geom_line() + geom_ribbon(aes(ymin = lower, ymax = upper), fill = dark_blue, alpha = 0.2) + 
  geom_point(data = adopt_vars_wide_by_wave, aes(x = IAI_7_AGEADOPT, y = cort_use), 
             color = dark_blue, alpha = 0.5)  + 
  geom_line(data = adopt_vars_wide_by_wave, aes(x = IAI_7_AGEADOPT, y = cort_use, 
                  group = IDENT_SUBID), color = dark_blue, alpha = 0.2) +
  ylab("Waking cortisol") + xlab ("Age Adopted (months)")
effect_plot

ggsave(effect_plot, file = "figures/waking_cort_by_age_adopt_raw.pdf", width = 6, height = 4)
```

### age adopt x age split (Age 13): NS
```{r}
adopt_vars_cort_df <- adopt_vars_cort_df %>%
  mutate(age_split = ifelse(corrected_cort_age_yrs < 13, "<13", ">=13"))

ageadopt_agesplit_mod <- lmer(cort_use ~ age_adopt_log*age_split +
                              #corrected_cort_age_yrs.c + 
                           index_wave.c + index_day.c + batch.c + 
                    meds.c + sex.centered + age_placed_log 
                     + (1 + index_wave.c | IDENT_SUBID), 
                      data = adopt_vars_cort_df)
summary(ageadopt_agesplit_mod)
anova(ageadopt_agesplit_mod, method = "III")

save(ageadopt_linear_mod, file = "model_output/ageadopt_agesplit_model_results.Rdata")
```

### Are piecewise effects moderated by age adopted?
This is hard to test... 

# 4) Testosterone
using sex-normed testosterone, generated in 2A cleaning script. 
```{r}
waking_cort_df$testos_normed <- waking_cort_df$testos_cleaned_log_wave_ave_gender_normed
waking_cort_df$testos_normed.c <- waking_cort_df$testos_normed- mean(waking_cort_df$testos_normed)
```

## descriptive info on testos:
```{r}

testos_df_by_subj <- waking_cort_df %>%
  group_by(IDENT_SUBID, index_wave) %>%
  dplyr::summarize(age = mean(corrected_cort_age_yrs), 
            sex = mean(DEM_3_GENDER_CHILD),
            testos_normed = mean(testos_normed, na.rm =T), 
            GROUP.n = mean(GROUP.n))


testos_demo <- testos_df_by_subj %>%
      mutate(GROUP = ifelse(GROUP.n == 0, "COMP", "PI")) %>%
    group_by(GROUP, index_wave) %>%
    dplyr::summarize(
              N_with_testos = sum(!is.na(testos_normed)),
              N_missing_testos = sum(is.na(testos_normed)),
              mean_age = mean(age[!is.na(testos_normed)]),
              sd_age = sd(age[!is.na(testos_normed)]),
              min_age = min(age[!is.na(testos_normed)]),
              max_age = max(age[!is.na(testos_normed)]),
              mean_testos = mean(testos_normed, na.rm = T),
              sd_testos = sd(testos_normed, na.rm = T),
              min_testos = min(testos_normed, na.rm = T),
              max_testos = max(testos_normed, na.rm = T))
testos_demo
save(testos_demo, file = "tables/testos_subsample_waking_cort_demo.Rdata")

```

## linear main effect, without age: SIG
```{r}
testos_linear_waking <- lmer(cort_use ~ 
                           GROUP + testos_normed + 
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c 
                     + (1 + index_wave.c | IDENT_SUBID), 
                      data = waking_cort_df)
summary(testos_linear_waking)
save(testos_linear_waking, file = "model_output/testos_waking_linear_model_results.Rdata")
```

## linear main effect, with age: BOTH SIG 
```{r}
testos_age_linear_waking <- lmer(cort_use ~ 
                           GROUP + testos_normed + corrected_cort_age_yrs.c +
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c 
                     + (1 + index_wave.c | IDENT_SUBID), 
                      data = waking_cort_df)
summary(testos_age_linear_waking)
save(testos_age_linear_waking, file = "model_output/testos_age_waking_linear_model_results.Rdata")
```

## testos x group without age: NS
```{r}
testosxgroup_linear_waking <- lmer(cort_use ~ 
                           GROUP * testos_normed + 
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c 
                     + (1 + index_wave.c | IDENT_SUBID), data = waking_cort_df)
summary(testosxgroup_linear_waking)
save(testosxgroup_linear_waking, file = "model_output/testosxgroup_waking_linear_model_results.Rdata")
```


## testos x group with age: NS
```{r}
testosxgroup_age_linear_waking <- lmer(cort_use ~ 
                           GROUP * testos_normed + corrected_cort_age_yrs.c +
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c 
                     + (1 + index_wave.c | IDENT_SUBID), data = waking_cort_df)
summary(testosxgroup_age_linear_waking)
save(testosxgroup_age_linear_waking, file = "model_output/testosxgroup_age_waking_linear_model_results.Rdata")
```


## age x group with tests: NS
so group x age effects are not significant when you control for testosterone...
```{r}
agexgroup_testos_linear_waking <- lmer(cort_use ~ 
                           GROUP * corrected_cort_age_yrs.c +testos_normed + 
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c 
                     + (1 + index_wave.c | IDENT_SUBID), data = waking_cort_df)
summary(agexgroup_testos_linear_waking)
save(agexgroup_testos_linear_waking, file = "model_output/agexgroup_testos_waking_linear_model_results.Rdata")
```


## 5) Power analysis of main model 

### observed power analysis
```{r}
# load model object 
load("../manuscript/model_output/waking_cort_piecewise_model.Rdata")
model <- data.frame(summary(lmer_piecewise_waking)$coefficients)
model$Coefficients <- rownames(model)

# run 'observed power" calculation for your model with niall's function
pow2 <- function (beta, se) {
  z1mb = abs(beta/se) - 1.96
  pow = (pnorm(abs(beta/se) - 1.96, mean = 0, sd = 1, lower.tail = TRUE))
  result <- list("power"= pow)
  return(result)
}


obs_pwr_groupxage1 <- pow2(beta = model$Estimate[model$Coefficients == "GROUPPI:age1"], se = model$Std..Error[model$Coefficients == "GROUPPI:age1"])

obs_pwr_groupxage2 <- pow2(beta = model$Estimate[model$Coefficients == "GROUPPI:age2"], se = model$Std..Error[model$Coefficients == "GROUPPI:age2"])

obs_pwr_groupxage1
obs_pwr_groupxage2

save(obs_pwr_groupxage1, obs_pwr_groupxage2, file = "model_output/obs_pwr_cort_waking_piecewise.rda")
```

### power analysis for future replication

Question: want to know how many subjects I need, or more within-subject time points, to reach 80% power for groupxage1 
```{r, include = F, eval = F}
# first generate fake data with expand.grid.
# let's say we have the same Number of subjects (240)
# but all with 3 data points at ages 4, 8, 12 for estimating groupage1
group <- c(0,1)
age1 <- c(4, 8, 12)
groupage1 <- age1*group
SUBID <- 1:240
#index_wave <- c(1,2,3) # don't need this, because repeated measures in age.

new_df <- expand.grid(groupage1 = groupage1, SUBID = SUBID)
nrow(new_df)

# Specify some fixed and random parameters --> from orig model. 
b <- c(model$Estimate[model$Coefficients == "(Intercept)"], model$Estimate[model$Coefficients == "GROUPPI:age1"]) # fixed intercept and slope
V1 <- 24.462 # random intercept variance
#V2 <- matrix(c(0.5,0.05,0.05,0.1), 2) # random intercept and slope variance-covariance matrix
#can't figure out how to get this.
s <- summary(lmer_piecewise_waking)$sigma  # residual standard deviation
#uild a model object
#Use the makeLmer or makeGlmer function to build an artificial lme4 object.

model1 <- makeLmer(y ~ groupage1 + (1 | SUBID), 
                   fixef=b, VarCorr=V1, sigma=s, data=new_df)

#Now we have “pilot” models, which can be used with simr.

powerSim(model1, nsim=50)

```