---
title: "longitudinal cort analysis"
author: "Michelle.VanTieghem"
date: "March 27, 2018"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
---


```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
```


# make specialized functions for this analysis used later.
```{r, include = F, echo = F}

b1 <- function(x, bp) ifelse(x < bp,x, bp)
b2 <- function(x, bp) ifelse(x < bp, bp, x) 

######### lmer Version #######
condslope.lmer.slopes <- function(x, z, c, y){
  # x = name of x variable
  # z = name of z variable (moderator)
  # c = conditional value of z
  # y = reg object
  # lmer model must be in this order x + z + x*z
# updated slightly by MVT on March 4, 2019 
# bcause tidy doesn't work. 
out <- summary(y)
xz <- paste(x, z, sep=":")
zx <- paste(z, x, sep = ":")
w0.intercept <- fixef(y)["(Intercept)"] + fixef(y)[z]*c
w1.slope <- fixef(y)[x] + fixef(y)[zx]*c 
#y.cond <- w0.intercept + w1.slope*xvalue
require(broom)
modtidy <- data.frame(summary(y)$coef)
modtidy$rownames <- rownames(modtidy)
coef2.var <- subset(modtidy, rownames == x)$Std..Error^2
coef4.var <- subset(modtidy, rownames == zx)$Std..Error^2
out.vcov <- vcov(y)
cond.se <- sqrt(coef2.var + (c) * (c) * coef4.var + 2 * (c) * out.vcov[x, zx])
t.val <- w1.slope/cond.se
p.val <- 2*(1-pt(abs(t.val), subset(modtidy, rownames == x)$df, lower.tail=T))
lower95 <- w1.slope-qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
upper95 <- w1.slope+qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
# z.out <- z.test(w1.slope, sigma.x= cond.se)
return(list(w0.intercept=round(w0.intercept, digits = 5), w1.slope=round(w1.slope, digits = 5),
            df = subset(modtidy, rownames == x)$df,
            t.val = round(t.val, digits = 5), p.val = round(p.val, digits = 4), 
            lower95 = round(lower95, digits = 5), 
            upper95 = round(upper95, digits = 5)))
}

condslope.lmer.pointestimates <- function(x, z, c, y){
  # x = name of x variable
  # z = name of z variable (moderator)
  # c = conditional value of z
  # y = reg object
  # lmer model must be in this order x + z + x*z
  
# updated slightly by MVT on March 4, 2019 
# bcause tidy doesn't work. 
out <- summary(y)
xz <- paste(x, z, sep=":")
w0.intercept <- fixef(y)["(Intercept)"] + fixef(y)[z]*c
w1.slope <- fixef(y)[x] + fixef(y)[xz]*c 
#y.cond <- w0.intercept + w1.slope*xvalue
require(broom)
modtidy <- data.frame(summary(y)$coef)
modtidy$rownames <- rownames(modtidy)
coef2.var <- subset(modtidy, rownames == x)$Std..Error^2
coef4.var <- subset(modtidy, rownames == xz)$Std..Error^2
out.vcov <- vcov(y)
cond.se <- sqrt(coef2.var + (c) * (c) * coef4.var + 2 * (c) * out.vcov[x, xz])
t.val <- w1.slope/cond.se
p.val <- 2*(1-pt(abs(t.val), subset(modtidy, rownames == x)$df, lower.tail=T))
lower95 <- w1.slope-qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
upper95 <- w1.slope+qt(0.975, subset(modtidy, rownames == x)$df)*cond.se
# z.out <- z.test(w1.slope, sigma.x= cond.se)
return(list(w0.intercept=round(w0.intercept, digits = 5), w1.slope=round(w1.slope, digits = 5),
            df = subset(modtidy, rownames == x)$df,
            t.val = round(t.val, digits = 5), p.val = round(p.val, digits = 5), 
            lower95 = round(lower95, digits = 5), 
            upper95 = round(upper95, digits = 5)))
}


```

# load data and organize
```{r}
load("../../../data/1_cort_data/Cortisol_Ready_to_analyze.Rdata")

range(cort_merge8$corrected_cort_age_yrs)

cort_merge8$time_of_day <- as.factor(with(cort_merge8,
                                   ifelse(time_of_day == "waking", "Waking", "Evening")))

summary(cort_merge8$time_of_day)

cort_merge8$index_day.c <- as.numeric(as.character(cort_merge8$index_day))
cort_merge8$meds.c <- as.numeric(as.character(cort_merge8$meds)) - mean(as.numeric(as.character(cort_merge8$meds)), na.rm = T)
sum(is.na(cort_merge8$batch))
sum(is.na(cort_merge8$cort_use))

# simplifying name of cort stuff 
cort_merge8$cort_use <- cort_merge8$cort_values_fixed_no_outliers 
data_use <- subset(cort_merge8, !is.na(cort_use) & 
                                                  !is.na(batch.c) & !is.na(meds.c) & !is.na(index_wave.c) & !is.na(index_day.c))
```

### make this extra summary dataset
this is so that we can plot the raw data, but plot means across both days instead of plotting 2 days per subj (per time of day, per wave...)
```{r}
cort_ave_days_m <- data_use %>%
  group_by(IDENT_SUBID, index_wave, time_of_day) %>%
  # this will provide average cort by each wave, and time of day...
  #aka average across 2 days if that subject has 2 days of data!
  dplyr::summarize(n_cort_samples = n(), 
            cort_use = mean(cort_use, na.rm = T),
            PDS_mean = mean(PDS_mean),
            corrected_cort_age_yrs = mean(corrected_cort_age_yrs), 
            GROUP.n = mean(GROUP.n), 
            sex.centered = mean(sex.centered), 
            meds.c = mean(meds.c),
            batch.c = mean(batch.c), 
            index_wave.c = mean(index_wave.c))

head(cort_ave_days_m)
cort_ave_days_m$GROUP <- as.factor(ifelse(cort_ave_days_m$GROUP.n == 0, "COMP", "PI"))
```


# waking + evening omnibus model 
## 1) Linear 

```{r}

lmer_linear <- lmer(cort_use ~ 
                           GROUP *time_of_day* corrected_cort_age_yrs  +
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c +
                      (1 + index_wave.c | IDENT_SUBID), 
                   data = data_use)
summary(lmer_linear)
save(lmer_linear, file= "model_output/full_cort_linear_model.Rdata")

```


## 2 ) Quadratic

make age squared variable
```{r, warnings = F}
data_use$age_squared <- data_use$corrected_cort_age_yrs ^2 
```


group x age_squared x time of day is significant!! 
```{r, warnings = F}
lmer_quad_nested <- lmer(cort_use ~ 
                           GROUP * time_of_day*age_squared + 
                           GROUP * time_of_day *corrected_cort_age_yrs +
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c +
                      (1 + index_wave.c | IDENT_SUBID), data = data_use)
summary(lmer_quad_nested)
cort_quadratic_model <- anova(lmer_quad_nested)
save(lmer_quad_nested, file = "model_output/full_cort_quadratic_model.Rdata")

```

## 3) Piecewise ORIGINAL VERSION

### Info:
Another approach would be to wrap the call to lmer in a function that is passed the breakpoint as a parameter, then minimize the deviance of the fitted model conditional upon the breakpoint using optimize. This maximizes the profile log likelihood for the breakpoint, and, in general (i.e., not just for this problem) if the function interior to the wrapper (lmer in this case) finds maximum likelihood estimates conditional upon the parameter passed to it, the whole procedure finds the joint maximum likelihood estimates for all the parameters.\

Reference: https://stats.stackexchange.com/questions/19772/estimating-the-break-point-in-a-broken-stick-piecewise-linear-model

### run optimizaton of breakpoint
notes: does work with time_of_day as a random slope, but does NOT work when using index_wave as a random slope.\ allowing age slopes to vary in both groups. \

```{r, warnings = F, message = F}

bp = min(data_use$corrected_cort_age_yrs) # starting point for breakpoint.

# Wrapper for Mixed effects model with variable break point
run_models <- function(bp){
  # this is the formula for your model 
  mod <- lmer(cort_use ~ 
                GROUP*b1(corrected_cort_age_yrs, bp)*time_of_day  +
                GROUP*b2(corrected_cort_age_yrs, bp)*time_of_day + batch.c +
         sex.centered + meds.c + index_wave.c + index_day.c +
                # do i have enough data for time of day to be included??!
                (1 + index_wave.c | IDENT_SUBID), data = data_use)
  deviance(mod, REML = F) # you're minimizing this, i think.
}

# set the range of breakpoints you'd allow. 
search.range <- c(min(data_use$corrected_cort_age_yrs)+
                    1,max(data_use$corrected_cort_age_yrs)-1)
# find the optimal model  based on breakpoint. 
mod.opt <- optimize(run_models, interval = search.range)
bp <- mod.opt$minimum # 

# get a confidence interval of your break point. 
CI.bp <- function(bp, tgt)
{
  run_models(bp) - tgt
}
# get info for confidence interval for break point! 
tgt <- mod.opt$objective + qchisq(0.95,1)
lb95 <- uniroot(CI.bp, lower=search.range[1], upper=bp, tgt=tgt)
ub95 <- uniroot(CI.bp, lower=bp, upper=search.range[2], tgt=tgt)
bp_lower <-lb95$root
bp_upper <- ub95$root
# save the result
optimize_piecewise_bp <- cbind(bp, bp_lower, bp_upper)
save(optimize_piecewise_bp, file = "model_output/cort_optimize_piecewise_breakpoint.Rdata")
```

optimal breakpoint is `r bp` with `bp_lower` and `bp_higher` 

### now rerun model with optimal age inflection
```{r}
# make sure that Waking is baseline level for time of day.
data_use$time_of_day <- factor(data_use$time_of_day, levels = c("Waking", "Evening"))
summary(data_use$time_of_day)
# changing this so that age1 and age2 are centered at breakpoint.
data_use$age1 <- b1(data_use$corrected_cort_age_yrs, bp) -bp 
data_use$age2 <- b2(data_use$corrected_cort_age_yrs, bp) -bp

piecewise_mod_waking_ref <- lmer(cort_use ~ 
                       GROUP*age1*time_of_day  +
                GROUP*age2*time_of_day + batch.c  + 
        sex.centered + meds.c + index_wave.c + index_day.c +
                (1 + index_wave.c | IDENT_SUBID), data = data_use)

# save main model ooutput
cort_piecewise_regular_model <-piecewise_mod_waking_ref
summary(cort_piecewise_regular_model)
save(cort_piecewise_regular_model, file = "model_output/full_cort_piecewise_regular_model.Rdata")

```


# WAKING ONLY 

```{r}
# subset data for only waking cortisol.
waking_cort_df <- subset(data_use, time_of_day == "Waking")

```

## 1) linear
```{r}

lmer_linear_waking <- lmer(cort_use ~ 
                           GROUP * corrected_cort_age_yrs +
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c 
                     + (1 + index_wave.c | IDENT_SUBID), 
                   data = waking_cort_df)
summary(lmer_linear_waking)
save(lmer_linear_waking, file= "model_output/waking_cort_linear_model.Rdata")
```



## 2) Quadratic 
```{r}
waking_cort_df$age_squared  <- waking_cort_df$corrected_cort_age_yrs^2
lmer_quadratic_waking <- lmer(cort_use ~ 
                           GROUP * age_squared + GROUP*corrected_cort_age_yrs + 
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c 
                     + (1 + index_wave.c | IDENT_SUBID), 
                   data = waking_cort_df)
summary(lmer_quadratic_waking)
save(lmer_quadratic_waking, file= "model_output/waking_cort_quadratic_model.Rdata")
```

## 3) Piecewise both groups vary 

### re-do breakpoint for confirmation! 
```{r, warnings = F}

bp = min(data_use$corrected_cort_age_yrs) # starting point for breakpoint.

# Wrapper for Mixed effects model with variable break point
run_models <- function(bp){
  # this is the formula for your model 
  mod <- lmer(cort_use ~ 
                GROUP*b1(corrected_cort_age_yrs, bp)  +
                GROUP*b2(corrected_cort_age_yrs, bp) + batch.c +
         sex.centered + meds.c + index_wave.c + index_day.c +
                (1 + index_wave.c | IDENT_SUBID), data = waking_cort_df)
  deviance(mod, REML = F) 
}

# set the range of breakpoints you'd allow. 
search.range <- c(min(waking_cort_df$corrected_cort_age_yrs)+
                    0.5,max(waking_cort_df$corrected_cort_age_yrs)-0.5)
# find the optimal model  based on breakpoint. 
mod.opt <- optimize(run_models, interval = search.range)
bp <- mod.opt$minimum 

# get a confidence interval of your break point. 
CI.bp <- function(bp, tgt)
{
  run_models(bp) - tgt
}
# get info for confidence interval for break point! 
tgt <- mod.opt$objective + qchisq(0.95,1)
lb95 <- uniroot(CI.bp, lower=search.range[1], upper=bp, tgt=tgt)
ub95 <- uniroot(CI.bp, lower=bp, upper=search.range[2], tgt=tgt)
bp_lower <-lb95$root
bp_upper <- ub95$root
# save output.
optimize_piecewise_bp <- cbind(bp, bp_lower, bp_upper)
save(optimize_piecewise_bp, file = "model_output/waking_optimize_piecewise_breakpoint.Rdata")
```

center ages on the breakpoint 
```{r}
# center these ages on the bp
# changing this so that age1 and age2 are centered at breakpoint.
waking_cort_df$age1 <- b1(waking_cort_df$corrected_cort_age_yrs, bp) -bp 
waking_cort_df$age2 <- b2(waking_cort_df$corrected_cort_age_yrs, bp) -bp
```

### model with optimal breakpoint.
```{r}
lmer_piecewise_waking <- lmer(cort_use ~ 
                                GROUP + age1 + age2 + 
                          GROUP:age1 + GROUP:age2 + batch.c + 
                          index_wave.c + index_day.c + 
                 sex.centered + meds.c +
                      (1 + index_wave.c | IDENT_SUBID), 
                   data = waking_cort_df)
#lmer_piecewise_waking
summary(lmer_piecewise_waking)
save(lmer_piecewise_waking, file= "model_output/waking_cort_piecewise_model.Rdata")

```

## Calculate ICC for mixed effects
```{r, warnings = F}
# from empty model
mod <- lmer(cort_use ~ (1 | IDENT_SUBID), data = waking_cort_df)
icc(mod)
ICC_cort <- icc(mod)[1]
save(ICC_cort, file = "model_output/ICC_for_waking_cort.Rdata")

# from full piecewise model
icc(lmer_piecewise_waking, adjusted = TRUE)
ICC_cort <- icc(lmer_piecewise_waking, adjusted=TRUE)[1]
save(ICC_cort, file = "model_output/ICC_for_waking_cort_full_piecewise_model.Rdata")

```

### prep for plotting
make new dataframe for plotting fitted effects
```{r}
# make specific new data for COMP age effect.
stuff <- data.frame(age1 = c(-9, -8, -7, -6.44, -5, -4, -3, -2, -1, -0.0001, rep(0,6)),
                    age2 = c(rep(0, 10), .0001, 2, 3, 4, 5, 5.9), 
                    GROUP = c(rep("COMP", 16), rep("PI",16)), batch.c = 0, 
       sex.centered = 0, meds.c = 0, index_wave.c = 0, index_day.c = 0)
nrow(stuff)

# specify design matrix with fixed effects and new df
fit.mat <- model.matrix(~  GROUP + age1 + age2 +
                          GROUP:age1 + GROUP:age2 + batch.c +
                          index_wave.c + index_day.c + 
                    sex.centered + meds.c, 
                    data = stuff) 
#head(fit.mat)
cis <- diag(fit.mat %*% tcrossprod(vcov(lmer_piecewise_waking), fit.mat))

# predict y values and lwr/upr bounds using model object and new data
# 
stuff$cort_Y <- predict(lmer_piecewise_waking, stuff, re.form = NA)
stuff$cort_lwr <- stuff$cort_Y-1.96*sqrt(cis)
stuff$cort_upr <- stuff$cort_Y+1.96*sqrt(cis)

# split into 2 dataframes and then stack. 
stuffage1 <- stuff %>%
  dplyr::select(-age2) %>%
  filter(age1 < 0) %>%
 rename(corrected_cort_age_yrs = age1) %>%
  mutate(corrected_cort_age_yrs = corrected_cort_age_yrs + bp)
stuffage2 <- stuff %>%
  dplyr::select(-age1) %>%
  filter(age2 > 0) %>%
  rename(corrected_cort_age_yrs = age2) %>%
  mutate(corrected_cort_age_yrs = corrected_cort_age_yrs + bp)

```

### make plot 
```{r}

waking_piecewise_PI_plot <- ggplot(stuffage1, aes(x = corrected_cort_age_yrs, 
                                                  y = cort_Y, color = GROUP)) + 
  geom_line(size = 1) + geom_line(data =stuffage2, aes(x = corrected_cort_age_yrs, 
                                                       y = cort_Y, color = GROUP), size = 1) + 
  my_colors + my_colors2 +
  ylab("Waking Cortisol (nmol/L)") + xlab("Age (years)") + 
  geom_point(data = subset(cort_ave_days_m, time_of_day == "Waking"),
             aes(x= corrected_cort_age_yrs, y= cort_use, 
                                         color = GROUP), size=1, alpha = 0.2)+ 
  geom_line(data = subset(cort_ave_days_m, time_of_day == "Waking"),
            aes(x = corrected_cort_age_yrs, y = cort_use, 
                color = GROUP, group = IDENT_SUBID), alpha = 0.2) + 
 scale_x_continuous(breaks = seq(4, 19, by = 2), limits=c(4.0, 19), expand = c(.02,0)) +
 scale_y_continuous(expand = expand_scale(mult = c(0, .1))) +
   theme(legend.position= c(0.10, 0.90),  legend.key.height = unit(0.05, "cm"),
         legend.key.width = unit(0.05, "cm"), legend.background = element_rect(color = "black")) + 
  geom_ribbon(data = stuffage1, aes(ymin = cort_lwr, ymax =cort_upr, fill = GROUP),  alpha = .1, color = NA)  + 
  geom_ribbon(data = stuffage2, aes(ymin = cort_lwr, ymax = cort_upr, fill = GROUP), alpha = 0.1, color = NA)

pdf("figures/waking_piecewise_model_plot_overlapped_CI.pdf", width = 6, height = 4)
waking_piecewise_PI_plot
dev.off()

# save to powerpoint
ppt_test2 <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with_vg(code = print(waking_piecewise_PI_plot)) %>%
  print(target = "../../../manuscript/figures_for_PNAS/powerpoint_vector_figures/orig_from_R/waking_piecewise_model_plot_overlapped_CI.pptx")
  
```


### prep for signficance plot of group differences across age
```{r}
bp = 13
age.search.range <- seq(from = min(data_use$corrected_cort_age_yrs)- bp, to =  max(data_use$corrected_cort_age_yrs) - bp, by = 0.25)

waking_SS_table <- data.frame() 
for (age in age.search.range){
  if(age < 0){ # 0 = breakpoint of age 13
    ss_test <- data.frame(Age = age + bp, condslope.lmer.pointestimates("GROUPPI", "age1", age, lmer_piecewise_waking))
   # print(ss_test)
  } else {
    ss_test <- data.frame(Age = age + bp, condslope.lmer.pointestimates("GROUPPI", "age2", age, lmer_piecewise_waking))
  }
  waking_SS_table <- rbind(waking_SS_table, ss_test)
}
#waking_SS_table
save(waking_SS_table, file = "model_output/waking_piecewise_orig_posthoc_group_diff.Rdata")

# age cut-offs
PI_greater_than_comp <- waking_SS_table %>%
  filter(p.val < 0.05 & w1.slope > 0) %>%
  summarize(min_age = min(Age), 
            max_age = max(Age)) %>%
  mutate(comparison = "PI > COMP")

PI_smaller_than_comp <- waking_SS_table %>%
  filter(p.val < 0.05 & w1.slope < 0) %>%
  summarize(min_age = min(Age), 
            max_age = max(Age)) %>%
  mutate(comparison = "PI < COMP")

waking_SS_agerange_table <- rbind(PI_greater_than_comp, PI_smaller_than_comp)
waking_SS_agerange_table
 save(waking_SS_agerange_table, file = "model_output/wakingpiecewise_orig_posthoc_ageranges_of_group_diff.Rdata")

```

### make plot
```{r}
waking_SS_plot <- ggplot(waking_SS_table, aes(x = Age, y = w1.slope)) + 
  geom_line() + geom_ribbon(aes(ymin = lower95, ymax = upper95), alpha = 0.2) + 
  geom_hline(aes (yintercept = 0), linetype = "dashed") + ylab("Group effect for waking cortisol (PI - COMP)") +
 scale_x_continuous(breaks = seq(4,19, by = 2), limits=c(4.0, 19), expand = c(.02,0)) +
 scale_y_continuous(expand = expand_scale(mult = c(0, .1))) 

pdf("figures/waking_piecewise_model_SS_significance_plot.pdf", width =6, height = 4)
waking_SS_plot
dev.off()

# save to powerpoint
ppt_test2 <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with_vg(code = print(waking_SS_plot)) %>%
  print(target = "../../../manuscript/figures_for_PNAS/powerpoint_vector_figures/orig_from_R/waking_piecewise_model_SS_significance_plot.pptx")
  
```

## option 1: combining interaction plot and significance plot with entire age range
```{r}
library(gridExtra)
waking_piecewise_PI_plot2 <- waking_piecewise_PI_plot + ylim(0, 50)
waking_SS_plot2 <- waking_SS_plot + ylim(-20, 20)
grid.arrange(waking_piecewise_PI_plot2, waking_SS_plot2)

#SAVING TO POWERPT 
ppt_test2 <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with_vg(code = print(grid.arrange(waking_piecewise_PI_plot2, waking_SS_plot2))) %>%
  print(target = "../../../manuscript/figures_for_PNAS/powerpoint_vector_figures/orig_from_R/stacked_waking_piecewise_SS_plots.pptx")
  
```

### post-hoc calculations for piecewise model 
```{r}
# calculate slopes for age effects, per group, before and after breakpoint.
comp_slope_age1 <- data.frame(group = "COMP", condslope.lmer.slopes( "age1","GROUPPI", 0, lmer_piecewise_waking))
PI_slope_age1 <- data.frame(group = "PI", condslope.lmer.slopes("age1", "GROUPPI", 1, lmer_piecewise_waking))
comp_slope_age2 <- data.frame(group = "COMP", condslope.lmer.slopes( "age2","GROUPPI", 0, lmer_piecewise_waking))
PI_slope_age2 <- data.frame(group = "PI", condslope.lmer.slopes("age2", "GROUPPI", 1, lmer_piecewise_waking))

waking_posthoc_slopes <- rbind(comp_slope_age1, PI_slope_age1, comp_slope_age2, PI_slope_age2)

names(waking_posthoc_slopes) <- c("GROUP", "Breakpoint Intercept", "Estimated Slope", "df", "t.value", "p.value", "lower", "upper")
waking_posthoc_slopes$'Age range' <- c("age1", "age1", "age2", "age2")
waking_posthoc_slopes
# save table 
save(waking_posthoc_slopes, file = "model_output/waking_piecewise_orig_posthoc_age_slopes.Rdata")
```


# EVENING ONLY 
```{r}
# subset data to only include evening cort
evening_cort_df <-  subset(data_use, time_of_day == "Evening")

```

## 1) Linear
```{r}
# center these ages on the bp
lmer_linear_evening <- lmer(cort_use ~ 
                                GROUP + GROUP*corrected_cort_age_yrs + 
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c +
                     + (1 + index_wave.c | IDENT_SUBID), 
                   data = evening_cort_df)
#lmer_piecewise_waking
summary(lmer_linear_evening)
save(lmer_linear_evening, file= "model_output/evening_cort_linear_model.Rdata")
```

## 2) Quadratic
```{r}
# center these ages on the bp

lmer_quadratic_evening <- lmer(cort_use ~ 
                               GROUP*age_squared + GROUP*corrected_cort_age_yrs +
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c +
                     + (1 | IDENT_SUBID), 
                   data = evening_cort_df)
#lmer_piecewise_waking
summary(lmer_quadratic_evening)
save(lmer_quadratic_evening, file= "model_output/evening_cort_quadratic_model.Rdata")
```

## 3) Piecewise both groups vary 
note: using same breakpoint as waking models.
```{r}
# center these ages on the bp
lmer_piecewise_evening <- lmer(cort_use ~ 
                                GROUP + age1 + age2 + 
                          GROUP:age1 + GROUP:age2 +
                           index_wave.c + index_day.c + batch.c + 
                    sex.centered + meds.c +
                     + (1 | IDENT_SUBID), 
                   data = evening_cort_df)
#lmer_piecewise_waking
summary(lmer_piecewise_evening)
save(lmer_piecewise_evening, file= "model_output/evening_cort_piecewise_model.Rdata")
```


# Comparing models

## omnibus models (evening and waking)
```{r}
anova(lmer_linear, lmer_quad_nested) # big difference from linear to quad
anova(lmer_quad_nested, piecewise_mod_waking_ref) # big difference from quad to piecewise

AIC_table <- AIC(lmer_linear, lmer_quad_nested, piecewise_mod_waking_ref)
BIC_table <- BIC(lmer_linear, lmer_quad_nested, piecewise_mod_waking_ref)
model_comparison_cort <- data.frame(cbind(AIC_table, BIC_table$BIC))
names(model_comparison_cort) <- c("DF", "AIC", "BIC")
rownames(model_comparison_cort) <- c("Linear", "Quadratic", "Piecewise")
model_comparison_cort$model <- rownames(model_comparison_cort)

#save 
model_comparison_cort
save(model_comparison_cort, file = "model_output/model_comparisons_cort.Rdata")


```


## Waking models only 
```{r}
anova(lmer_linear_waking, lmer_quadratic_waking) #, big difference from linear to quadratic
anova(lmer_quadratic_waking, lmer_piecewise_waking) # big difference from quadratic to piecewise

# to simplify my life, just going to report the 3 models for now 
AIC_table <- AIC(lmer_linear_waking, lmer_quadratic_waking, lmer_piecewise_waking)
BIC_table <- BIC(lmer_linear_waking, lmer_quadratic_waking, lmer_piecewise_waking)
model_comparison_waking <- data.frame(cbind(AIC_table, BIC_table$BIC))
names(model_comparison_waking) <- c("DF", "AIC", "BIC")
rownames(model_comparison_waking) <- c("Linear", "Quadratic", "Piecewise")
model_comparison_waking$model <- rownames(model_comparison_waking)
save(model_comparison_waking, file = "model_output/model_comparisons_waking.Rdata")

model_comparison_waking
```


## evening models only 
```{r}
anova(lmer_linear_evening, lmer_quadratic_evening) #, no difference from linear to quadratic
anova(lmer_quadratic_evening, lmer_piecewise_evening) # big difference from quadratic to piecewise

# to simplify my life, just going to report the 3 models for now 
AIC_table <- AIC(lmer_linear_evening, lmer_quadratic_evening, lmer_piecewise_evening)
BIC_table <- BIC(lmer_linear_evening, lmer_quadratic_evening, lmer_piecewise_evening)
model_comparison_evening <- data.frame(cbind(AIC_table, BIC_table$BIC))
names(model_comparison_evening) <- c("DF", "AIC", "BIC")
rownames(model_comparison_evening) <- c("Linear", "Quadratic", "Piecewise")
model_comparison_evening$model <- rownames(model_comparison_evening)
save(model_comparison_evening, file = "model_output/model_comparisons_evening.Rdata")

model_comparison_evening
```
