---
title: 'Detecting Outliers in updated cortisol data'
author: "Michelle VanTieghem"
date: "May 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# NOTES 
## calculate outliers for longitudinal cortisol  \
We want to calculate outliers by Group, Wave, day & time of day \
in other words, subset the data by index20 (each time point per sub)  \
and then split into PI and comparisons \
based on those groups (40) calculate outliers 3SD beyond mean. \

## updated May 2018 
to account for the fact that we will be removing the youngest kids, and oldest kids, to balance groups! \
remove outliers AFTER that fact. \

## updated june 2018:
make sure you're using corrected_cort_age_yrs and cort_values_fixed  \


```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
```


# load cleaned data with correct ages and times, which includes all data entry master data.
```{r}
# load data, already reformatted and with *corrected* times!!
load(file= "../../../data/1_cort_data/0_Long_cort_corrected_ages_and_times.Rdata")

nrow(cort_cleaned)
```

# remove COMPS < age 4 before calculating outliers
```{r}

cort_cleaned2 <- subset(cort_cleaned, !(IDENT_SUBTYPE == 0 & corrected_cort_age_yrs < 4))

range(cort_cleaned2$corrected_cort_age_yrs, na.rm = T)
nrow(cort_cleaned2) #
```

# get rid of extreme / impalusible values first 
```{r}

plot(cort_cleaned2$cort_values_fixed)

cort_cleaned2 <- subset(cort_cleaned2, cort_values_fixed < 200)
nrow(cort_cleaned2)
```


# calculate outliers by Group, wave, time of day

```{r}
# group the data by GROUP (PI or comp) and CORT INDEX20 (e.g. cort 1-20)
# note: index_20 separates each cortisol data collection time point 
# cort 1-4 = wave 1, day 1, 4 samples per day
# cort 5-8 = wave 1, day 2, 4 samples per day
# cort 9-13 = wave2, day 1, 4 samples per day
# cort 14-16 = wave2, day 2, 4 samples per day
# cort 17-20 = wave 3, day 1 (no day 2 at wave 3), 4 samples per day

outlier_groups <- group_by(cort_cleaned2, IDENT_SUBTYPE,index20)

# get the means, SD, upper and lower limit (3 SD above or below mean) for *each* grouping of cort daya
outlier_info <- dplyr::summarize(outlier_groups, mean= mean(cort_values_fixed, na.rm = TRUE), sd= sd(cort_values_fixed, na.rm = TRUE), upper_limit= mean(cort_values_fixed, na.rm =TRUE) + 3*sd(cort_values_fixed, na.rm = TRUE), lower_limit=mean(cort_values_fixed, na.rm = TRUE) - 3*sd(cort_values_fixed, na.rm = TRUE))

# make an empty dataframe to store new stuff in 
cort_cleaned2.mark.outliers <- data.frame()

# make an empty column in outlier_info to store N outliers 
outlier_info$N_outliers <- NA

# for each grouping of cortisol data in outlier_info 
for (i in 1:nrow(outlier_info)){
  print(i)
  # which group is this? PI or comp
  group <- outlier_info$IDENT_SUBTYPE[i]
  # which cort index? 1-20
  index <- outlier_info$index20[i]
  print(index)
  # set the max and minimum values of cortisol that we will use as outlier threshold
  lower_limits <- outlier_info$lower_limit[i]
  upper_limits <- outlier_info$upper_limit[i]
  # get all of the original data by group & cort index  
  data_chunk <- subset(cort_cleaned2, index20 == index & IDENT_SUBTYPE == group)
  # mark the outliers as True 
  data_chunk$cort_mark_outliers <- ifelse(data_chunk$cort_values_fixed >= upper_limits
                          | data_chunk$cort_values_fixed <= lower_limits, T, F)
  # replace outlier values with NAs in a new variable 
  data_chunk$cort_values_fixed_no_outliers <- ifelse(data_chunk$cort_mark_outliers == T, NA, data_chunk$cort_values_fixed)
  print(nrow(data_chunk))
  # count how many NAs there are - how many outliers there are.
  N_outliers <- sum(data_chunk$cort_mark_outliers, na.rm = TRUE)
  # save the N in your outlier table at the correct index 
  outlier_info$N_outliers[i] <- N_outliers 
  # save this into a new dataframe 
  cort_cleaned2.mark.outliers <- rbind(cort_cleaned2.mark.outliers, data_chunk) 
} 

#outlier_info

#head(cort_cleaned2.mark.outliers)
```

## save this table for later! 
```{r, evaluate = F}
# save the outlier table!! 
write.table(outlier_info, "../../../data/1_cort_data/Marta_version_Cortisol_outlier_info_table.txt", col.names = T, quote= F)

```


# Remove the outliers from your cleaned data and save
```{r, evaluate = F}
nrow(cort_cleaned2)
# first, we have to subset out the outliers (right now they are just marked)
cort_cleaned.no.outliers <- subset(cort_cleaned2.mark.outliers, !is.na(cort_values_fixed_no_outliers))

save(cort_cleaned.no.outliers, file = "../../../data/1_cort_data/1A_Long_cort_removed_outliers.Rdata")


```