---
title: "Calculating age intervals for longitudinal cortisol"
author: "Michelle VanTieghem"
date: "January 23, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# updated june 2018:
make sure you're using corrected_cort_age_yrs and cort_values_fixed 



```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
```


# Load the data 
```{r, include = F}
load(file= "../../../data/1_cort_data/1A_Long_cort_removed_outliers.Rdata")

nrow(cort_cleaned.no.outliers)
names(cort_cleaned.no.outliers)
```

# set up all of these variables o be in the right format! 
```{r}
cort_cleaned.no.outliers$GROUP <- as.factor(cort_cleaned.no.outliers$GROUP.x)
cort_cleaned.no.outliers$index_wave <- as.factor(cort_cleaned.no.outliers$index_wave)
# make this effect coded! 
cort_cleaned.no.outliers$index_day <- as.factor(ifelse(cort_cleaned.no.outliers$index_day == 1, -1, 1))

      
# make this a factor! 
cort_cleaned.no.outliers$time_of_day <- as.factor(ifelse(cort_cleaned.no.outliers$index4 == 1 | cort_cleaned.no.outliers$index4 == 2, "waking","evening"))

cort_cleaned.no.outliers$gender.centered <- cort_cleaned.no.outliers$DEM_3_GENDER_CHILD - mean(cort_cleaned.no.outliers$DEM_3_GENDER_CHILD, na.rm = TRUE)
```

## making an index2 variable, which will average over the morning samples and evening samples to give a single slope for each subject. 
```{r}
cort_cleaned.no.outliers$index3.f <- as.factor(ifelse(cort_cleaned.no.outliers$index4 == 1 | cort_cleaned.no.outliers$index4 == 2, "waking", ifelse(cort_cleaned.no.outliers$index4 == 3, "afternoon", "evening")))

cort_cleaned.no.outliers$index2.f <- as.factor(ifelse(cort_cleaned.no.outliers$index4 == 1 | cort_cleaned.no.outliers$index4 == 2, "waking",  "evening"))

```

## check ages and fill in NAs with session age. 
```{r}
# not all ages are fixed, because some just need to be session. 
sum(is.na(cort_cleaned.no.outliers$corrected_cort_age_yrs))

cort_cleaned.no.outliers$corrected_cort_age_yrs <- with(cort_cleaned.no.outliers, ifelse(
  is.na(corrected_cort_age_yrs), corrected_session_age_yrs, corrected_cort_age_yrs))
# check they are all fixed! 
sum(is.na(cort_cleaned.no.outliers$corrected_cort_age_yrs))

hist(cort_cleaned.no.outliers$corrected_cort_age_yrs)

cort_cleaned.no.outliers$age.centered <- cort_cleaned.no.outliers$corrected_cort_age_yrs - mean(cort_cleaned.no.outliers$corrected_cort_age_yrs, na.rm = TRUE)

range(cort_cleaned.no.outliers$corrected_cort_age_yrs)
```



## make a variable calculating age-intervals between waves 
super slow function, but it works!
```{r}
cort_cleaned.no.outliers$cort_age_wave1 <- NA 
cort_cleaned.no.outliers$cort_age_wave2 <- NA
cort_cleaned.no.outliers$cort_age_wave3 <- NA
for (i in 1:nrow(cort_cleaned.no.outliers)){
  n <- cort_cleaned.no.outliers$IDENT_SUBID[i]
  subset <- subset(cort_cleaned.no.outliers, IDENT_SUBID == n)
  if (1 %in% subset$index_wave){
    cort_cleaned.no.outliers$cort_age_wave1[cort_cleaned.no.outliers$IDENT_SUBID == n] <- subset$corrected_cort_age_yrs[subset$index_wave == 1][1]
  }
  if (2 %in% subset$index_wave){
    cort_cleaned.no.outliers$cort_age_wave2[cort_cleaned.no.outliers$IDENT_SUBID == n]<- subset$corrected_cort_age_yrs[subset$index_wave == 2][1]
  } 
  if (3 %in% subset$index_wave){
    cort_cleaned.no.outliers$cort_age_wave3[cort_cleaned.no.outliers$IDENT_SUBID == n] <- subset$corrected_cort_age_yrs[subset$index_wave == 3][1]
  } 
}

tail(cort_cleaned.no.outliers$cort_age_wave3) 
```

## now calculate age intervals
```{r}
cort_cleaned.no.outliers$age_intervals <- ifelse(cort_cleaned.no.outliers$index_wave == 3 & !is.na(cort_cleaned.no.outliers$cort_age_wave1), (cort_cleaned.no.outliers$cort_age_wave3 - cort_cleaned.no.outliers$cort_age_wave1), 
                           ifelse(cort_cleaned.no.outliers$index_wave == 3 & !is.na(cort_cleaned.no.outliers$cort_age_wave2), (cort_cleaned.no.outliers$cort_age_wave3 - cort_cleaned.no.outliers$cort_age_wave2),
                             ifelse(cort_cleaned.no.outliers$index_wave == 3 & is.na(cort_cleaned.no.outliers$cort_age_wave2) & is.na(cort_cleaned.no.outliers$cort_age_wave1), 0,
                            ifelse(cort_cleaned.no.outliers$index_wave == 2 & !is.na (cort_cleaned.no.outliers$cort_age_wave1), 
                       (cort_cleaned.no.outliers$cort_age_wave2 - cort_cleaned.no.outliers$cort_age_wave1),
                       ifelse(cort_cleaned.no.outliers$index_wave == 2 & is.na (cort_cleaned.no.outliers$cort_age_wave1), 
                       0,  0)))))

summary(cort_cleaned.no.outliers$age_intervals)

```

## save the new age_intervals variable to the dataframe object for future!
```{r}

save(cort_cleaned.no.outliers, file= "../../../data/1_cort_data/1B_Long_cort_added_age_intervals.Rdata")
nrow(cort_cleaned.no.outliers)

```
