---
title: "Cleaning testosterone"
author: "Michelle.VanTieghem"
date: "June 11, 2018"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# clean and organize data for for longitudinal TESTOSTERONE
- based on sex - within sex normed percentages? 
- significant effects of time of day (waking > evening)
- a lot of evening testos data missing from wave 1 
- so we will use only waking testos values! 

```{r, include = F}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lme4)
library(lmerTest)
main_dir <- "/Users/michellevantieghem2/Documents/Columbia/LabDATA/Projects/git_repo_backups/structural_analysis_2018/"

```

# load data that has cort + testosterone, 
with correct ages and times, which includes all data entry master data.
and has already removed cortisol outliers. 
so we are just going to find testosterone for these samples. 
```{r}
# load data, already reformatted and with *corrected* times!!
load(file=  "../../../data/1_cort_data/1C_long_cort_added_puberty_and_meds.Rdata")
nrow(cort_cleaned.no.outliers)

cort_cleaned.no.outliers$gender.f <- with(cort_cleaned.no.outliers, as.factor(ifelse(DEM_3_GENDER_CHILD == 0, "male", "female")))

range(cort_cleaned.no.outliers$corrected_cort_age_yrs)

```


```{r}

names(cort_cleaned.no.outliers)[grepl("test", names(cort_cleaned.no.outliers))]
summary(cort_cleaned.no.outliers$testosterone_notes)
summary(cort_cleaned.no.outliers$testosterone)
range(cort_cleaned.no.outliers$testosterone, na.rm = T)

```

# Cleaning 
## get rid of extreme / impalusible values first 
but don't remove entire ROW, just fill in NA for testosterone
because we want to keep the good cort values! 
how do I know what plausible values are supposed to be? 
```{r}
hist(cort_cleaned.no.outliers$testosterone)
cort_cleaned.no.outliers$testos_cleaned <- ifelse(cort_cleaned.no.outliers$testosterone > 500, NA, 
                                                 as.numeric(cort_cleaned.no.outliers$testosterone))

hist(cort_cleaned.no.outliers$testos_cleaned)
```

## log transformation
from dominics paper. 

```{r}
# how many zeros? 
zeros <- sum(ifelse(cort_cleaned.no.outliers$testos_cleaned == 0, 1, 0), na.rm  = T)
zeros # these two values will be come infinity, change to 0.001
cort_cleaned.no.outliers$testos_cleaned2 <- ifelse(cort_cleaned.no.outliers$testos_cleaned == 0, 0.1, 
                                                 cort_cleaned.no.outliers$testos_cleaned)


cort_cleaned.no.outliers$testos_cleaned_log <- log(cort_cleaned.no.outliers$testos_cleaned2)
hist(cort_cleaned.no.outliers$testos_cleaned_log)
summary(cort_cleaned.no.outliers$testos_cleaned_log)
```

## check what time of day most testos collected
data was collected across all times of day... 
```{r}
check_td <- cort_cleaned.no.outliers %>% 
  mutate(index4.f = ifelse(index4 == 1, "waking", ifelse(index4 == 2, "waking+45", ifelse(index4 == 3, "5pm", "8pm")))) %>%
  group_by( index_wave, index_day, index4, index4.f) %>% #, index_day, index_wave) %>%
  summarize(N_samples = n(),
            N_testos = sum(!is.na(testos_cleaned_log)),
            N_missing_testos = sum(is.na(testos_cleaned_log)))
check_td

```


## graph raw data 
```{r}

ggplot(cort_cleaned.no.outliers, aes(x = index4, y = testos_cleaned_log)) + 
  geom_smooth() + theme_classic() + geom_point (alpha = 0.5) + facet_grid (gender.f ~ index_wave)
```


## test whether there are timeof day effects on testosterone
```{r}
testos_timeofday <- lmer(testos_cleaned_log ~ batch + index_day + time_of_day + index_wave + GROUP + corrected_cort_age_yrs + (1 | IDENT_SUBID), 
     data = cort_cleaned.no.outliers)
summary(testos_timeofday)
save(testos_timeofday, file = "../supplemental/tables/testos_time_of_day_model.Rdata")
```

## waking > evening for testosterone.
```{r}
library(effects)
plot(effect("time_of_day", testos_timeofday))
```


## average testos measures: by wave across all times of day
```{r}
# calculate the mean testosterone for each subject on each day for each wave. 
testos_wave <- cort_cleaned.no.outliers %>% 
  filter(time_of_day == "waking") %>%
  group_by(IDENT_SUBID, index_wave) %>%
  summarize(testos_cleaned_log_wave_ave = mean(testos_cleaned_log, na.rm = T))

hist(testos_wave$testos_cleaned_log_wave_ave)

# merge back into long dataset. for future use. 
cort_cleaned.no.outliers <- left_join(cort_cleaned.no.outliers, testos_wave, by = c("IDENT_SUBID", "index_wave"))

nrow(cort_cleaned.no.outliers)
```



## make a df for testsos specifically 
note: only 10 PI males with testos at wave 3! be careful with 3 way ints.
```{r}
# can't be duplicated across all of these times of day and days 
# just one row per wave per subj.

# note: index day is atually already mean-centered so we need -1! 
testos_df <- subset(cort_cleaned.no.outliers, index4 == 1 & index_day == -1)
nrow(testos_df)
summary(testos_df$index_wave)

table <- testos_df %>% 
  group_by(GROUP, gender.f, index_wave) %>%
  summarize(n = n())
table
```


## graph cleaned testos by group and sex.
```{r}
testos_df$index_wave.n <- as.numeric(as.character(testos_df$index_wave))
ggplot(testos_df, aes(x =index_wave.n, y = testos_cleaned_log_wave_ave, color = GROUP)) + 
  theme_classic()+ geom_point(alpha = 0.5) + geom_smooth() + facet_grid (GROUP ~ gender.f)  


```

## graph cleaned testos by batch.
```{r}
ggplot(testos_df, aes(x = as.factor(batch), y = testos_cleaned_log_wave_ave)) + 
  geom_violin () + theme_bw() + geom_jitter (alpha = 0.5)


```

# calculate outliers by gender & group... there are no outliers!
```{r}
# group the data by GENDER, and wave (not day or time of day)
# using the transformmed day average varaiable for testosterone. 
outlier_groups <- group_by(testos_df, GROUP, gender.f)

# get the means, SD, upper and lower limit (3 SD above or below mean) for *each* grouping of cort daya
outlier_info <- summarize(outlier_groups, mean= mean(testos_cleaned_log_wave_ave, na.rm = TRUE), sd= sd(testos_cleaned_log_wave_ave, na.rm = TRUE), upper_limit= mean(testos_cleaned_log_wave_ave, na.rm =TRUE) + 3*sd(testos_cleaned_log_wave_ave, na.rm = TRUE), lower_limit=mean(testos_cleaned_log_wave_ave, na.rm = TRUE) - 3*sd(testos_cleaned_log_wave_ave, na.rm = TRUE))

# make an empty dataframe to store new stuff in 
testos.outliers <- data.frame()

# make an empty column in outlier_info to store N outliers 
outlier_info$N_outliers <- NA

# for each grouping of cortisol data in outlier_info 
for (i in 1:nrow(outlier_info)){
  group <- outlier_info$GROUP[i]
  gender<- outlier_info$gender.f[i]
  # set the max and minimum values of cortisol that we will use as outlier threshold
  lower_limits <- outlier_info$lower_limit[i]
  upper_limits <- outlier_info$upper_limit[i]
  # get all of the original data by group & cort index  
  data_chunk <- subset(testos_df, gender.f == gender & GROUP == group)
  # mark the outliers as True 
  data_chunk$testos_mark_outliers <- ifelse(data_chunk$testos_cleaned_log_wave_ave >= upper_limits
                          | data_chunk$testos_cleaned_log_wave_ave <= lower_limits, T, F)
  # replace outlier values with NAs in a new variable 
  data_chunk$testos_cleaned_log_wave_ave <- ifelse(data_chunk$testos_cleaned_log_wave_ave == T, NA, data_chunk$testos_cleaned_log_wave_ave)
 # print(nrow(data_chunk))
  # count how many NAs there are - how many outliers there are.
  N_outliers <- sum(data_chunk$testos_mark_outliers, na.rm = TRUE)
  # save the N in your outlier table at the correct index 
  outlier_info$N_outliers[i] <- N_outliers 
  # save this into a new dataframe 
  testos.outliers <- rbind(testos.outliers, data_chunk) 
} 

outlier_info
# ok, good identical.
nrow(testos.outliers)
nrow(testos_df)


```

# check models testos 
## test for differences by sex

```{r}

testos_gender <- lmer(testos_cleaned_log_wave_ave ~ gender.f + GROUP + corrected_cort_age_yrs + index_wave + batch + 
                      + (1 | IDENT_SUBID), data = testos_df)

anova(testos_gender)

```


## age x gender effect... and group effect.
```{r}
testos_gender_int <- lmer(testos_cleaned_log_wave_ave~ GROUP + gender.f*corrected_cort_age_yrs + index_wave + batch  + 
                       (1 | IDENT_SUBID), data = testos_df)

anova(testos_gender_int)
```


## no group x sex effect
```{r}
testos_gender_int <- lmer(testos_cleaned_log_wave_ave~ GROUP*gender.f + gender.f*corrected_cort_age_yrs + index_wave + batch  + 
                       (1 | IDENT_SUBID), data = testos_df)

anova(testos_gender_int)
```

### plot age x sex effects on testos (expected in literature)
```{r}
ggplot(testos_df, aes(x =corrected_cort_age_yrs, y = testos_cleaned_log_wave_ave, color = gender.f)) + 
  geom_smooth(method = "lm") + theme_bw() + geom_jitter (alpha = 0.5) #+ facet_grid (~ gender.f)

```


## thought there might be group x age x gender effect
there is not - good, not supposed to be! 
```{r}
testos_gender_int <- lmer(testos_cleaned_log_wave_ave ~  GROUP * gender.f* corrected_cort_age_yrs + index_wave 
                      + (1 | IDENT_SUBID), data = testos_df)

anova(testos_gender_int)
```

### plot this effect
confirming that age-related increases are the same across PIS and COMPS
```{r}
ggplot(testos_df, aes(x =corrected_cort_age_yrs, y = testos_cleaned_log_wave_ave, color = GROUP)) + 
  geom_smooth(method = "lm") + theme_bw() + geom_jitter (alpha = 0.5) + facet_grid (~ gender.f)

```



# rank order within sex - (across both groups)

## first get min and max for each sex 
```{r}
testos_min_max <- testos_df %>%
  group_by(DEM_3_GENDER_CHILD) %>% 
  filter(!is.na(testos_cleaned_log_wave_ave)) %>% 
  summarize(min=min(testos_cleaned_log_wave_ave), 
            max= max(testos_cleaned_log_wave_ave), 
            mean=mean(testos_cleaned_log_wave_ave))

testos_min_max

```

## now gender-norm the testos values 
subtract min value within each group from individual testos levels for each subject
scale this by dividing the diff between max and min values withn each group
obtain a percent of maximum possiblity score for each participant
Cohen 1999 (from Dominic 2015)
```{r}
min_female <- testos_min_max$min[testos_min_max$DEM_3_GENDER_CHILD == 1]
min_male <- testos_min_max$min[testos_min_max$DEM_3_GENDER_CHILD == 0]

# then calculate diff max and min 
diff_female <- testos_min_max$max[testos_min_max$DEM_3_GENDER_CHILD == 1]- testos_min_max$min[testos_min_max$DEM_3_GENDER_CHILD == 1]
diff_male <- testos_min_max$max[testos_min_max$DEM_3_GENDER_CHILD == 0]-
  testos_min_max$min[testos_min_max$DEM_3_GENDER_CHILD == 0]


# substract the min for each sex - so min - min will = zero. 
# then also divide the result by the diff of that sex.
# will yield a perentage! 
testos_df$testos_cleaned_log_wave_ave_gender_normed <- with(testos_df,
  ifelse(gender.f == "female",
  (testos_cleaned_log_wave_ave - min_female) / diff_female, 
  ifelse(gender.f == "male", (testos_cleaned_log_wave_ave -min_male)/diff_male, NA)))


summary(as.numeric(testos_df$testos_cleaned_log_wave_ave_gender_normed))
# why are there 3 NAs? those subjects migt not have testos? yes. or they were outliers. 

hist(testos_df$testos_cleaned_log_wave_ave_gender_normed)
```

## graph the tranformed data 
```{r}

ggplot(testos_df, aes(x = testos_cleaned_log_wave_ave,
                               y = testos_cleaned_log_wave_ave_gender_normed,
                               color = gender.f)) + 
  theme_bw() + geom_jitter (alpha = 0.5) #+ facet_grid (~ gender.f)

```

## conduct group differences with the normed data 
still a group difference!! 
```{r}

testos_gender_int <- lmer(testos_cleaned_log_wave_ave_gender_normed ~ GROUP + gender.f  + batch + 
                          + corrected_cort_age_yrs + index_wave +
                       (1 | IDENT_SUBID), data = testos_df)

anova(testos_gender_int)

```

## merge the normed testosterone things with the data
```{r}
# keep only the testosterone data 
testos_merge <- testos_df %>% 
  select(IDENT_SUBID, index_wave, testos_cleaned_log_wave_ave, testos_cleaned_log_wave_ave_gender_normed)

# remove the old testos (uncleaned) data from the giant cortisol df 
cort_merge7_no_testos <- cort_cleaned.no.outliers %>%
  select(-starts_with("testos"))

# now left join, so that the testos-wave-averages get added to the cortisol data 
cort_merge8 <- left_join(cort_merge7_no_testos, testos_merge, by = c("IDENT_SUBID", "index_wave"))

save(cort_merge8, file ="../../../data/1_cort_data/1D_long_cort_testos_cleaned.Rdata")
```
