---
title: "Growth chart for Hippocampus volume ages 4+"
author: "Michelle.VanTieghem"
date: "October 16, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---


# Readme:
conducting main analysis, removing wave3 (different scanner) 

```{r, warnings = F, include = F, message = F}
source("../../0_R_analysis_setup_file.R")
keep_legend <- theme(legend.position= c(0.90, 0.10),  legend.key.height = unit(0.05, "cm"), legend.key.width = unit(0.05, "cm"), legend.background = element_rect(color = "black"))
```

# load data
```{r}
load("../../../data/2_fs_data/3_cleaned_fsdata_for_growth_chart_brain_age_yrs4_19.Rdata")

fsdata6 <- fsdata6 %>%
  filter(index_wave < 3)
```

# Hippocampus VOLUMES
## plots of raw data for left hipp
```{r, warnings = F}
 ggplot(fsdata6, aes(x =brain_age_yrs, y = Left.Hippocampus, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) +  stat_smooth() + 
  my_colors + my_colors2
```


## test differences between hemispheres? 
### if not, averbrain_age_yrs them and use for subsequent analyses 
```{r, warnings = F}
RA <- ggplot(fsdata6, aes(x = Right.Hippocampus, y = Left.Hippocampus, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth(method = 'lm') + 
  my_colors + my_colors2
RA
```


## Bilateral Hippocampus 
probably quadratic in comps. 
```{r, warnings = F}
ggplot(fsdata6, aes(x =brain_age_yrs, y = Hipp_ave, color = GROUP)) +
  geom_point( ) + theme_classic() + 
  geom_line(aes(group = IDENT_SUBID), alpha = 0.5) + stat_smooth() + 
  my_colors + my_colors2

```

# fix up variables
## rescale variables 
for coefficients to be interpretable
```{r}

fsdata6$Hipp_ave <- fsdata6$Hipp_ave/100
# NOTE: ICV NEEDS TO BE RESCALED FOR SIMILAR SD RANGE AND INTERPRETABLE COEFFICIENTS
fsdata6$ICV <- fsdata6$ICV/100000

```


## mean center stuff

```{r}
fsdata6$brain_age_yrs.c <- fsdata6$brain_age_yrs - mean(fsdata6$brain_age_yrs)
fsdata6$sex.c <- fsdata6$sex - mean(fsdata6$sex)
# main effect of age =squared 
fsdata6$age_squared <- fsdata6$brain_age_yrs^2
fsdata6$age_squared.c <- fsdata6$age_squared - mean(fsdata6$age_squared)
fsdata6$ICV.c <- fsdata6$ICV - mean(fsdata6$ICV)
fsdata6$meds.c <- fsdata6$ICV - mean(fsdata6$ICV)
fsdata6$motion_ave_new.c  <- fsdata6$motion_ave_new - mean(fsdata6$motion_ave_new)

```


# AGE MODELS FOR Hippocampus
do brain_age_yrs and group interact to predict change in Hippocampus? 
using lmer, mixed effects - random intercepts for subject only. 


## linear model with age X group interaction
```{r, warnings = F}
# same for ICV. 
hipp_linear <- lmer(Hipp_ave ~ brain_age_yrs *GROUP + sex.c + ICV.c 
                     + motion_ave_new.c + 
                 (1 | IDENT_SUBID), 
                 data = subset(fsdata6, brain_age_yrs < 20))


summary(hipp_linear)
save(hipp_linear, file = "model_output/hipp_no_wave3_linear_model_results.Rdata")

```


## main effects linear model - no interaction
```{r, warnings = F}
# same for ICV.
hipp_linear_main_effects <- lmer(Hipp_ave ~ brain_age_yrs + GROUP + sex.c + ICV.c 
                     + motion_ave_new.c + 
                 (1 | IDENT_SUBID), 
                 data = subset(fsdata6, brain_age_yrs < 20))


summary(hipp_linear_main_effects)
save(hipp_linear_main_effects, file = "model_output/hipp_no_wave3_linear_main_effects_model_results.Rdata")

AIC(hipp_linear_main_effects)
```


## Calculate ICC for mixed effects
```{r, warnings = F}
# from simple model 
mod <- lmer(Hipp_ave ~ (1 | IDENT_SUBID), data = fsdata6)
icc(mod)
ICC_hipp <- icc(mod)[1]
save(ICC_hipp, file = "model_output/ICC_for_hippocampus.Rdata")


# from final model with group main effect
ICC_hipp_full_model <- icc(hipp_linear_main_effects)
ICC_hipp_full_model
save(ICC_hipp, file = "model_output/ICC_for_hippocampus_full_model.Rdata")


# very similar!! 
```


## graph model fit with raw data
note: plotting 2 lines, but NOT plotting results of NS interaction. 
```{r, warnings = F}

#  graph the model and confirm the post-hoc tests!! 
effect_hipp <- as.data.frame(effect("brain_age_yrs:GROUP", 
                                    hipp_linear_main_effects, confint=TRUE))

hipp_linear_plot_with_raw <- ggplot(effect_hipp, 
                                    aes(x = brain_age_yrs, y = fit, color = GROUP)) +
   geom_line(size = 1) + #facet_grid (~GROUP) + 
   theme_classic() + ylab(bquote('Hippocampus Volume '*(mm/100)^3*'')) + 
  xlab ("Age (years)") + my_colors + my_colors2 + 
   # add raw data points! 
   geom_point(data = fsdata6, aes(x = brain_age_yrs, y = Hipp_ave, color = GROUP),
              size = 1, alpha = 0.2) + 
   geom_line(data = fsdata6, aes(x = brain_age_yrs, y = Hipp_ave, group = IDENT_SUBID), 
             alpha = 0.2) + 
  scale_x_continuous(breaks = seq(4,21.5, by = 2), limits = c(4, 21.5), expand = c(.02,0)) +
 scale_y_continuous(expand = expand_scale(mult = c(0, .1))) +
  theme(legend.position="none")

pdf (file = "figures/Hipp_no_wave3_linear_model_with_raw.pdf", width = 6 , height = 4)
hipp_linear_plot_with_raw
dev.off()

hipp_linear_plot <- hipp_linear_plot_with_raw +  geom_ribbon(aes(ymin=lower, ymax=upper, 
                  group = GROUP, fill = GROUP), alpha =0.1, color = NA)
pdf (file = "figures/Hipp_no_wave3_linear_model_with_raw_CI.pdf", width = 6 , height = 4)
hipp_linear_plot
dev.off()
```

SAVING TO POWERPT 
```{r}

#ppt_test2 <- read_pptx() %>%
 # add_slide(layout = "Title and Content", master = "Office Theme") %>%
  #ph_with_vg(code = print(hipp_linear_plot)) %>%
  #print(target = "../../../manuscript/figures/powerpoint_vector_figures/Feb_2020/Hippocampus_linear_model_with_raw_CI.pptx")
  
```

